% This is part of the Avaneya Project Crew Handbook.
% Copyright (C) 2010, 2011, 2012
%   Kshatra Corp.
% See the file License for copying conditions.

% Preface chapter...
\StartChapter{Preface}

The {\it Avaneya Project Crew Handbook} is a guide for new and current contributors to the Avaneya project, a cerebral science fiction game for GNU. 

This document will likely continue to be revised and improved with time. Therefore, because it is a living document, it is probably better to pass on a \href{http://bazaar.launchpad.net/~avaneya/avaneya/trunk/view/head:/Documentation/Contributors/Handbook/Avaneya\%20Project\%20Crew\%20Handbook.pdf}{link} to the latest revision within our Bazaar repository\footnote[]{See \in{section}[Revision Control Management] for details in accessing the Bazaar source repository.} than it is to send a static locally stored copy to another person.

This handbook had several purposes:

\startitemize[4]
\item
To force those responsible for the game's inception to further refine their ideas to the point where they could be solidified on paper;

\item
To define the project as clearly as possible so that everyone involved in contributing understands it;

\item
To provide the project with a definition in a single, consolidated, canonical location;

\item
To politely consider those already engaged in the project from having to explain those things redundant with the handbook.\footnote[rtfm]{It is usually a good idea to ask questions. But as we say in the hacker community, albeit in a slightly more colourful form, consider reading the documentation first.}
\stopitemize

This book is probably not as useful to normal users as it is to contributors. It may even spoil elements of the game for the former, so consider yourself warned. On the other hand, normal users are just as likely to enrich the game through constructive feedback which is best buttressed with a well informed understanding of the project.

It is also important to bear in mind that it is not necessary for users to be familiar with the relevant details contained within this handbook anymore than it is for patrons of theatre to be familiar with the details stage technicians working with lighting and acoustics are. However, if you happen to be a stage technician, you need to be.

\StopChapter

