% This is part of the Avaneya Project Crew Handbook.
% Copyright (C) 2010, 2011
%   Kshatra Corp.
% See the file License for copying conditions.

% Time section...
\StartSection{Time}

Time keeping in the game is different from what you are used to on Earth. You need to know this in order to understand at least the story's chronology and the user interface. Although Terrans encountered in the game would likely use their native system, for the sake of consistency and to minimize confusion to the user, all fictional dates of the future are represented in the Arcadian format described below.

\StartSubSection{Prime Meridians & Timezones}

% Timezones...
A prime meridian\index{prime meridian} is an imaginary longitudinal line of reference on a planet that is used as the basis for offsetting its timezones\index{timezones}. On Earth, this is the Coordinated Universal Time (UTC)\index{Coordinated Universal Time} overlying Greenwich, London. On Mars, the crater {\it Airy-0}\index{Airy-0} had been used as the prime meridian since 1969 A.D.. This latter convention was established based on early photographs from the Mariner 6 and 7 missions\index{Mariner missions}.

After asserting their independence,\footnote{See \in{section}[Civil Disobedience and False Flags] for more Arcadian efforts at independence.} Arcadians changed the prime meridian they were using on Mars to their native city of Arcadia instead of {\it Airy-0}, the latter being roughly on the other side of the planet. Even with the change, however, there still remained 24 timezones as before. The difference meant that the zones would now have to be offset relative to Arcadia, and thus Arcadian clocks are set to {\it Coordinated Mars Time\index{Coordinated Mars Time}\index{MTC}} (MTC) since the time of its declaration of independence.

\StartSubSection{Seconds, Minutes, Hours, Days, and Years}

% Explain seconds, minutes, and hours...
A Martian second, minute, and hour are still the same as their Terran counterparts. That is, a Martian hour is still the same as 3600 Terran seconds.

% Explain solar day...
A {\it sol\index{sol}} is a solar day. This is the Martian analogue to a Terran day, but slightly longer by 2.7 \%. This conversion means that there are 1.027491 Terran days for every Martial sol. Arcadian clocks \quote{time slip}\index{time slip} at midnight for 39 minutes, 40 seconds to makeup for the difference. This allows its residents to use the 24-hour clock they were originally accustomed to on Earth.

% Explain year...
A Martian year, abbreviated {\it MYr} and pronounced "m-year", contains 668.6 sols, or 689 days by a terrestrial metric. That means there are 88775.245 seconds in a Martian year with 1.8876712 Martian years for every Terran year (1 year, 320 days, and 18.2 hours). These years are enumerated relative to a temporal datum as described in \in{chapter}[Timeline].

% Solstice and Equinox subsection...
\StartSubSection{Solstice and Equinox}

A solstice happens twice a year when the Sun's apparent position in the sky over Mars, like Earth, reaches its northernmost or southernmost extremes. The equinox also happens twice a year when the tilt of its axis is inclined neither away from nor towards Sun, but laying directly within the plane of the planet's equator.

% Calendar subsection...
\StartSubSection{Calendar}

The colonists use a different calendar than on Earth for both political and practical reasons. The political reasons are two fold. The original Arcadians were predominantly secular and the Gregorian calendar's year zero has a theological significance to many. The other reason being that they coveted a calendar which was meaningful in the context of {\it their own} indigenous history.

The practical reason is simple. The Gregorian calendar is useless in a world where the orbital period, seasons, and length of a solar day are different than those of Earth. The calendar's months, for instance, are not meaningful in the absence of the natural cycles of the Earth's Moon. As a consequence, they needed a calender that is meaningful to the physical idiosyncrasies of their own planet.

Had an attempt at employing the equipartitioned Terran months on Mars been made, one would have realized that they could never be of nearly equal length as they are on Earth because the Martian orbit has a high degree of orbital eccentricity\index{orbital eccentricity}. In orbital mechanics, this is the amount an orbital path deviates from a perfect circle to form an ellipse. Kepler's second law\index{Kepler's second law} states that the line adjoining a planet with the Sun will sweep through equal areas in equal time. This means its seasons cannot be of equal length because the planet speeds up as it gets close to the Sun and then slows down again as it becomes distant.

The Arcadian solution to the problem of the Martian calendar was inspired by 20\high{th} century aerospace engineer, Dr. Robert Zubrin.\footnote{See \in{definition}[definition:Mars Direct] for more on Zubrin.} He noted that since the Martian seasons, and therefore months, are not of equal length, to be useful, the calendar must be divided into equal segments that correspond to equal angles about the Sun. This enables Arcadians to not only predict the seasons, but to also keep the familiar concept of twelve months, each one therefore still remaining as a \math{30^{\circ}} (\math{360^{\circ} / 12 = 30^{\circ}}) arc. 

Where they parted from their Terran counterparts was in the frame of reference. In ancient times, Terrans of a geocentric view named the months after whatever zodiacal constellation the Sun, as seen from Earth, appeared to be positioned in. The Arcadians, now having to reconcile the fact that Earth is no longer the only place our solar system harbours human life, instead relied upon a more universal heliocentric perspective in using the location of Mars in the zodiac as seen {\it from} the Sun, as opposed to the Terran inverse. This is possible because the zodiac is physically meaningful to any planet within our solar system since all of the planets it contains lie in the same orbital plane. The zodiac, therefore, could further be relied upon to furnish the Arcadian calendar with the names of its months. The result of these calculations are contained in \in{table}[table:Martian Months].
\crlf

\placetable[force][table:Martian Months]{The Martian calendar of twelve months.}
{
    \bTABLE[split=repeat,option=stretch]
    \setupTABLE[column][5]
        [width=.52\textwidth,
        align=yes]
    \setupTABLE[row][each][align=center]
    \setupTABLE[5][1][align=center]

\bTABLEhead
    \bTR[bottomframe=on]
      \bTH  Month \eTH
      \bTH  L\low{s}\footnote{This is the range of angles Mars makes with respect to the Sun, as seen from the Sun, within the given month.} \eTH
      \bTH  First\\Sol \eTH
      \bTH  Total\\Sols \eTH
      \bTH  Notes \eTH
    \eTR
\eTABLEhead

\bTABLEbody
    \bTR
      \bTC Gemini \eTC
      \bTC \math{[0^{\circ}, 30^{\circ})}\eTC
      \bTC 1 \eTC
      \bTC 61 \eTC
      \bTC Beginning of the year begins with the vernal equinox on Gemini 1 (beginning of spring)\eTC
    \eTR
    
    \bTR
      \bTC Cancer \eTC
      \bTC \math{[30^{\circ}, 60^{\circ})} \eTC
      \bTC 62 \eTC
      \bTC 65 \eTC
      \bTC  \eTC
    \eTR
    
    \bTR
      \bTC Leo \eTC
      \bTC \math{[60^{\circ}, 90^{\circ})} \eTC
      \bTC 127 \eTC
      \bTC 66 \eTC
      \bTC Aphelion on Leo 24 (farthest from Sun)\eTC
    \eTR
    
    \bTR
      \bTC Virgo \eTC
      \bTC \math{[90^{\circ}, 120^{\circ})} \eTC
      \bTC 193 \eTC
      \bTC 65 \eTC
      \bTC Northern hemisphere's summer solstice on Virgo 1 (Sun at its northernmost)\eTC
    \eTR
    
    \bTR
      \bTC Libra \eTC
      \bTC \math{[120^{\circ}, 150^{\circ})} \eTC
      \bTC 258 \eTC
      \bTC 60 \eTC
      \bTC  \eTC
    \eTR
    
    \bTR
      \bTC Scorpius \eTC
      \bTC \math{[150^{\circ}, 180^{\circ})} \eTC
      \bTC 318 \eTC
      \bTC 54 \eTC
      \bTC  \eTC
    \eTR
    
    \bTR
      \bTC Sagittarius \eTC
      \bTC \math{[180^{\circ}, 210^{\circ})} \eTC
      \bTC 372 \eTC
      \bTC 50 \eTC
      \bTC Autumnal equinox on Sagittarius 1 (Sun over equator)\eTC
    \eTR
    
    \bTR
      \bTC Capricorn \eTC
      \bTC \math{[210^{\circ}, 240^{\circ})} \eTC
      \bTC 422 \eTC
      \bTC 47 \eTC
      \bTC Dust storm season starts \eTC
    \eTR
    
    \bTR
      \bTC Aquarius \eTC
      \bTC \math{[240^{\circ}, 270^{\circ})} \eTC
      \bTC 469 \eTC
      \bTC 46 \eTC
      \bTC Perihelion on Aquarius 16 (nearest to Sun)\eTC
    \eTR
    
    \bTR
      \bTC Pisces \eTC
      \bTC \math{[270^{\circ}, 300^{\circ})} \eTC
      \bTC 515 \eTC
      \bTC 48 \eTC
      \bTC Northern hemisphere's winter solstice on Pisces 1 (Sun at its southernmost) \eTC
    \eTR
    
    \bTR
      \bTC Aries \eTC
      \bTC \math{[300^{\circ}, 330^{\circ})} \eTC
      \bTC 563 \eTC
      \bTC 51 \eTC
      \bTC Dust storm season ends \eTC
    \eTR 

    \bTR
      \bTC Taurus \eTC
      \bTC \math{[330^{\circ}, 360^{\circ})} \eTC
      \bTC 614 \eTC
      \bTC 56 \eTC
      \bTC End of year on Taurus 56 \eTC
    \eTR 
\eTABLEbody

\eTABLE
}

As an example, a date expressed by an Arcadian might be {\it 7 Virgo, 32 A.R.}. This denotes the seventh day of the fourth month, Virgo, of the thirty-second Martian year following Arcadia's assertion of independence. See \in{chapter}[Timeline] for a description of the suffix that followed the year.

