% This is part of the Avaneya Project Crew Handbook.
% Copyright (C) 2010-2013 Cartesian Theatre <info@cartesiantheatre.com>.
% See the file Copying for details on copying conditions.

% Euphemisms section...
\StartSection{Euphemisms}

\StartTimelineDate{9 Scorpius, 3 B.R.}
Arda attends a ceremony commissioning the first settlement fire brigade.
\StopTimelineDate

% At 1,000 new arrivals every year, each Tarikin capable of transporting 500 
% immigrants twice every year, this would take about ten Terran years or 5.6 
% MYrs...
\StartTimelineDate{14 Scorpius, 3 B.R.}
The Arcadian Settlement's annual census notes a population now in excess of 10,000 colonists and growing.
\StopTimelineDate

\StartTimelineDate{22 Scorpius, 3 B.R.}
Originally published by the {\it RANND Corp}, Edward Geisler's seminal work, {\it Transplanetary Commerce and Mars: The Red Horn of Plenty}, is hailed by industry leaders for its influence as the day is marked with the first of a series of private launches {\it en route} to Mars. Beginning from Kiruna, European Union, a heavy lift spacecraft bears a payload of both personnel and cargo to be received at a starport in the Arcadian Settlement. The application for commercial activity was validated by UNSA several years in advance.

The launch is celebrated by various industry tycoons as a landmark transformation as the first transnational corporations promote themselves to become the first transplanetaries. The principle stakeholders consist of an international private consortium representing a variety of sectors ranging from soft drink vendors and agricultural biotechnology to aerospace.

Geisler's paper discussed the theoretical means and opportunities of achieving off--world commerce with Mars and the strategies to offset the initial high risks. It draws on the historical analogy of Europe's expansion into the \quote{New World} of the Americas and the great wealth that was generated for export.
\StopTimelineDate

\StartTimelineDate{1 Sagittarius, 3 B.R.}
Arda receives a UNSA directive to terminate several of Henrik's orthomolecular medical research programs occupying settlement laboratory space. In response to Henrik's request for clarification, a UNSA liaison declines to respond in writing, but informally indicates intentions to lease the vacated facilities to pharmaceutical giant GlaxoKlineSmith.
\StopTimelineDate

\StartTimelineDate{41 Sagittarius, 3 B.R.}
Arda receives a UNSA directive to terminate the settlement's kombucha brewery and cultured foods research program. All equipment and most other material assets are to remain behind in the vacated facilities. 

Although consumption of the \quote{Arcadian Manowar}\index{Arcadian+Manowar}\footnote{The Portuguese man o' war is a species of jellyfish ({\it Physalia physalis}) from which the beverage draws its name. It is said to resemble the \quote{mushroom} used in the drink's fermentation.} beverage had been a cultural practise in the settlement for years due to the simplicity in preparation, its widespread adoption had been preceded with Henrik's discovery that those exposed to higher than average cosmic radiation, which is to be expected during extended EVAs, had on average benefited from its consumption. His research conducted within settlement laboratories using {\it in vitro} trials provided compelling evidence that the fermented tea aided in the repair of radiation--induced cellular DNA damage. 

Until now, the beverage had been a minor export back to Earth where it had long been prized, widely sought after, and considered superior to Terran variants. The biochemistry of the archaic assortment of bacteria and yeast that acts as a symbiotic colony for its fermentation, as used for centuries on Earth, but metabolized now on Mars under a lower gravitational field, was a common explanation in Terran urban legends to explain its unique properties.

The difficulty in obtaining a genuine bottle from a very limited stock drawn from a marketplace already saturated with Asian Union counterfeits ensured it remained rarely more than an expensive novelty item on Earth. This had been the case since the first railgunned\footnote{See \in{section}[Railguns] for an explanation and diagram of a railgun.} imports were received on Earth and celebrated symbolically as the first off--world consumables. Since that time the possession became symbolic of personal wealth and social status on Earth. With the termination of the settlement's program, the diminishing modest stocks remaining on Earth were expected to be auctioned off at significant value to private collectors.
\StopTimelineDate

\StartTimelineDate{22 Capricorn, 3 B.R.}
Arda receives a UNSA directive to reduce Nayana's greenhouse water usage. The savings are intended to offset the requirements of a recently opened Coke bottling facility leasing the site of the settlement's former cultured foods research facility.
\StopTimelineDate

\StartTimelineDate{47 Capricorn, 3 B.R.}
Arda receives a UNSA directive to terminate all of Henrik's double blind herbal research studies being conducted by settlement laboratories conducting orthomolecular research. UNSA cites the NAU Congress and EU Parliament having recently passed legislation declaring herbs currently under settlement research, such as St. John's Wort, echinacea, and a number of others, as controlled substances.

Henrik pleads with UNSA to ban GlaxoKlineSmith and all other lobbyists from influencing UNSA policy. He relies on several hundred academic publications authored by his laboritories to validate the value of their research, but is unsuccessful.
\StopTimelineDate

\StartTimelineDate{30 Aquarius, 3 B.R.}
Arda receives a UNSA directive to provide logistical support to recently landed Clown Food and subsidiary contractors requiring logistical assistance in erecting several kilometers of facilities to be used in the production of genetically modified beef, pork, and potatoes. The production is essential to provide for the supply chain of the newly expanded restaurant franchise across the settlement and neighbouring starport.

Nayana requests Arda's intervention. Arda pleads with UNSA to reconsider their decision. He points out that the availability of tillable land required for warm--blooded herbivores for human consumption is not efficient, but must be created through a process which is resource intensive. He adds that most of the energy used in their raising would be radiated off into space as heat long before it could be of any use as food. He stresses that resource management is critical and the margin for error much smaller than on Earth, and therefore Clown Food's requirements would place an undue burden on the Arcadian Settlement.

UNSA does not respond directly. Arda attempts to initiate revocations of the contractors' visas. UNSA overrides the revocation, privately reprimands him, and councils his compliance with the original directive.
\StopTimelineDate

\StartTimelineDate{4 Pisces, 3 B.R.}
Arda receives a UNSA directive to terminate several of Henrik and Nayana's mushroom ranches. The ranches were used to recycle plant waste into high quality sources of edible protein. The energy used in maintaining ranch temperature is reallocated for railgun use. 

Nayana advises Arda that all plant waste destined for the ranches, or as composted greenhouse fertilizer, must now be transported to newly created outdoor garbage fills across the planitia. Outdoor atmospheric conditions and cosmic radiation ensure the sterilized organic matter will neither decompose nor become reusable.
\StopTimelineDate

\StartTimelineDate{19 Aries, 3 B.R.}
Arda receives a UNSA directive to secure all new settlement energy requirements from private energy contractor Haliburton. Haliburton will provide UNSA with nuclear reactors that will superannuate the settlement's standard practise of expanding primarily through new geothermal wells and photovoltaic panels. The directive specifies that these purchases are to be absorbed into the already allocated settlement's Terran sustainability research budget. 
\StopTimelineDate

\StartTimelineDate{51 Taurus, 3 B.R.}
Arda receives a UNSA directive to terminate several settlement research laboritories currently investigating Terran sustainability. Some of these include key photobioreactor farms of Henrik's that produce kelp fertilizer for greenhouse use. The resources freed are to be redirected to increased deuterium production for Terran export and its requisite energy demanding railgun use. Arda has no choice and reluctantly complies with the directive. UNSA cites technical reasons while Arda cites what he suspects is settlement public debt in private to his confidants.
\StopTimelineDate

