% This is part of the Avaneya Project Crew Handbook.
% Copyright (C) 2010-2013 Cartesian Theatre <info@cartesiantheatre.com>.
% See the file Copying for details on copying conditions.

% Civil Disobedience & False Flags section...
\StartSection{Civil Disobedience & False Flags}

\StartTimelineDate{1 Gemini, 2 B.R.}
The beginning of a new year is marked with the first of a long series of partisan meetings held in secret. The clandestine symposium, self titled {\it Free Arcadia},\index{Free Arcadia} discusses a number of grievances with Earth and the prospects for settlement independence. Arda, Henrik, Khalid, Nayana, and Senka are in attendance. Leonard is not directly informed.
\StopTimelineDate

\StartTimelineDate{4 Pisces, 1 B.R.}
The settlement holds an inauguration ceremony following Arda having won a landslide election. He accepts executive office as President of a newly formed interim settlement government of limited self--imposed independence. Arda and his staff operate from within the newly erected {\it Arcadia Hall}. UNSA acknowledges neither the interim government nor the occurrence of an election, though rumours on Earth surface following a high level leak to alternative media purporting emergency briefings being conducted within the North American Union's National Security Council.

Leonard strongly cautions Arda in private against the {\it "formation of a redundant administration, exceeding the mandate, and any action which may undermine the Office's strategic interests or sovereignty."} In New York the United Nations Security Council immediately convenes the first of a series of emergency meeting on the situation as it continues to unfold.
\StopTimelineDate

\StartTimelineDate{7 Pisces, 1 B.R.}
Promoted from the position of United Nations Deputy Secretary--General, Security Council Resolution 12661 is adopted appointing Leonard to the newly created \quote{\it Office of the Governor of the Arcadian Settlement}. Though UNSA does not explicitly acknowledge the election in the wording of the resolution, it mandates that the \quote{\it Governor's seniority shall precede all other Arcadian Settlement Public Officials and Civil Servants.} Arcadia Hall does not release a public statement in response to the resolution.
\StopTimelineDate

\StartTimelineDate{50 Taurus, 1 B.R.}
Arda puts the contentious issue of settlement independence before a public referendum of 15,343 colonists. The {\it Mars Enhanced Telecommunications Orbiter} is taken temporarily offline by {\it Bronfman--Murdoch Aerospace} personnel on Earth citing routine maintenance. The vote carries at 98\%, but the news consequently does not return to Earth through mainstream media.
\StopTimelineDate

\StartTimelineDate{51 Taurus, 1 B.R.}
The asteroid {\it 52048 Varuna} aerobreaks into the Martian atmosphere, disintegrating as predicted. Meteor showers are reported impacting the south pole's dry ice sheets by the Terraformation Team.
\StopTimelineDate

\StartTimelineDate{55 Taurus, 1 B.R.}
Arcadia Hall issues a public advisory to alert the settlement of Arda having gone missing. The advisory is recalled when Nayana sets out by rover and locates Arda drafting a document under the light of a late night desert aurora.
\StopTimelineDate

\midaligned{\bf \star \hspace[star_hspace] \star \hspace[star_hspace] \star}

% Year zero...
\StartTimelineDate{1 Gemini, 0 A.R. (Year Zero)}
Intending to reach the general public of both planets, Arda formally addresses the Secretary--General of the United Nations through an open letter also available on the solnet.

\startTimelineCorrespondenceDocument
    \input Timeline/Arda_Independence_Letter.tex
\stopTimelineCorrespondenceDocument

Arda's publicly accessible transmission over the solnet is accompanied with copies of all settlement research to date. He also includes a copy of the {\it Rubicon Act}. 

The {\it Rubicon Act}, also known more formally as the {\it Constitution of the Republic of Arcadia Planitia}, is the first of the settlement's self--declared legislation. Its aim is to initiate and secure the colony's independence. The document was the product of years of underground drafting through covertly held partisan meetings and, if successful, would beget the first off--world autonomous city--state in human history. The cultural significance to Arcadians is such that their calendar embraces the event as its year zero epoch. Summaries of notable portions follow.

\startTimelineGeneralDocument
    \input Timeline/Rubicon_Act_Summary.tex
\stopTimelineGeneralDocument

A failed denial--of--service attack lodged against the Terran solnet downlink station at UNSA's Jet Propulsion Laboratory on Earth attempts to disrupt reception of Arda's transmission. Khalid advises the station's technicians that the attack originated from within the Terran station's own local intranet from behind a restricted access zone using several zero day vulnerabilities in the station's transceiver software. His findings are published and popularized in {\it Phrack},\index{Phrack} an electronic periodical for hackers.

Arda's independence letter propagates rapidly over peer to peer networks. Responses on Earth among Terrans following Martian events through independent media are predominantly in favour of the settlement's commitment to independence. Mainstream media sensationalize and relate the situation as a costly mission in crisis at the mercy of a Mission Commander of questionable intentions and character.
\StopTimelineDate

\StartTimelineDate{2 Gemini, 0 A.R.}
Arda arrives by dirigible at {\it Valles Marineris}\index{Valles Marineris}, a 4,000 km long canyon reaching up to 7 km in depth. He spends the sol in solitary reflection.

The former NASA had long ago salvaged aluminium recovered in 2001 A.D. from the ruins of the former World Trade Centers in lower Manhattan for reuse in their Mars rover {\it Spirit's} cable shield. To remind future Arcadians of the reasons for independence, Arda has the {\it Constitution} engraved in aluminium recovered from the derelict craft as it came to finally rest at its UNESCO World Heritage Site on Harrowdown Hill, Mars. The plate is kept at Arcadia Hall on public display where it becomes known as the {\it Lamothe Plate} in reverence to the late Dr Richard Lamothe, a renowned Nobel prize winning French astrophysicist, areologist, and author whose every volume is among Arda's private collection.

NASA had originally selected the metal for reuse with {\it Spirit} as a symbol of perseverance in the face of evil. Arda, convinced both the metal and the message it embodied was still relevant, selected it for reuse in the Lamothe Plate.
\StopTimelineDate

\StartTimelineDate{14 Gemini, 0 A.R.}
A Survey Team is deployed on a routine assignment using ground based shallow radar to scout for artesian aquifers. Travelling solo, Arda departs Arcadia to rendezvous later with the team as previously scheduled. Midway he performs an EVA to dismount the rover and check one of its tires. While outside, he is knocked from his feet as an explosion occurs on the other side of the rover from within the cabin. Although his forearm is shattered, he sustains no life threatening injuries. With the rover's first aid kit destroyed, he improvises a make shift caste around the suit's elasticized mesh by applying liquid water to solidify coatings of supportive frozen regolith around the wounded limb.

Arda inspects the rover to find the cabin fully depressurized and the airlock's seal too badly damaged to maintain positive pressure. The rover's electronic controls are non--responsive, rendering the machine derelict in the middle of a Martian desert. The nearest rescue effort that acknowledges his emergency VHF transponder is an aquifer drilling rig crew at least 42 hours away. With only 35 hours of liquid oxygen available for his respirator, he extends it in time by syphoning cryogenic liquid oxygen from the rover's propulsion system.

The damaged rover is later towed back to a service bay in the city. Senka finds fragments of a pressurized container with a gaseous methane / oxygen signature that had been rated at 35,000 kPa. The mechanics rule out the rover's own equipment since its fuel cells are methanol based and no other known onboard standard equipment required methane. 

In an effort to determine who last serviced the vehicle, Senka notes that the electronic maintenance log is missing. Remnants in the cabin of what appear to be fragments of a timer based detonator are recovered with its components not matching any of the rover's equipment.
\StopTimelineDate

\StartTimelineDate{10 Cancer, 0 A.R.}
Arcadia Hall amends the Constitution by passing the {\it Humanoid Act}, stripping corporate legal fictions of the rights of human beings and denying money as a form of free speech. Transplanetaries with registered offices within Arcadia's territorial region do so now without personhood legal protection.
\StopTimelineDate

% Coke pulled out of Vietnam the year before the United States military did...
\StartTimelineDate{42 Cancer, 0 A.R.}
{\it Coke} quietly closes down its production and bottling facilities. Several other transplanetaries, including the successful fast food franchise {\it Clown Food}, are seen at Starport Arcadia making preparations to return to Earth. None of the transplanetaries preparing to return to Earth issue public statements.
\StopTimelineDate

\StartTimelineDate{62 Cancer, 0 A.R.}
No. \#2 Railgun is taken offline for unscheduled rail and insulator surface repair by Terran UNSA contractor, {\it Solar Urban Moving Systems}. Senka and her Mechanics Team are not informed.
\StopTimelineDate

\StartTimelineDate{20 Leo, 0 A.R.}
A longshoreman\index{longshoreman} at No. \type{#}2 Railgun's blockhouse reports an explosion that has taken out their compulsator power supply. The single largest loss of life ever to occur on Mars unfolds as two--hundred and fifty three people are killed and seven more injured as an adjacent residential district's inflatable dome in Arcadia's Fourth Precinct depressurizes from shrapnel. Henrik and staff rapidly triage the surviving casualties, rushing them to the city's hospital.

Arda convenes an emergency meeting with all state secretaries present at Arcadia Hall. Forensic evidence of a cryobomb's methane / oxygen chemical residue is disclosed. Analysis suggests that at least the detonator is of probable non--Martian origin. 

Before the meeting adjourns, Terran media implicate the alleged identities of the \quote{railgun bombers} as Free Arcadia partisan extremists with ties to Arda.

Shortly after the meeting adjourns, Arda begins reading a presidential address from Arcadia Hall covering his staff's preliminary findings. However, the live broadcast is cut short prematurely from reaching a Terran audience due to technical difficulties with the {\it Mars Enhanced Telecommunications Orbiter}. Khalid tries to re--establish the uplink but discovers the satellite used to broker nearly all communication between the two planets has been remotely taken offline for an unscheduled systems diagnostic by {\it Bronfman--Murdoch Aerospace} contractors at UNSA's Jet Propulsion Laboratory on Earth. Terran mainstream media networks reassure the Office of the President that Arda's presidential address will be rescheduled for replay later simultaneous the Terran Superbowl.

Leonard consults with Arda privately and advises him to step down from his political position of President in light of the optics and amidst growing pressure from UNSA. However, he maintains that UNSA still supports his role in a strictly technical capacity of Mission Commander due to the influence and respect he wields across the colony. Arda raises that while this is true, the real reason is because his compliance is the most reliable means of overseeing the immediate resumption of stable deuterium exports in a minimum time frame. The Governor ends his counselling with the necessity that Arda \quote{\it dismantle all indigenous self--governance apparatus immediately due to his inability to no longer guarantee the safety of the colonists}.
\StopTimelineDate

\StartTimelineDate{21 Leo, 0 A.R.}
Arcadia Hall brings forward evidence before the newly elected and sworn in Arcadian Justice of the Peace with allegations that Leonard is a NAU--CIA asset. Justice Liau withdraws to his chamber to review the material. The judge issues an arrest warrant and authorizes the confiscation, revocation, and destruction of Leonard's visa and red UN Laissez--Passer diplomatic passport.

Arcadia Hall directs ground crew at Starport Arcadia to begin fuelling {\it ERV Bravo}, Earth--bound, taking advantage of a minimum--energy launch window. Arda deputizes the first peace officers of a newly formed RAP police detachment, superannuating the minimal United Nations contingent of deputy North American Union Marshals present to date.

Deputy Commissioner Knudsen leads a security detail of the newly formed detachment. Leonard is identified, presented with the warrant, resists arrest, but is taken into custody while being read his rights. The irate Governor is escorted to the starport for subsequent deportation. An automated solo departure lifting off later that evening will return him back to Earth with non--emergency flight controls locked down from manual user override.
\StopTimelineDate

\StartTimelineDate{31 Leo, 0 A.R.}
Nayana on--site at the south pole with her Terraformation Team report findings of dry ice sublimation. Her team concludes this is likely in response to {\it 52048 Varuna's} impact.
\StopTimelineDate

\StartTimelineDate{33 Leo, 0 A.R.}
Footage of alleged riots erupting across Arcadia with security forces responding disproportionately are aired across Earth. Mainstream media attribute the alleged riots to grievances held by dissenters at variance with \quote{\it extremist Free Arcadia partisans}. According to the line taken by the networks, the riots began in response to Arcadia Hall's alleged use of force to suppress political dissent. The alleged malcontents and dissenters are coined \quote{\it Red Unionists}\index{Red Unionists} and advocate a return to UNSA sovereignty.

Several critics note the equivalence of what appears to be Terran gravity in portions of the footage. They are rapidly dismissed by mainstream media as \quote{\it cranks} and \quote{\it conspiracy theorists}.
\StopTimelineDate

