% This is part of the Avaneya™ Project Crew Handbook.
% Copyright (C) 2010-2017 Cartesian Theatre™ <info at cartesiantheatre dot com>.
% See the file Copying for details on copying conditions.

% Countdown section...
\StartSection{Countdown}

\StartTimelineDate{38 Leo, 0 A.R.}
The United Nations Security Council passes resolution 12664 where it is adopted unanimously. Announcement of the resolution is not yet made public, but is to be deferred until 28 Scorpius. The resolution contains the following:

\startTimelineGeneralDocument
The Security Council,

Reaffirming the principles and purposes of the Charter of the United Nations,

Determined to combat by all means threats to transplanetary peace and security caused by terrorist acts,

Deploring the gross and systematic violation of human rights, including the repression of peaceful demonstrators, expressing deep concern at the deaths of civilians, and rejecting unequivocally the incitement to hostility and violence against the civilian population and infrastructure excited through rogue elements within the United Nations Arcadian Settlement's local administration,

Recognizing the inherent right of individual or collective self--defence in accordance with the Charter,

\startitemize[R]
\item Unequivocally condemns in the strongest terms the horrifying terrorist attacks which took place recently at the site of the No. \type{#}2 Railgun on Mars in the United Nations Arcadian Settlement and regards such acts, like any act of international terrorism on Earth, as a threat to transplanetary peace and security;

\item Expresses its deepest sympathy and condolences to the victims, their families, and of the United Nations Arcadian Settlement;

\item Calls on all Mars Treaty signatories to work together urgently to bring to justice the perpetrators, organizers and sponsors of these terrorist attacks and stresses that those responsible for aiding, supporting or harbouring the perpetrators, organizers and sponsors of these acts will be held accountable;

\item Calls on all Mars Treaty signatories to redouble their efforts to prevent and suppress terrorist acts including by increased cooperation and full implementation of the relevant anti--terrorist conventions and Security Council resolutions;

\item Expresses its readiness to take all necessary steps to respond to the terrorist attacks of the Arcadian Settlement's railgun bombing, and to combat all forms of terrorism, in accordance with its responsibilities under the Charter of the United Nations;

\item Authorizes Mars Treaty signatories that have notified the Secretary--General, acting nationally or through regional organizations or arrangements, and acting in cooperation with the Secretary--General and the Office of the President of the Space Agency, to take all necessary measures to protect civilians, civilian populated areas, and assets under threat of attack in the United Nations Arcadian Settlement, and requests those signatories concerned to inform the Secretary--General immediately of the measures they take pursuant to the authorization conferred by this paragraph which shall be immediately reported to the Security Council;

\item Decides to remain actively seized of the matter.%\footnote{Most Security Council resolutions end with this phrase. Article 12 of the United Nations Charter states that while it \quote{\it is exercising in respect of any dispute or situation...the General Assembly shall not make any recommendation...unless the Security Council so requests.} In other words, the General Assembly of many is being instructed by the Security Council of few to not interfere in an affair the latter still wishes to handle exclusively for the time being.}
\stopitemize
\stopTimelineGeneralDocument

The resolution recommends that the {\it United Nations Department of Peacekeeping Operations (UNDPKO)} immediately place the multinational contingents under its permanently standing {\it Rapid Reaction Force} on high alert. Article 51 of the United Nations Charter is cited as the legal basis for the use of force which allows for \quote{\it the inherent right of individual or collective self--defence if an armed attack occurs against a Member of the United Nations.}
\StopTimelineDate

\StartTimelineDate{44 Leo, 0 A.R.}
The {\it Non--Aligned Movement}\index{Non--Aligned Movement} issues a \quote{\it Statement to the United Nations Security Council on the Situation Between the United Nations and the Republic of Arcadia Planitia}. The statement recalls the principles of \quote{\it non--intervention, non--interference, the sovereignty of the Arcadian people, and their right to self--determination.}

The organization is an association of Terran states advocating the doctrine of non--alignment. They define this as \quote{\it respect for independence, sovereignty, territorial integrity, and security of states as they strive for freedom from imperialism, all forms of colonialism, racism, foreign aggression, occupation, domination, and interference.} The constituents include the states of Venezuela, Iran, Ecuador, Cuba, Syria, and others.
\StopTimelineDate

\StartTimelineDate{2 Virgo, 0 A.R.}
The United Nations Department of Peacekeeping Operations' {\it Joint Chiefs of Staff} assemble in Brussels, European Union, where they liaise with government officials and other high level military and UN servants for the first day of a three month summit. The summit is conducted under maximum security with no access to the press.

Though the summit's existence is made public, the high degree of discretion exercised attracts public attention. The lack of transparency draws widespread criticism from the press due to the presence of a number of high level state officials from across the world.

Rumours rapidly promulgate purporting the purpose of the summit as one of planning the first off--world military campaign in history. Leaked partially redacted documents suggest the summit's agenda is to assess the feasibility, strategic planning, intelligence gathering, logistical, personnel, economic, and political implications to conduct such an operation. Known high level UNSA liaisons, including incumbent president Dr. Samuel Lieberman, are captured on camera being hustled in and out of the summit's secured area in unmarked vehicles.

Public speculation is increased when representatives of major international financial institutions are caught on camera leaving the vehicle parkade by independent journalists. 

The NAU--CIA director's presence is later confirmed when a local escort service discloses to municipal level media the director's refusal to settle a bill for unpaid services at a hotel the director was residing in. For the duration of the summit the hotel was known to be exclusively restricted to summit attendees and their support staff for security reasons. In response to media inquiries, the agency issues a press release reiterating generic public relations statements on the agency's commitment to national security.
\StopTimelineDate

\StartTimelineDate{27 Libra, 0 A.R.}
UNDPKO's summit in Brussels concludes with the quiet establishment of the newly formed {\it Mars Command (MARSCOM)}. This is established under the civilian administered UNDPKO and is now responsible for overseeing all of the latter's peacekeeping operations at the strategic level, in transit to, in the vicinity of, and on the Red Planet. The headquarters is located at Peterson Air Force Base, North American Union, and is jointly staffed by an international consortium of military personnel from all nations signatory to the Mars Treaty.

A large volume of procurement requests for military hardware are disseminated to a multitude of defence and aerospace contractors.
\StopTimelineDate

\StartTimelineDate{34 Libra, 0 A.R.}
INTERPOL issues a \quote{red notice} international arrest warrant for Arda.
\StopTimelineDate

\StartTimelineDate{36 Libra, 0 A.R.}
Lockheed Martin--Boeing's Skunk Works advanced research facilities begin refitting UNSA's {\it Tarikin III} personnel ferry into {\it Kali}, a militarized remotely piloted spacecraft carrying a payload of personnel and materials. This is in response to a UNSA request for quotation for a classified mission payload specification drafted according to MARSCOM's Joint Chiefs of Staff.
\StopTimelineDate

\StartTimelineDate{22 Scorpius, 0 A.R.}
Selected Rapid Reaction Force personnel are assigned to North American Union's School of the Americas and UNSA training facilities in Antarctica and Spain. The latter facilities are now administered by MARSCOM.

A prominent senior level civilian scientist stationed at the now military administered UNSA training facility publicly addresses the {\it Antarctic Treaty Secretariat}\index{Antarctic Treaty Secretariat} headquartered in Buenos Aires, South American Union. Dr. Ishmael expresses concerns that Article I of the {\it Antarctic Treaty System}\index{Antarctic Treaty System} of 1961 A.D. prohibits the use of \quote{\it any measures of a military nature, such as the establishment of military bases and fortifications, and the carrying out of military manoeuvres in Antarctica.} The Secretariat releases a public statement acknowledging the build--up of military personnel and expresses her concern by reaffirming her commitment to the treaty with its official depositary, the North American Union.

The North American Union's Secretary of State\index{North American Union Secretary of State} responds by clarifying that the UNSA training facilities are still under a civilian mandate, and therefore a military presence for peaceful purposes remains consistent with treaty obligations.
\StopTimelineDate

\StartTimelineDate{23 Scorpius, 0 A.R.}
Arda receives an encrypted message warning of an imminent plan drafted by a consortium of influential Terran interests. The informant claims to have allegedly drawn the information from leaked source material made available through software vulnerabilities in the proprietary firewall protecting UNSA intranets. 

The anonymous tip--off claims these interests have influence over the United Nations and warn that the latter is mobilizing to land a brigade of peacekeepers to Martian theatre. The brigade ({\it MARSBRIG}) would operate under UNSA's newly formed MARSCOM, which itself operates as a newly formed organ of the United Nations, which in turn, according to the informant, answers to these private interests.

Field commanders on the ground are to be tasked with several prioritized strategic objectives while in theatre. Among the digital dossier provided by the informant, Arda locates a memorandum outlining all of the alleged objectives described in detail in the file {\tt COSMIC_TOP_SECRET_PROTECTED_C-000521-MJCS-SETREC}.

\startTimelineCorrespondenceDocument
COSMIC TOP SECRET
DISTRIBUTION RESTRICTED / FOR NAMED RECIPIENTS ONLY
MARSCOM JOINT CHIEFS OF STAFF
\color[colour_item]{MEMORANDUM ON PRIORITIZED STRATEGIC OBJECTIVES FOR ARCADIAN SETTLEMENT\crlf RECOVERY}

\startitemize[n]
    
    \item The Joint Chiefs of Staff have considered the enclosed Memorandum for UNSA which responds to a request of the Executive Office of the Secretary-General for brief but precise descriptions of methods to be exercised in the reclamation of the Arcadian Settlement through military intervention.
    
    \item The Joint Chiefs of Staff recommend that the enclosed Memorandum be forwarded as a preliminary submission suitable for planning purposes. It is assumed that there will be similar submissions from other agencies and that these inputs will be used as a basis for developing a time-phased plan. Individual projects can then be considered on a case-by-case basis.
    
    \item Further, it is assumed that a single agency will be given the primary responsibility for developing military and para-military aspects of the basic plan. It is recommended that this responsibility for both overt and covert military operations be assigned to the Joint Chiefs of Staff.
    
\stopitemize

\startitemize[4]

\item \color[colour_item]{ALPHA} -- TERMINATE ATATÜRK'S COMMAND\footnote{See \in{section}[ATATÜRK] for an explanation of this cryptonym given to Arda.}

    \startitemize[n]
    \item HIGH VALUE TARGET. IF CAPTURED ALIVE, PERFORM IMMEDIATE EXTRACTION TO KALI, AREOSTATIONARY ORBIT. HOLD IN CUSTODY. AWAIT EXTRAJUDICIAL RENDITION TO EARTH.
    \item TERMINATE W/ EXTREME PREJUDICE IF UNABLE TO CAPTURE ALIVE. IMPLICATE RED UNIONISTS. MEDIA PREBRIEFED. BHADRA 2 COMMSAT WILL STANDBY TO BROADCAST NEWS TO EARTH.
    \stopitemize

\item \color[colour_item]{BRAVO} -- CAPTURE ARCADIA
    \startitemize[n]
    \item INSERT NEW PLAINCLOTHES NAU-CIA, SPEC FORCE, AND BLACKWATER ASSETS INTO ARCADIA TO SUPPLY INTEL.

\page

    \item RULES OF ENGAGEMENT AUTHORIZE NEUTRALIZATION OF ANY AND ALL LOCAL RESISTANCE BY ANY MEANS NECESSARY. MISSION AIM TIME CONSTRAINTS REQUIRE NO QUARTER. EXPECT MINIMAL, IF ANY, FIGHTING IN BUILT UP AREAS. INTERNATIONAL HUMANITARIAN LAW IN CONTEXT OF INTERNAL ARMED CONFLICT NON-APPLICABLE TO NON-STATE ACTORS, INCLUDING SIERRA GUERRILLA FORCES.
    \item DISARM SIERRA AND CONFISCATE ALL SMALL ARMS.
    \item PERFORM ENHANCED INTERROGATION OF SUSPECTED CAPTURED FAP\footnote{Free Arcadia partisans.} TO LOCATE SIERRA WEAPONS' CACHES AND OTHER INTEL. EXTRAJUDICIAL DETENTION, TRIAL, AND EXECUTION WITHIN MINIMAL TIME FRAME ESSENTIAL. ATATÜRK EXCEPTION, VIZ OBJECTIVE ALPHA(ONE).
    \item COMMAND AND CONTROL OF STARPORT ARCADIA CRITICAL WITH ECONOMY OF EFFORT.
    \item SEIZE MAX NO OF MATERIAL ASSETS INTACT.
    \item RESTORE UNSA ADMIN, BROKERED VIA MARSCOM INITIAL COMMAND AND CONTROL UNTIL SITN STABILIZED.
    \stopitemize

\item \color[colour_item]{CHARLIE} -- EXTRACT ALREADY DEPLOYED INTEL ASSETS
    \startitemize[n]
    \item EXTRACT AND DEBRIEF NAU-CIA INTEL OPS CURRENTLY IN THEATRE SINCE ARRIVAL 28 MONTHS PRIOR VIA TARIKIN III FLIGHT SERIAL TA15.
    \stopitemize

\item \color[colour_item]{DELTA} -- ECONOMIC REFORMS
    \startitemize[n]
    \item REINSTATE KEY SIERRA LABOUR AND PERSONNEL NECESSARY FOR RESUMPTION OF STABLE DEUTERIUM AND OTHER CORE EXPORTS.
    \item PROTECT COMBAT SUPPORT SQUADRON'S ERECTION OF 'FIRST CENTRAL BANK OF THE ARCADIAN SETTLEMENT'.
    \item REVERT SETTLEMENT'S LEGAL TENDER FROM JENYA TO BANCOR.
    \item CAPTURE ALL RHODIUM HELD IN RESERVE.
    \stopitemize

\item \color[colour_item]{ECHO} -- POLITICAL REFORMS
    \startitemize[n]
    \item IMPLEMENT MARTIAL LAW W/ SUFFICIENT TIME TO ACHIEVE ALL STRATEGIC OBJECTIVES.
    \item ASSIST NAU-CIA ASSETS ENGINEER ARCADIA HALL ELECTION FOR OFFICE OF THE PRESIDENT. ALL BALLOT OPTIONS MUST BE COMPROMISED W/ ANY FRIENDLY CANDIDATE FROM SCHEDULE D.
    \stopitemize
\stopitemize

\page

IN UNLIKELY EVENT MARSBRIG UNABLE TO HOLD ARCADIA, EMERGENCY CONTINGENCY PLAN'S NEW STRATEGIC OBJECTIVES FOLLOW:
\startitemize[4]
\item \color[colour_item]{FOXTROT} -- INSERT AND ACTIVATE FIFTH COLUMN SLEEPER CELL NETWORK.

    \startitemize[n]
    \item ALL ASSETS, EXCEPT OBJECTIVE FOXTROT(TWO)'S, FALLBACK TO RENDEZVOUS TANGO AND CONFIRM LOC REP W/ GRID REF OVER CRYPTOCODE JULIET. ORBITAL FIRESUPPORT AVAILABLE. AWAIT EXTRACTION TO KALI IN AREOSTATIONARY ORBIT AND STANDBY FOR FURTHER ORDERS.

    \item UPON OBJECTIVE BRAVO(ONE), THESE ASSETS REMAIN BEHIND IN SIERRA SAFE HOUSES AS RESIDENT FIFTH COLUMN CONTRAS. IF POSSIBLE, W/O COMPROMISING COVER, FIFTH COLUMN WILL JAM SIERRA'S COMM AND PROTECT RETREATING MARSBRIG ASSETS.

    \item FIFTH COLUMN CONDUCTS INSURGENCY OPS TO EXEC FALSE FLAGS AS RED UNIONISTS. IF COMM UPLINK BROKEN WILL CONDUCT OPS INDEP W/O NEED FOR LOG RESUPPLY. PROVIDES MARSCOM W/ CLANDESTINE UNIT BEHIND ARCADIAN DEMARCATION LINE FOLLOWING UNLIKELY STRAT OR TACTICAL RETREAT.

    \item IF OBJECTIVE DELTA(THREE) FAILS, FIFTH COLUMN WILL RAISE FUNDS IN LOCAL JENYAS TO MAINTAIN OPS. SIERRA'S SEIZED LABS POSSIBLY CONDUCIVE TO DOMESTIC SYNTHETIC NARCOTIC PRODUCTION. SYNTHETIC IN SITU METHOD PROVIDES PLAUSIBLE DENIABILITY AND GREATER RELIABILITY W/O NEED FOR TERRAN IMPORT, BYPASSING STARPORT AUTHORITY / CUSTOMS.
    \stopitemize
\stopitemize
\stopTimelineCorrespondenceDocument

The anonymous informant provides detailed information on troop movement to Terran MARSCOM administered training facilities, size of units, weapons and tactics to be employed, probable location of encampment, and route of approach. The brigade will be deployed to Martian theatre under the banner of the {\it United Nations Mars Expeditionary Peacekeeping Force (UNMEPF)}. The force draws elements mostly from North American Union and European Union units specializing in airborne and mechanized infantry, armoured reconnaissance, tactical UAVs, self--propelled artillery, and logistical support.

The message is encrypted using a public key of Arda's not used since his long departed days of a military posting with the European Union's Mountain and Commando Brigade in Hakkari. Khalid determines the digital signature to be authentic. Arda and his cabinet deliberate before concluding the message to be credible.
\StopTimelineDate

\StartTimelineDate{24 Scorpius, 0 A.R.}
Khalid advises Arda that he can take out UNSA's orbital surveillance capabilities by destroying the {\it Mars Enhanced Telecommunications Orbiter} while the settlement still has access. Arda issues an executive order to do so. UNSA detects a settlement login attempt and transmits a command sequence instructing the orbiter to revoke Arcadia's credentials but is several minutes too late. The orbiter burns up its entire hydrazine attitude control fuel reserves, deorbits, and incinerates during atmospheric entry.

All solnet communications between Earth and Mars that had been brokered by METO are now effectively disabled. However, the orbiter had also been used for reliable Arcadian intercommunication. Khalid improvises in METO's absence by refitting key settlement communications equipment to combine line--of--sight VHF and, where that fails, ionosphere reflected shortwave communications. Neither method requires a satellite.

UNSA has the Jet Propulsion Laboratory lock down all four {\it Mars Positioning Satellites} and places them in safe mode in anticipation of an RAP denial--of--service attack, though none comes. Arda's staff reason the satellites have no use as UNSA surveillance assets and are still useful for their own navigation on the Red Planet where ground based transponders are still unavailable in many remote regions.
\StopTimelineDate

\StartTimelineDate{27 Scorpius, 0 A.R.}
Dr. Ishmael is found dead within his residence at the UNSA facility in Antarctica. The Garrison Commander's\index{Garrison Commander} chain of command authorizes him to release a redacted coroner's report attributing the cause of death to \quote{\it haemorrhage due to incised wounds of the left wrist with blood chemistry indicating an overdose of painkillers} -- but rules out foul play. 

Ishmael's wife, a physician, who had been living at the facility with her husband is interviewed by independent media where she claims the incisions were non--fatal, that there was insufficient blood on the scene as expected from a major artery, that the report did not indicate any liquids at the scene that could have been used to consume painkillers, and that the dosage had been non--fatal.

Mounting public pressure leads to the Garrison Commander\index{Garrison Commander} authorizing the release of additional previously redacted details. The new documents show alcohol had been present at the scene. Ishmael's wife responds that he was a devout Muslim who never consumed alcohol, but even if he had, his blood chemistry analysis in the toxicology report showed no alcohol.
\StopTimelineDate

\StartTimelineDate{28 Scorpius, 0 A.R.}
The United Nations Secretary--General's press security holds a press conference where the Security Council's previously passed resolution 12664 to invade Mars is made public for the first time. The timing coincides with the saturation of Terran media with a recent celebrity adultery scandal. Scarce media coverage that does take place highlights Terran heads of states citing the importance of the United Nations doctrine of the {\it responsibility to protect}, with emotional references to alleged Martian atrocities.
\StopTimelineDate

\StartTimelineDate{29 Scorpius, 0 A.R.}
Arda holds an emergency cabinet meeting at {\it Arcadia Hall} in secrecy. Several hours into the meeting, Arda summons a group of three of his best geologists and three of his best civil engineers. They are tasked in private before departing on rovers to fulfill a high priority assignment overseen by Senka.
\StopTimelineDate

% Expediency is plausible if partisan meetings had been held at such a location where surveying had already been done in the past...
\StartTimelineDate{30 Sagittarius, 0 A.R.}
Arcadia Hall issues a city--wide emergency evacuation order\index{evacuation order} that is carried out over the next two weeks. All non--essential personnel are transported to an unmarked installation constructed within a subterranean lava tube. The installation is heated and powered by a geothermal power station with a supply of artesian aquifer liquid water.
\StopTimelineDate

% If Hohmann Type II, assume 8.5 Terran months since he left Mars...
\StartTimelineDate{48 Sagittarius, 0 A.R.}
{\it ERV Bravo} carrying Leonard returns to Earth landing at Edwards Air Force Base, North American Union. The spacecraft is seized immediately. Leonard is debriefed before departing for the City of London\index{City of London} the following evening.
\StopTimelineDate

\StartTimelineDate{50 Sagittarius, 0 A.R.}
Arcadia completes its evacuation of all non--essential personnel. The movement of non--essential materials and Arcadia Hall's administrative staff out of the city to the underground installation follows immediately. A make--shift facility for skeleton staff and functionality of Arcadia Hall is erected within the shelter and given the same name.

Excavation begins for the construction of unpressurized underground passageways linking the shelter outside of the city with airlocks within it. Usage of tunnel entrances within the city avoids obvious vehicle tracks from orbital telemetry leading to the location of the underground installation outside of it. 

Senka predicts the tracks created by excavation equipment still present on the surface will be erased through surface wind deposition and erosion during the seasonal dust storm if one should occur. This normally occurs during the orbital perihelion when Mars is at its closest to the Sun.

Within the underground installation, and with most of the city evacuated, Arda gathers the settlers to provide an explanation for the recent commotion and events leading to their evacuation. Influenced by Vattel's {\it Law of Nations}, Arda compels them that it is within their \quote{natural rights} the option of collective self--defence since, regardless of whether they decide to take up arms or not, they must realize that they will soon enough be confronted with someone else's.

Arda reminds the settlers that their constitution already permits the raising of militias. A referendum is held on the raising of several units to provide protection for Arcadians and the city. The settlers are generally reluctant, but concede to it as a last resort. The vote carries with a majority.

Arda explains to his citizens that since prior military service is not unusual in the field of manned space exploration, putting out a call for all such settlers would be a reasonable place to start. A list is assembled with more than a hundred personnel responding.
\StopTimelineDate

\StartTimelineDate{2 Capricorn, 0 A.R.}
Relying on a combination of military training manuals remaining on Arcadian local solnet caches, personal experiences, knowledge of the current state of Arcadia's industrial infrastructure, such as milling and robotic assembly machines, Arda begins the first of several day's with his cabinet on a measured response to an expected confrontation with a United Nations military force.

The sol marks the first of a series of meetings held over the next several months on the recruitment, organization, and equipping of an indigenous militia. The meetings are held by an {\it ad hoc} advisory group called the {\it City Security Committee}. The CSC consists of settlement personnel with experience ranging from industrial processes, ex--military, mechanics, various managerial roles, chemists, engineers, and others. Acting as a think tank, they are formed as an advisory group.

Arda reasons that Arcadia has a minimum of nine months to prepare for the arrival of a military force, assuming the latter experiences no delays in the equipping, training, launch, and arrival. If they take longer the militia can expand its capabilities as much as time permits. To console doubts Arda further reasons that the United Nations would not assume the risk of deploying a military force across the solar system with all the political and economic liabilities that accompany it without the necessary confidence that the settlers would not be capable of defending themselves. He likens MARSCOM's overconfidence to an Achilles' heel.

The {\it Arcadian Militia} is established on paper, but its creation does not extend beyond that. Arda assumes responsibility over the null portfolio as Commander--in--Chief and begins with authorizing the creation of the first {\it Arcadian School of Asymmetric Warfare\index{asymmetric warfare}} with professional development courses available to personnel of all ranks under the leadership of former military personnel.
\StopTimelineDate

\StartTimelineDate{3 Capricorn, 0 A.R.}
With settlement approval Arda authorizes the use of subterranean areas of Arcadia for the garrison of several militia units yet to be created. These would include everything from drill halls and supply depots to training facilities. The units planned are few in number but reflect the settlement's current capabilities and resources within the minimum time remaining before the expected landing of the United Nations brigade.

\startitemize[4]
\item {\it 1\high{st} People's Own Arcadian Mechanized Infantry}. An infantry regiment to be equipped with armoured personnel carrier rovers. Its purpose is to seek out, close ranks with, and neutralize the enemy's capabilities.

\item {\it 2\high{nd} People's Own Arcadian Dragoons}. An armoured regiment to provide armoured reconnaissance support.

\item {\it 3\high{rd} People's Own Arcadian Combat Support Engineers}. An engineering battalion to provide heavy equipment, signals (communication), and logistical support.
\stopitemize
\StopTimelineDate

\StartTimelineDate{12 Capricorn, 0 A.R.}
Plans for a feasible service rifle capable of operating under the extreme external environmental conditions Mars presents are proposed before the {\it City Security Committee} by a team of highly skilled machinists. The plans are well received. 

The R1A1 service rifle is a gas--operated, self--loading, air--cooled, magazine--fed, shoulder--fired, kinetic energy weapon. It is a modernized adaptation of the former Terran Coldwar era FN FAL. Everything required for its simple and rapid manufacture is already at Arcadia's disposal.

The receiver and heavy barrel are made of an alloyed magnesium. Though the original design called for heavier steel, the magnesium is lighter and when alloyed allows for flexibility. These metal components are treated with a protective finish that matches in tan colour the sands of {\it Arcadia Planitia}. Even with the original weapon weighing only a third on Mars what it would on Earth, the design modifications still render a weapon that is lighter.

The stock is modified to use a lightweight composite hemp plastic. The material is stiff, has a high heat tolerance, flame retardant, and incidentally biodegradable. The stock contains a battery power source to provide energy to an integrated heating element within the receiver to reduce misfires due to frosting.

The weapon features both magnesium alloy sights and an ROA1 detachable optical scope for target acquisition, heavily adapted for Martian ballistics where the Terrestrial point blank needs to be moved from a 200 metre mark up to 700 metres. The front and rear sights bear small sealed tritium markings that provide a small glow to aid in night sighting. The optical scope has night vision, thermal imaging, and laser range finder capabilities.

The receiver is chambered to the already well field tested Terran 7.62x51mm round. A magnesium alloy is selected for the cartridge casing. The ejection port is modified with a cover which the original design did not have to reduce the likelihood of very fine Martian iron oxide dust as small as 1.5μm from entering the breech and fouling the action.

The gas regulator calls for another departure from the original design to manage the greater gas differential of the high chamber pressure contrasted with the near vacuum of Mars at the end of the muzzle.

An RGA1 anti--personnel breech loaded launcher is optionally mounted beneath the barrel and can chamber a variety of cartridges. These include magnesium flares, smoke grenades, or fragmentation \quote{cryonades}. Cryonades are small explosive shells that contain a volatile mixture of cryogenic liquid oxygen and methane. Armourers are instructed that these fragmentation cartridges must be stored when not in use at cryogenic temperatures to prevent them from becoming inert through degassing.

Although the original design that formed the basis for this new adaptation is antiquated, the weapon is relatively simple to manufacture, operate, and already well proven in the extreme arctic and high altitude climates of Earth. The committee notes that this is important since these climates share the closest environmental similarities with Mars. The weapon should be able to operate in temperatures as low as \math{-90^{\circ}}C. The committee suggests that any frosting issues that may arise will have to be addressed later with field trials and special chemically engineered lubricants.
\StopTimelineDate

% They need a strategic transport...
\StartTimelineDate{33 Capricorn, 0 A.R.}
{\it Kali's} launch window is missed due to low--Earth orbit having become increasingly populated with dangerous space debris\index{space debris}. UNSA engineers fear they may be approaching the Kessler effect\index{Kessler effect} of causing a runaway chain reaction which would reduce all objects in orbit. Since this would potentially threaten not only {\it Iterum Shipyard}, but all Terran space missions, {\it Kali's} trans--Martian injection is rescheduled for the next available launch window of 25 months.
\StopTimelineDate

\StartTimelineDate{35 Capricorn, 0 A.R.}
The first non--experimental production ready R1A1 service rifles emerge from the assembly line and are delivered to unit quartermasters. The weapons' receivers are engraved with an Ancient Persian\index{Ancient Persia} maxim, {\it ride well, shoot straight, and speak the truth}. Arda's receiver is stamped with serial {\it 109450}.
\StopTimelineDate

% Should occur during a perihelion season...
\StartTimelineDate{16 Aquarius, 0 A.R.}
In keeping with Senka's prediction, most of Mars, including {\it Arcadia Planitia}, is enveloped in a seasonal dust storm. The day marks the planet's orbital perihelion where it is nearest to the Sun.
\StopTimelineDate

% Kali takes a free-return trajectory taking 180 days and drops off personnel and material. Should take place on 33 Capricorn, 0 A.R. + 25 months for the rescheduled launch from Earth + 180 days of transit...
\StartTimelineDate{11 Cancer, 2 A.R.}
{\it Kali} completes its trans--Martian journey, aerobreaking into areostationary orbit. The ship carries a payload of remotely operated equipment destined for Phobos, the Sun--synchronous photographic reconnaissance satellite {\it Bhadra I},\footnote{In Hindu mythology, one of Shiva's servants and goddess of the hunt.} the areosynchronous communications satellite {\it Bhadra II}, MARSBRIG military personnel, as well as some civilian defence personnel and contractors. The combined human resources and equipment represent the {\it United Nations Mars Expeditionary Peacekeeping Force (UNMEPF)}.
\StopTimelineDate

% Two month after arrival in orbit to give Thor Battery time to finish construction...
\StartTimelineDate{9 Leo, 2 A.R.}
While in areostationary orbit, {\it Kali's} Commanding Officer issues warning orders down the ship's chain of command to prepare for insertion. Mechanized infantry company, platoon and section commanders are briefed before loading Mars Ascent--Descent Vehicles (MADVs). Armour and artillery follow.
\StopTimelineDate

