% This is part of the Avaneya™ Project Crew Handbook.
% Copyright (C) 2010-2017 Cartesian Theatre™ <info at cartesiantheatre dot com>.
% See the file Copying for details on copying conditions.

% Aftermath & Diaspora section...
\StartSection{Aftermath & Diaspora}

\StartTimelineDate{42 Leo, 2 A.R.}
A disarmed and dismounted UNMEPF brigade is escorted on both flanks by dual columns of Arcadian dragoons and infantry. All critically wounded {\it hors de combat} are triaged and transferred to an Arcadian hospital. The hospital staff are augmented with UNMEPF medical personnel, including former lawful combatants now under Arcadian custody. The dead temporarily remain behind frozen. The escorted prisoners of war are destined for Arcadia's underground emergency shelter to be exchanged with the previous inhabitants, the city's evacuees.

A minimally functional Arcadia Hall with skeleton staff is repopulated and oversees the return of evacuees from the shelter back to their homes.

The shelter, whose means of movement in and out are few in number, becomes a strategically ideal subterranean location to garrison UNMEPF prisoners of war. The latter's vehicles and weapons are dispersed across various Arcadian units.

Arda reassures Nilhara that, although he is confident UN forces would not have abided by the Geneva Conventions, it is an opportunity for Arcadia to demonstrate to both worlds that it is a legitimate, sovereign and civilized state in tending to the sick and wounded of either side without reproach or prejudice for race, religion, creed, or nationality.
\StopTimelineDate

\StartTimelineDate{45 Leo, 2 A.R.}
Nilhara agrees to provide Arcadia Hall with access credentials to the {\it Bhadra I/II} satellites. Arcadia's solnet connectivity with Earth is re--established.

Arda contacts the Secretary--General to establish a back door diplomatic communication's channel. From this sol and the eight that succeed it, the following terms are successfully negotiated over diplomatic cables. The concessions of the Security Council are as follows.

\startitemize[R]
\item The General Assembly shall acknowledge the Republic of Arcadia Planitia as a legitimate and independent state and shall not infringe upon its sovereignty.

\item The Republic of Arcadia Planitia shall be represented in the General Assembly on Earth with a representative of its own choosing. Arda is reassured that he will be given amnesty whether he chooses to fulfil that office himself or not, though he declines.

\item The United Nations shall cease peacekeeping operations in the proximity of, and within, Arcadia. The term \quote{hostilities} is never used. The Secretary--General, on behalf of the Security Council, does not unambiguously reaffirm any commitment to the prohibition of the deployment of all forms of weapon systems and munitions of war, suborbital, orbital, or in outer space, and therefore, does not directly acknowledge {\it Thor Outpost}. Questions of its use, consequently, remain quietly unresolved.
\stopitemize

The concessions of the Republic of Arcadia Planitia are as follows:

\startitemize[R]
\item All UNMEPF prisoners of war are to be treated humanely in accordance with the Geneva Conventions.

\item All UNMEPF prisoners of war are to be repatriated to Earth as soon as logistically feasible.

\item The Republic of Arcadia Planitia shall become a signatory to the Vienna Convention on Diplomatic Relations, allowing the United Nations to be represented with a permanent diplomatic mission in Arcadia. This shall include a single embassy, but possibly one or more consulates as agreed upon in the future.

\item A limited number of Terran corporations are permitted to resume commercial activity in Arcadia. With their registered offices on Earth, the status of corporate personhood within Arcadia remains unresolved.
\stopitemize

Arcadia Hall prepares formal requests to various Terran nations for the opening of embassies in the North American Union, European Union, Venezuela, Cuba, and others. In turn, it receives a submission of Leonard's credentials for entry into the UN's diplomatic mission in Arcadia. Arcadia Hall rejects Leonard's application as a {\it persona non grata} as permitted under Article 9 of the Vienna Convention.
\StopTimelineDate

\StartTimelineDate{55 Leo, 2 A.R.}
Arcadian forces transport UNMEPF personnel, including the dead, to the dropsite for Kali's MADVs. Arcadian immigration officials offer all UNMEPF personnel, including former combatants, the opportunity for Arcadian citizenship. Some accept and will remain behind. 

Senka assists with the MADV's refuelling and preflight checks. Arda returns Nilhara's sidearm, accompanied with a bottle of \index{Arcadian+Manowar}Arcadian Manowar. The pair salute. The spacecraft departs with Nilhara.
\StopTimelineDate

\StartTimelineDate{60 Leo, 2 A.R.}
Henrik is promoted to the Office of the Secretary of State. For his courage in the defence of the city, he is awarded the {\it Order of Arcadia}\index{Order of Arcadia}, the republic's highest civic and military honour -- and the first of it. Unit citations are further awarded to the {\it 1\high{st} People's Own Arcadian Mechanized Infantry}, {\it 2\high{nd} People's Own Arcadian Dragoons}, and the {\it 3\high{rd} People's Own Arcadian Combat Support Engineers}.
\StopTimelineDate

% Assume 180 days to return after launch...
\StartTimelineDate{40 Scorpius, 2 A.R.}
Kali's crew return to Earth. Nilhara is immediately taken into custody and debriefed. He is court--martialed on charges of having violated Article 104 of the Uniform Code of Military Justice in providing Arda with access credentials to the {\it Bhadra} satellites. He is found guilty of aiding enemy combatants and sentenced to life imprisonment without parole at the Midwest Joint Regional Correctional Facility, Forst Leavenworth, North American Union.
\StopTimelineDate

\StartTimelineDate{6 Capricorn, 2 A.R.}
Arda holds a public forum at Arcadia Hall on the future of the republic. The forum ends with a public vote carrying a motion to begin an Arcadian diaspora\index{Arcadian+diaspora} across Mars. This would begin with the settling of the rest of Arcadia Planitia. The rationale being that a nation comprised of a sole city has little hope of security and independence in a world still easily dominated by powerful Terran interests.
\StopTimelineDate

\StartTimelineDate{9 Capricorn, 2 A.R.}
An inbound subsonic aircraft crashes in the desert of {\it Arcadia Planitia} 40 km south of Starport Arcadia at 0700 MST. Emergency responders arrive on scene to find the dismembered wreckage bearing red cross emblems. The pilot is found without vital signs, pronounced dead, and the first officer missing. An attached RAP police constable locates 3600 kg of cocaine dislodged from within false bottom coffins bearing UNMEPF emblems.

The transportation secretary informs Arcadia Hall of the police investigators preliminary findings. The tail number was registered to an NAU-CIA contractor for UN diplomatic flights. Further, chemical analysis determined the origin of the narcotic payload to be of probable Martian origin of a more addictive and dangerous synthetic form.

The International Committee for the Red Cross provides documentation confirming that it had no such aircraft in its inventory, nor chartered by any of its affiliated organizations, and that such usage of the distinctive red cross emblem violated international law.

That evening Arcadia Hall liaises with the newly opened North American Union consulate in Arcadia and, as a matter of first official business, summons their diplomat.
\StopTimelineDate

