% This is part of the Avaneya Project Crew Handbook.
% Copyright (C) 2010, 2011, 2012 Cartesian Theatre <kip@thevertigo.com>.
% See the file Copying for details on copying conditions.

% Rubicon Act's notable portions summarized...
\startarticle
\item % Article I

...declares the colony an independent, sovereign, self--governed, constitutional republic with the right to self determination. It is self--styled the {\it Republic of Arcadia Planitia (RAP)}, colloquially known by its capital and sole city, {\it Arcadia}. The United Nations claims this area falls under its extraterritorial mandate and acknowledges it only as the Arcadian Settlement.

\item % Article II

...defines the head of state, the President. The President and his secretaries are democratically elected and serve for as long as the electorate permit. Since Arcadia is only a single city at this time, there is no large legislative assembly defined akin to a congress of sorts as sometimes found in the governance of nations of greater size.

\item % Article III

...discharges all of Arcadia's public debt currently held by the {\it International Monetary Fund} and the {\it World Bank}, ceasing interest payments effective immediately. A debt limit is imposed and the borrowing of money for public use can be done, but only against the public's credit and without accruing interest.

The Terran central banks had been using the settlers' registered biological property, that is, their UN issued extraterritorial birth certificates, and the security of the bearer's labour as monetized collateral against a debt that preceded the settlement's very existence. This was done without the knowledge of the bearer.

\item % Article IV

...defines an independent judiciary whose judges are democratically elected. They serve for as long as the electorate permit.
\stoparticle

\startarticle[start=7]
\item % Article VII

...superannuates the Terran bancor fiat currency with the {\it jenya}, a rhodium standard. This currency is composed of a platinum metal considered more rare and precious than gold. It is declared the exclusive legal tender within the Republic, necessitating that all Terran interests acquire Arcadian goods and services exclusively in jenyas, as opposed to in their own currencies of opportunity.

\item % Article VIII

...defines the state's disposition towards personal arms and a military. It provides that the right of the people to responsibly bear arms must never be infringed as part of the state's natural checks and balances. The underlying idea is that this is done principally as a safe guard for the protection of the people from either their own government, or that of another's -- but not for the sake of enabling vigilante justice.

The creation of a permanent standing military force at the government's disposal is strictly forbidden. The raising of militias is permitted, but conditional only upon the source of the chain of command's executive authority vested exclusively in none other than the people of the republic, not of government. 

Some of the militia's many restrictions are that it is forbidden from deployment outside of Arcadia's territorial regions; that its purpose is exclusively defencive -- of people, but of neither government nor corporations; that both its commissioned officers and enlisted ranks be drawn exclusively from Arcadian citizenry; that membership among these ranks remains strictly voluntary; that it shall not exercise a legal framework distinct from the civilian's in the discipline of its soldiers; and that it cannot be garrisoned or deployed domestically as an aid to civil power under any circumstance, real or speculative.

\item % Article IX

...denies the deployment of munitions of war, suborbital, orbital, in outer space, or anywhere outside of Arcadia's territorial regions.
\stoparticle

\startarticle[start=11]
\item % Articles XI to XXVI

...enumerates the fundamental freedoms and human rights Arcadians enjoy with respect to their new state. They are adaptations from Earth's former {\it European Convention on Human Rights} drafted in 1950 A.D.. These articles include the

    \startitemize[R,2*broad][start=11,before=\startlinecorrection,after=\stoplinecorrection]
    \setupitemize[left=(, right=)]
    \item[item:rubicon_act_fundamental_rights_first] right of everyone to their life;
    \item right of peaceful enjoyment of one's possessions;
    \item prohibition of torture;
    \item prohibition of slavery;
    \item right of liberty and security;
    \item right to a fair trial;
    \item prohibition of retrospective criminalization of acts;
    \item right of one's privacy;
    \item right of freedom of thought, conscience, and religion;
    \item right of freedom of expression;
    \item right of freedom of assembly;
    \item right of any of marriageable age to marry and establish a family;
    \item prohibition of discrimination;
    \item right to not be denied an education;
%    \item right to regular, free, and fair elections;
    \item[item:rubicon_act_fundamental_rights_last] prohibition of the death penalty.
    \stopitemize

\stopitemize

\startarticle[start=29]
\item % Article XXIX

...prohibits the state from providing itself with a not--withstanding clause, preventing a legal override of any part of the Act.
\stoparticle

