% This is part of the Avaneya Project Crew Handbook.
% Copyright (C) 2010-2014 Cartesian Theatre <info@cartesiantheatre.com>.
% See the file Copying for details on copying conditions.

% Timeline chapter...
\StartChapter{Timeline}
The world Avaneya takes place in is highly self--referential. A fictional timeline is positioned relative to a {\it year zero\index{year zero}}, with years preceding suffixed {\it B.R.\index{B.R.}}, before the republic; while those following are suffixed {\it A.R.\index{A.R.}}, after the republic. This is the convention adopted by Arda and his settlers at the time they assert their independence.

Note that a year is defined as a Martian year (MYr) and is not the same as a Terran year. You may want to refer to \in{section}[Time] for a detailed explanation of how Arcadians keep track of time on Mars.

Year zero's placement relative to the Gregorian calendar\index{Gregorian calendar} is deliberately never made explicit. Inferentially, however, one can reason that it is proximate enough for a contemporary audience to still feel relevant, yet distant enough to afford some creative license. This is important because, although Avaneya takes place in the future, it deals with many contemporary issues.

\startlines
\page \input Timeline/01_Promises_Made.tex
\page \input Timeline/02_Erecting_Home.tex
\page \input Timeline/03_Euphemisms.tex
\page \input Timeline/04_Civil_Disobedience_and_False_Flags.tex
\page \input Timeline/05_Countdown.tex
\page \input Timeline/06_Contact.tex
\page \input Timeline/07_Aftermath_and_Diaspora.tex
\page \input Timeline/08_The_Users_Predicament.tex
\stoplines

\StopChapter

