% This is part of the Avaneya Project Crew Handbook.
% Copyright (C) 2010-2018 Cartesian Theatre™ <info@cartesiantheatre.com>.
% See the file Copying for details on copying conditions.

% Audio section...
\StartSection{Audio}

\startitemize[4]

\head {\em libav / libasound_module_pcm_a52}

The {\tt libasound_module_pcm_a52} library is a plugin for ALSA and therefore only available on supported platforms  -- usually GNU/Linux systems. The plugin is a wrapper for the lower level encoder provided through {\tt libav}. The latter provides a way to encode a \index{surround sound:+digital}digital surround sound stream in A52 format to a pass--through device (e.g. S/PDIF). The user's home theatre receiver then decodes and amplifies the stream to the user's surround sound system.

We need to be careful for legal reasons. See \in{section}[Restricted Data Considerations] for more information on why.

%\head {\em libmikmod}

%We use libmikmod for playback of tracker music modules. Some people were requesting tracker support. Since it should not take very much effort to integrate into our engine's audio subsystem, as described in \in{section}[Audio], we considered this a reasonable request. One caveat is that we need to make sure the library is compiled thread safe.

\head {\em OpenAL}

OpenAL is used for 3D spatial audio rendering. Actual decoding of audio data is done through other APIs, such as {\tt SDL_audio}. Audio programming under GNU/Linux is unfortunately still a mess,\footnotecite[tuxradar2010_linux_audio_mess] but APIs like this help make it sane for game developers.

\head {\em SDL_audio}

This SDL module is used to provide audio decoding. It provides a standardized interface to various decoder backends, such as Ogg Vorbis and others.
\stopitemize

