% This is part of the Avaneya Project Crew Handbook.
% Copyright (C) 2010-2018 Cartesian Theatre™ <info@cartesiantheatre.com>.
% See the file Copying for details on copying conditions.

% Developer tools section...
\StartSection{Developer Tools}

\startitemize[4]
\head {\em Apport}

Apport intercepts program crashes. It collects debugging information about the crash, obtains anonymous information about the user's platform, and uploads to us in a standardized format by integrating directly into Launchpad. It is available on supported GNU operating systems.

\head {\em CppUnit}

The CppUnit library is a unit testing framework module for the C++ programming language. It is used to perform self diagnostics of various components of the game engine and tools. They are executed at build time.

\head {\em GNU Autotools}

The GNU Autotools are industry standard and akin to the \quote{construction scaffolding} surrounding a new building under development. They can be used to reconcile and harmonize the many idiosyncrasies of different platforms to make \index{portability} portability easier.

\head {\em GNU Compiler Collection}

The GNU Compiler Collection is used for the engine's dependency calculation, compilation, and linking. GCC is not a compiler, but actually a collection of compilers for different platforms. Ports of them are available for virtually every platform under the Sun, including every single one we are targeting.

\head {\em GNU MPFR}

The GNU MPFR Library allows us to control how floating point computations are done in C++. The goal is to make programs give reliable and reproducible results. This is important because differences in machine generated code, numeric handling libraries, dedicated hardware floating point processors and optimizations can yield results that are inconsistent across different environments.

A {\it libre} game is likely to be ported to a variety of hardware platforms. Players interacting over Solnet could be using very different platforms where operations on the same values produce different results. The MPFR library will ensure that calculations are performed consistently across disparate hardware. This way all players will play by the same rules.
\stopitemize

