% This is part of the Avaneya Project Crew Handbook.
% Copyright (C) 2010-2017 Cartesian Theatre™ <info@cartesiantheatre.com>.
% See the file Copying for details on copying conditions.

% freedesktop.org section...
\StartSection{freedesktop.org}

The freedesktop.org community is committed to providing specifications and applications for a consistent user experience across various desktop environments. This frees developers to focus on having to do things only one way, the standards compliant way, as opposed to many ways or having to make bad assumptions about the user's preferred desktop environment. 

\startitemize[4]
\head {\em D--Bus Specification}

The D--Bus (Desktop Bus) API is a simple inter--process communication (IPC) system for software applications to communicate with one another. Avaneya can use it for things like determining the state of the user's network connection. We currently use it heavily in the {\it Avaneya: Viking Lander Remastered DVD}.

\head {\em Desktop Application Autostart Specification}

The autostart specification allows applications to safely self launch when an appropriate medium is inserted. It could be traditional optical media or a USB stick as examples. This can be useful where user's are accustomed to a graphical user interface appearing on legacy systems after inserting an optical media for their favourite game.

\head {\em Desktop Entry Specification}

Several freedesktop.org compliant desktop environments have adopted this convention. This allows applications to provide \quote{desktop entries} describing where an application is, how it is to be launched, where it should appear in the user's menu, and localization strings.

\head {\em Media Player Remote Interfacing Specification}

MPRIS is a standardized interface for compliant media players. We could provide an option in Avaneya that stops the user's media player if it is currently playing music and then instruct it to resume playback on exit.

\head {\em Startup Notification Specification}

This specification defines a method that allows any compliant desktop environment to monitor Avaneya as it initializes and provide the user with feedback.
\stopitemize

