% The Characters chapter...
\startchapter{The Characters}

Avaneya's main characters aim to be rich in detail and multifaceted. Their character attributes are generally revealed through indirect means, such as body language, object symbolism, and dialogue. This may happen during a cinematic sequence, for instance.

\startsection{Arda Baştürk}
Arda was born 43 B.R., Istanbul, Turkey, to Dursun and Ayla Baştürk. His mother, Ayla, was a bank teller, and Dursun, an aeronautical engineer with the NAUSA Jet Propulsion Laboratory and later military intelligence officer.

Dursun had been active politically in opposing Turkey's membership in the European Union which happened shortly after Arda's birth. Several attempts were made on Dursun's life, culminating in a car bomb that left him paralysed before finally succumbing to his condition a year later.

Ayla was cautious in discussing the identity of the suspected assailants with Arda, as she did not want to risk putting her son in danger. Still, Arda sensed she had suspected that it had not been the work of marginalized extremists, but had been the work of elements within their own government on behalf of the Rothschilds.

As part of a working class family, with only his mother to support him, Arda had few opportunities that would not come without extending great effort on his part. 

In an era defined by scarcity, they were of modest means and she felt it important to continue imparting Dursun's philosophy of a positive work ethic. With food shortages rampant, high unemployment, and riots on a weekly basis in Istanbul, she knew it was his only ticket out, if he should have any at all.

Arda was described in his youth as highly perceptive and intelligent, although generally reserved, and further, reserving his intellectual gifts for his principle love, science. With the exception of what little he knew of his father, he felt the political landscape was largely the realm of {\it "bumbling idiots"}. This, combined with his privately held view that the world was irreversibly falling apart, he preferred to devote himself to what was immutable, studying the forces governing the world of natural phenomena.

Arda studied at Istanbul Technical University, made possible through a combination of scholarships and his salary as a part time army reserve officer cadet. He graduated with a degree in aeronautical engineering, 21 B.R, where he was then commissioned into the army as a lieutenant. He completed his research doctorarate in heliophysics two years later, completing his habilitation one year later, with rumours he solved a central problem that plagued his habilitation thesis while deployed on a field exercise.

Although wanting to focus on his research, even with his qualifications, employment was scarce and the army reserves was one of the few opportunities available that was stable and left him with enough time of his own. He was well respected in the army as a natural born leader, though his privately held view of the institution was one of resentment, corruption, and misuse.

He later underwent commando training with the 3rd Commando Brigade, Siirt, before being posted to the Mountain and Commando Brigade, Hakkari. He rose through the ranks before acting as the unit's liaison officer, overseeing North American Union troops undergo training at their facilities.


