% This is part of the Avaneya Project Crew Handbook.
% Copyright (C) 2010-2013 Cartesian Theatre <info@cartesiantheatre.com>.
% See the file Copying for details on copying conditions.

% Preface...
\StartChapter{Preface}

Avaneya is a cerebral science fiction game set on Mars currently under active development. The game is a {\it sui generis} combining the perfect blend of the city builder social simulation, real time strategy, and cooperative multiplayer. It explores mature themes in social justice, politics, ecology, economics, and more. It is geared towards those with a passion for hard science fiction and software {\it libre}, in particular users of the GNU operating system\footnotecite[gnu_users_who_have_never_heard_of_gnu] and its various incarnations, such as Ubuntu.\footnotecite[about_ubuntu]

This handbook is intended for Avaneya's ever--growing community of contributors who would like to get involved in this very exciting project. This is not a guide for people that simply want to play the game. For information on actual game play, readers will eventually be able to refer to the player manual.

This is a living document and will likely continue to be revised frequently and improved with time, especially while this project is in its early stages. Therefore, it is better to pass on a direct \href{\LatestHandbookURL}{link} to the most up to date revision than to a potentially outdated local copy.

Our purposes were several in number in authoring this book:

\startitemize[4]
\item
To encourage contributors to refine their ideas, think them through, and commit them to paper. This aids in deliberation and provides a place for public review;

\item
To define the project as clearly as possible within a single, consolidated, canonical location so contributors have something to reference;

\item
To politely assist current contributors from having to provide redundant explanations to newcomers with those already contained within the book;\footnote[rtfm]{It is always a good idea to ask questions, but as we say in the hacker community, don't forget to read the documentation first.}

\item
Lastly, to appease the odd doubtful naysayer that we did indeed undertake thoughtful due diligence, that the project was never idle tavern talk, and that we are prepared to see it through to completion.
\stopitemize

This book is intended for contributors in our community, and may not be as appropriate for players. Indeed, it may even spoil elements of the game. Consider yourself warned!

On the other hand, our eventual players of Avaneya are just as likely to enrich the game through constructive feedback -- feedback best buttressed with a well--informed understanding of the project's goals outlined in this comprehensive volume. 

\StopChapter
