% This is part of the Avaneya Project Crew Handbook.
% Copyright (C) 2010-2013 Cartesian Theatre <kip@thevertigo.com>.
% See the file Copying for details on copying conditions.

% Launchpad section...
\StartSection{Launchpad}

Launchpad is a project started by Canonical, the organization behind the Ubuntu project. It is a service that allows for online project collaboration. It does this by providing methods for revision control management, blueprints, support, issue tracking, and more. The Avaneya project's source code is hosted there.

If you want to join, make sure you remember to add yourself to the Launchpad Avaneya Project Crew.

\blank[2*big]
\startnarrower[3*left]
Project on Launchpad:\crlf
\fullahref{https://launchpad.net/avaneya}

\blank[2*big]
Avaneya Project Crew on Launchpad:\crlf
\fullahref{https://launchpad.net/~avaneya}
\stopnarrower

% Issue Tracking subsection...
\StartSubSection{Issue Tracking}
Launchpad's issue tracker, usually referred to as a bug tracker, is actually useful for tracking any kind of issue -- not just bugs. For this reason, we refer to it as an issue tracker. We use it for triaging resources, keeping track of what still needs to be fixed, and other things. 

Anyone with a free Launchpad account can report an issue and need not be a member of the project crew. This is important since Avaneya could crash for anyone.

\startnarrower[3*left]
\fullahref{https://bugs.launchpad.net/avaneya}
\stopnarrower

% Blueprints subsection...
\StartSubSection{Blueprints}
Launchpad provides a feature called Blueprints which allows you to track the design process from specifications all the way to implementation. Our project's currently available blueprints are as follows.

\startitemize[4]
\head {\em ares-architectural-design}

This document describes the AresEngine's internal architectural design. See {\tt Avaneya/Documentation/Contributors/AresEngine/UML/ReadMe} for details.

\startnarrower[3*left]
\fullahref{https://blueprints.launchpad.net/avaneya/+spec/ares-architectural-design}
\stopnarrower

\head {\em ares-package-design}

The AresPackage manifest file format and documentation is provided here. See \in{section}[AresPackages] for details.

\startnarrower[3*left]
\fullahref{https://blueprints.launchpad.net/avaneya/+spec/ares-package-design}
\stopnarrower

\head {\em ares-events-design}

The event type schema and definitions for built--in event types is provided here in the form of Events.xsd and Events.xml respectively. See \in{section}[Events] for details.

\head {\em avaneya-project-crew-handbook}

What you are reading now, of course.
\stopitemize

