% This is part of the Avaneya Project Crew Handbook.
% Copyright (C) 2010-2013 Cartesian Theatre <info@cartesiantheatre.com>.
% See the file Copying for details on copying conditions.

% Leitmotifs chapter...
\StartChapter{Leitmotifs}

This is probably the most important chapter in the entire book in helping readers to determine whether this is the project for them to be involved in or not. The game has several recurring ideas, or {\it leitmotifs}\index{leitmotifs}, within it. While they may seem unrelated, they are useful cues in reconstructing a virtual world that is intended as a metaphor of sorts for some of our most defining contemporary problems.

We would like to keep this \quote{leitmotif set} as minimal as possible. Anything else that we might wish to express, it is hopefully already possible to derive where obvious from those already present. However, if you feel the set should be modified in some way, please bring it up on any of the communication media described in \in{section}[Communication]. 

These leitmotifs are sometimes difficult to establish as definitive truths in the real world, however, we do our best to provide what we can where we can. What follows are given in no particular order of import.

\startitemize[4]
\head {\em Foreign and domestic policy measures predicated on staged terrorism or other orchestrated events are not unusual.}

It is not unusual for significant foreign and domestic policy measures in the more influential nations of the world to be predicated on acts of staged terror or other orchestrated events by elements within governments or their intelligence services for whatever reason. Although the goals vary, the method of providing false pretexts to socially engineer public opinion in favour of granting concessions they may not have otherwise has been a standard operating procedure for as long as governments have claimed a social contract of one form or another dependent on public opinion, or one that at least maintains the perception of one.

There is an overwhelming body of historical evidence indicating the synthetic nature of such events. Prominent examples include at least seminal moments in the First World War,
\footnotecite[extras={, pp.~264--269.}][the_creature_from_jekyll_island] % First World War...
the Second World War,
\footnotecite[extras={ p.~193.}][shirer1960]\footnote{{\it Ibid.} pp.~518--520.} % Second World War's Reichstag fire and the Gleiwitz incident...
the 1953 Anglo--American backed regime change in Iran,
\footnotecite[gasiorowski2000] % Operation AJAX to overthrow Mossadegh in Iran...
the Lavon Affair,
\footnotecite[curtiss1992_lavon_affair]
Cold War,
\footnotecite[marshall2008_gladio] % Cold War...
\footnotecite[northwoods]
the attack on the USS Liberty,
\footnotecite[underbakke2010_uss_liberty]
the Vietnam War,
\footnotecite[bumiller2010] % Vietnam...
regime change in Angola,
\footnotecite[goodman2009_staged_angola_intel]
the Troubles in Ireland,
\footnotecite[turbeville2010_staged_terror] % IRA...
the Iran--Iraq 1980--1988 war,
\footnotecite[presstv_cia_iraq_gas_attack] % Iran / Iraq war, 1980-1988...
the Oklahoma City bombing,
\footnotecite[griffin_andrew2010_murrah_bombing_survivor] %  Oklahoma City bombing in 1995...
\footnotecite[berger2009_mcveigh]
\footnotecite[watson2006_mcveigh]
the alleged terrorist attacks of September 11\high{th} 2001, % 9/11...
\footnotecite[gourley2012] 
\footnotecite[debunking_911_debunking]
\footnotecite[harrit2009]
\footnotecite[letter_to_minister_regarding_911]
\footnotecite[macqueen_911_peace_movement]
\footnotecite[watson2011_tim_osman_dead]
Venezuela's CIA--backed coup d'etat in 2002, % CIA in Venezuela in 2002
\footnotecite[golinger2010_venezuela_cia]
the subsequent occupations of Iraq % Iraq...
\footnotecite[doward2009_fake_un_drone_over_iraq]  
\footnotecite[beaumont2007_greenspan_iraq_oil]
\footnotecite[presstv2013_blair_lied_iraq_war]
\footnotecite[chossudovsky2005_iraq_sas]
and Afghanistan, % Afghanistan...
\footnotecite[reuters2011_scientific_evidence_undermines_afghan_war] 
\footnotecite[extras={ p.~74.}][grad2009]
\footnotecite[washingtons_blog_wesley_clark_regime_change]
various post--9/11 drama, % Post 9/11 related...
\footnotecite[king2010] 
including the 2005 London bombings, % London bombings in 2005...
\footnotecite[secker2010] 
the 2008 Mumbai attack, % Mumbai attack in 2008...
\footnotecite[chossudovsky2005_mumbai] 
countless other FBI hatched plans, % Compilation of FBI hatched plans...
\footnotecite[shipler2012] 
\footnotecite[schwartz2013]
calls for Western intervention in Iran, % Cheney Iranian Navy Seals in 2007...
\footnotecite[shakir2008] 
\footnotecite[ingersol2012_iran_false_flag]
the overthrow of Colonel Muammar Gaddafi in 2011, % NATO 2011 campaign in Libya
\footnotecite[rt2013_gold_dinar_gaddafi]
\footnotecite[brown2011_libya_central_bank]
the 2012 UK Olympic Games alleged terrorist threats, % 2012 UK Olympic Games staged terrorism...
\footnotecite[nimmo2012_staged_olympic_terrorism]
calls for US intervention in Syria in 2012--2013, % Syrian intervention...
\footnotecite[quinn2013_syrian_chemical_weapons_false_flag] 
\footnotecite[rt2013_saudi_black_op_damascus_gas]
\footnotecite[istreams2013]
\footnotecite[watson2012_syria_massacre_hoax]
\footnotecite[presstv_cia_syria]
\footnotecite[gucciardi2013_syrian_chemical_weapons_false_flag]
\footnotecite[watson2013_saudi_russian_olympic_threat] % Saudis threaten Russia with terrorist attacks for defending Syria...
\footnotecite[bbc2013_del_ponte_syrian_sarin] % US allies / proxies behind gas attacks in Syria
\footnotecite[gavlak2013_saudi_bandar_gas]
\footnotecite[corsi2013_syrian_gas] 
\footnotecite[gucciardi2013_syrian_gas]
\footnotecite[roberts2013_syrian_gas]
\footnotecite[rt2013_turkey_finds_sarin]
\footnotecite[watson2013_syrian_rebel_admits_chemical_weapons]
\footnotecite[milmo2013_syrian_fluoride]
and probably the 2013 Boston bombing. % Boston Bombing in 2013...
\footnotecite[interfax2013] 

Nevertheless, despite the synthetic nature and the overwhelming body of evidence in support of this, even in the age of information, this phenomenon is still met with hostility by conventionalists,\footnotecite[wood2013_what_about_building_7] in addition to remaining either unknown,\footnotecite[rethink911_story_nyt_missed] or an unregistered interest treated as anecdotal at best, by the mainstream media.\footnotecite[cspan_wtc]\footnotecite[cspan_wtc2]\footnotecite[cspan_wtc3]\footnotecite[cspan_wtc4]\footnotecite[letter_to_minister_regarding_911_enza_uda] 

If Senecca were alive today, he might have remarked that terrorism is real by the common man, contrived by the wise, and useful by those who rule. Nevertheless, we must remember that, while the reasons soldiers are given may be contrived,\footnotecite[extras={, pp.~\href{http://goo.gl/6hDCPk}{194}.}][bernstein2005_the_final_days] the hardships they endure are not.
\footnotecite[cbc2012_malaria_drug_dnd]

\head {\em Neoclassical economics is junk science.}

The major axioms of neoclassical economics, the backbone of contemporary economic theory, are wrong. Markets are not natural systems akin to the laws of thermodynamics.
\footnotecite[adbusters_issue_85]
\footnotecite[lasn2012]
\footnotecite[adbusters_jamming_neoclassical_economics]

Commercial advertising assists consumers in becoming less informed and making irrational choices. This is the opposite of how markets are allegedly expected to behave. People do not always exercise rational preference, seek self maximization, always act on full and relevant information, or stand the most to gain through competition over cooperation.
\footnotecite[equality_trust]

The word economy is rooted in the Greek {\it οἰκονόμος}. It originally meant {\it good household management}. Yet some kinds of economic growth can be destructive.
\footnotecite[ilo1998_child_prostitution]
\footnotecite[h2oil]
The market mechanism of the profit motive does not always seek to maximize the public good.
\footnotecite[ge_light_bulb_planned_obsolescence]
\footnotecite[the_light_bulb_conspiracy]
\footnotecite[mohan2010_manulife_clown_food]
\footnotecite[boyd2009_insurance_tobacco]
Prices do not always reflect true costs.
\footnotecite[roberts2013]
\footnotecite[cobb1999]
\footnotecite[oxfam2004]
\footnotecite[raj2010]
\footnotecite[coca_cola_case]
\footnotecite[extras={ See sections 4.6 and 4.11.}][agenda_21]
A nation's economic prosperity, as conventionally measured with the GDP, does not reliably correlate with its level of happiness.
\footnotecite[anielski2007_economics_of_happiness] Therefore, if an economy were a utility, it is not clear then what exactly it is expected to do.

Every single critical life support system on our planet currently is in a steady state of decline. This is reflected in virtually any of the scientific literature over the past several decades, in contrast to the aforementioned field still unwilling to grow up. It is a pseudo--science inebriated with physics envy masquerading as a science.

\head {\em Drug policies do not reflect reality.}

The supply and distribution of a significant proportion of the more dangerous illicit narcotics found on the streets originate through the covert assistance of the more influential governments and financial institutions of our world.
\footnotecite[ruppert_confronts_deutch] % Domestic
\footnotecite[banks_drug_cartels]
\footnotecite[webb1996]
\footnotecite[last_white_hope]
\footnotecite[cia_drug_plane_crash] % Mexico
\footnotecite[newman2011_mexican_drug_lord_cia]
\footnotecite[michaelson2011_vicente_discovery_motion]
\footnotecite[conroy2013_camarena_fell_cia_operation]
\footnotecite[afghan_opium_kingpin] % Afghanistan
\footnotecite[cia_drugs_list]

Public drug policies in the more influential nations are frequently criticized as ineffective. Conventional thinking asserts that either reforms are essential because such existing policies undermine the very safety of those they were allegedly intended to protect,\footnotecite[geddes2013_politics_of_pot] the general public, or more severe measures are necessary to continue to protect the general public.\footnotecite[extras={, p.~46.}][canadian_conservatives_platform2011] Both persuasions require an assumption that the legislation that made these drugs illicit were drafted in good faith -- however naïve, misinformed, or ideologically driven its legislators may have been.

Unfortunately this assumption that public safety was ever a genuine consideration to begin with must be discharged if there is absolutely any compelling evidence that any amount of any type of illicit narcotic's production or distribution is coordinated within the highest levels of government. The point is only that it is more reasonable to say that the United States, as an example, was never serious to begin with than it is to explain away its failure in its self described \quote{War on Drugs} merely as incompetence. 

Drugs have been used for centuries by those in positions of influence to do everything from regulating our behaviour\footnotecite[wilson2007]\footnotecite[cook2000_canadian_war_rum]\footnotecite[samuel2010_french_bread_cia_lsd] to undermining the health and sovereignty of foreign nations.\footnotecite[wikipedia_opium_wars] It should then come of little surprise that there is virtually no rhyme or reason in the social perception of drugs and how they correlate with the level of harm implicated by each.

Heroin was once a household children's cough medicine\footnotecite[edwards2011_bayer_heroin_ads_for_children] and opium was in the snuffbox of every gentleman of distinction. Today we reserve methamphetamine for the junky; Oreos for the office;\footnotecite[schroeder2013_oreos_like_drugs] and socially validated\footnotecite[ubc_wine_research_centre_home] alcohol overall more harmful\footnotecite[extras={ See figure 2.}][nutt2010] than virtually anything found on the street for every occasion. These arbitrary social values are whatever we need them to be, rationalized only after the fact that a precedent had already been set.

\head {\em No one is free.}

To be free is to be independent of influence, which no one truly is. We are all subject to influence; otherwise, there would never have been any prospects for commercial advertising, information would never have become a corporate commodity,\footnotecite[santoso2008]\footnotecite[nagata2011] and an honest media would never have become such a rarity, unavailable when it matters the most,
\footnotecite[watson2013_hastings_crash]
\footnotecite[rethink9112013_open_letter_to_nyt]
\footnotecite[rethink9112013_ads_debate]
\footnotecite[letter_to_minister_regarding_911_enza_uda] 
available when it doesn't.\footnotecite[cbc2013_tim_hortons_coin_thrower]

\head {\em Legal systems regulate the power of those without any, or those that threaten it.}

The powerful in the more influential nations of the world are rarely held to the same standard of justice as the underclass.
\footnotecite[kissinger1975_illegal_immediately]
\footnotecite[piff2012_social_class_unethical_behaviour]
\footnotecite[fielding2013_british_mp_paedophiles]
\footnotecite[watson2012_prostitution]
\footnotecite[long2013_cheney_in_toronto] 
Government is among the leading causes of death\footnotecite[statistics_of_democide] and reveals itself with occasion throughout history as there to protect the interests of a handful of principle benefactors,
\footnotecite[wilson2007]
\footnotecite[lukacs2013_csec_and_ebenezers] 
\footnotecite[bowie2012]
\footnotecite[favel2008]
\footnotecite[hurricane_katrina_alleged_executions]
\footnotecite[bender2011_pentagon]
not to make changes in the public interest.
\footnotecite[report_from_iron_mountain]

Some influential international government institutions may espouse a commitment to universal human rights, equality before the law, and impartiality, but they are not really serious.
\footnotecite[extras={ {\it vid.} 26 minutes, 35 seconds onwards.}][un2013_secretary_general_golfing_partners]

Governments tolerate democracy only so far as it remains merely cosmetic and is not permitted to actually function in any meaningful way.\footnotecite[extras={ Select the {\it Votes} column to sort demand by vote.}][occupy_poll] Real democracies have no need for off switches here\footnotecite[wikipedia_order_in_council] and there.\footnotecite[notwithstanding_clause] 

This should not discourage people into believing that everyone in government ought to be treated with suspicion, or that government can never be a constructive influence in everyday civic life. On the contrary, many in positions of (minimal) authority strive to improve the quality of life for everyone.

\head {\em Man digs his own grave with his knife and fork; conventional treatment of illness is institutionalized fraud.}

Disease begins on the plate.\footnotecite[extras={ According to Campbell, all of the top three causes of death in Canada in 2007 identified by StatsCan are entirely dietary.}][statscan_leading_causes_of_death] This is the reason for the disease state that we live in since nutrition and exercise are not as lucrative as merely treating symptoms. What we eat can sometimes kill us,
\footnotecite[seralini2011_gmo_safety]
\footnotecite[the_china_study]
make us stupid,
\footnotecite[cbc2013_toxic_sugar]
\footnotecite[case_against_fluoride]
or heal us.
\footnotecite[cavusoglu2009]

Conventional medicine is allopathic medicine. Allopathic medicine is pharmaceutical medicine. Pharmaceutical medicine culminated in 106,000 fatalities in 1994 due to adverse drug reactions (ADR), just in the United States, not counting overdoses, and under the direction of credentialed allopaths. This places the paradigm between the United States' fourth and sixth leading causes of death in that year.\footnotecite[lazarou1998_adr] By estimation then, approximately 2,700,000 fatalities can be attributed to the use of allopathic medications between the years of 1983--2010.\footnotecite[prousky2012_toxicology]\footnotecite[lindsey2012_vitamins_and_diabetes]\footnotecite[leape2000] 

Treatment aside, even just the allopathic approach to preventative medicine has been at least as damning.\footnotecite[miller2011_vaccination_infant_mortality] Like the allopathic preventative approach, the health insurance industry has nothing useful to bring to the table.\footnotecite[mohan2010_manulife_clown_food]\footnotecite[boyd2009_insurance_tobacco]

By contrast, in considering the dangers of natural treatments, but 12 fatalities were allegedly attributable to vitamin supplements during that same time span.\footnotecite[aapcc_annual_reports] Note that these are allegations only and remain unproven. 

Natural treatments are frequently referred to as \quote{alternative medicine}. Given that these natural approaches have been used by humans for at least 20,000 years,\footnotecite[hardy2012] with the allopathic approaches being relatively late comers in recent centuries, merit of either paradigm aside, it is more rational to reserve the term \quote{alternative} to that which proposes a deviation from an established norm.

\head {\em Agribusiness is destroying the planet.}

Agriculture is our principle means of interacting with the planet. Changing how we produce, distribute, and select our food solves many global problems.\footnotecite[foodwaste]\footnotecite[food_inc]\footnotecite[food_matters]

\head {\em Today's conquests are made through debt, not the sword.}

Usury, including fractional reserve banking, is a form of slavery.\footnotecite[the_creature_from_jekyll_island]\footnotecite[minutes_bank_of_canada] We might consider trading in debt to be the analogue of human trafficking, therefore. Debt is the tool of choice in an era of globalization.\footnotecite[perkins2005] This theme will be explored in \in{section}[Combat Globalization].

\head {\em Corporations are not people.}

The notion that corporations deserve the same rights as living human beings has proven itself absolutely catastrophic.\footnotecite[the_corporation]\footnotecite[coca_cola_case] Human civilization, since it first began building ships and mapping out the stars, had some how managed to engage in commerce for thousands of years without ever having required legal fictions. This theme will be explored in sections \in{}[Corporate Personhood] and \in{}[Money As Free Speech].

\head {\em Genes are a cop out.}

Very little of what we call human nature is actually genetic. Genes specify human needs, not behaviour. There is no such thing as {\it default} human behaviour. Most of what we call human nature is probably cultural and environmental influences that are not universal. Everything from our theories of morality to the way we reason about space and time are subject to environmental influences.\footnotecite[henrich2010] This will be elaborated upon in \in{section}[Socio-economic Simulation].
\stopitemize

There you have it -- leitmotifs for what we hope will be a great game to come. If you have made your way thus far and without suffering from a heart attack, you will probably find your involvement rewarding. Otherwise it may not be the best place for you. We are neither interested in authoring a casual game, nor free of the well justified accusation of shit disturbers, if it wasn't apparent already.

There are no requirements that anyone believe or do anything here. Everyone comes in having made a personal choice and through their own research. But rest assured, there are many other {\it libre} projects that could use the talents of those readers already infuriated, or at least still with serious reservations.

But regardless of whatever prospective contributors decide, creativity is essential. We always want to try to encourage our players to think for themselves in the process of coming to, what we hope will be, one of the preceding leitmotifs resonating with them. To paraphrase Socrates, {\it we cannot teach anyone anything. We can only make them think}.

\StopChapter

