% This is part of the Avaneya Crew Handbook.
% Copyright (C) 2010, 2011
%   Kshatra Corp.
% See the file License for copying conditions.

% Resources for everyone section...
\StartSection{Resources For Everyone}
These resources are useful for every kind of contributor, from researchers and writers to modellers and engineers.

You might consider reading some of these resources in order to enrich your understanding of the game, whose fabric is drawn from {\it at least} all of them. The more background knowledge we have, the greater the quality of the game. The same goes for really any kind of complex simulation, and not just Avaneya.

\startitemize[3]
\setupwhitespace[big]
\item
{\it A People's History of the United States: 1492-Present} (book)\crlf
Howard Zinn, HarperCollins, ISBN 9780060528423, 2003.

\item
{\it Adbusters} (periodical)\crlf
Adbusters Media Foundation, ISSN 08479097.

\item
{\it Blue Mars: Mars Trilogy Bk. 3} (book)\crlf
Kim Stanley Robinson, Voyager, ISBN 9780586213919, 1997.

\item
{\it Death by Government} (book)\crlf
R. J. Rummel, Transaction Publishers, ISBN 9781560009276, 1997.

\item
{\it Debunking 9/11 Debunking} (book)\crlf
David Ray Griffin, Olive Branch Pr, ISBN 9781566566865, 2007.

\item
\href{http://is.gd/QXTVmH}{Drug Harms in the UK: a Multicriteria Decision Analysis} (paper)\crlf
Nutt et al., The Lancet 376.9752 (2010): 1558-565.

\item
{\it Food, Inc.} (film)\crlf
Magnolia Home Entertainment, ASIN B0027BOL4G, 2008.

\item
{\it Free Software, Free Society: Selected Essays of Richard M. Stallman} (book)\crlf
Richard M. Stallman, Free Software Foundation, ISBN 9781882114986, 2002.

\item
{\it Green Mars: Mars Trilogy Bk. 2} (book)\crlf
Kim Stanley Robinson, Voyager, ISBN 9780586213902, 1994.

\item
{\it How to Live on Mars: A Trusty Guidebook to Surviving and Thriving on the Red Planet} (book)\crlf
Robert Zubrin, Three Rivers Press, ISBN 9780307407184, 2008.

\item
{\it Journal of Orthomolecular Medicine} (periodical)\crlf
International Schizophrenia Foundation.

\item
\href{http://www.thevertigo.com/html/9-11/}{Kip's Open Letter to Minister of Public Safety Regarding 9/11}\footnote{This letter was written by the project lead and summarizes much of his understanding of the world at the time it was composed.} (letter)\crlf
Kip Warner, 2011.

\item
{\it Merchants of Doubt: How a Handful of Scientists Obscured the Truth on Issues from Tobacco Smoke to Global Warming} (book)\crlf
Naomi Oreskes et al., Bloomsbury Press, ISBN 9781596916104, 2010.

% Anyone have a reliable online link including pp. 461-500?
\item
{\it Minutes of Proceedings and Evidence Respecting the Bank of Canada} (paper)\crlf
Standing Committee on Banking and Commerce, Printer to the King's Most Excellent Majesty, 1939, pp. 461-500.

\item
{\it Red Mars: Mars Trilogy Bk. 1} (book)\crlf
Kim Stanley Robinson, Collins, ISBN 9780586213896, 1993.

\item
\href{http://www.teachpeace.com/Report_from_Iron_Mountain.pdf}{Report From Iron Mountain: On The Possibility And Desirability Of Peace}\footnote{If in doubt of the document's authenticity, refer to p.524 of Griffin's {\it The Creature From Jekyll Island}.} (paper)\crlf
John Kenneth Galbraith et al., The Dial Press, 1967.

\item
{\it Statistics of Democide: Genocide and Mass Murder since 1900 (Macht Und Gesellschaft, Bd. 2)} (book)\crlf
R. J. Rummel, Lit Verlag, ISBN 9783825840105, 1999.

\item
{\it Terraforming: The Creating of Habitable Worlds (Astronomers' Universe)} (book)\crlf
Martin Beech, Springer, ISBN 9780387097954, 2009.

\item
{\it The Case Against Fluoride: How Hazardous Waste Ended Up in Our Drinking Water and the Bad Science and Powerful Politics That Keep It There} (book)\crlf
Paul Connett et al., Chelsea Green Publishing, ISBN 9781603582872, 2010.

\item
{\it The Case for Mars: The Plan to Settle the Red Planet and Why We Must} (book)\crlf
Robert Zubrin et al., Free Press, 1997.

\item
{\it The China Study: The Most Comprehensive Study of Nutrition Ever Conducted and the Startling Implications for Diet, Weight Loss and Long-term Health} (book)\crlf
T. Colin Campbell, Benbella Books, 2006.

\item
{\it The Corporation} (film)\crlf
Zeitgeist Films, ASIN B0007DBJM8, 2004.

\item
{\it The Creature from Jekyll Island: A Second Look at the Federal Reserve} (book)\crlf
G. Edward Griffin, American Media, ISBN 9780912986395, 2010.

\item
\href{http://www.equalitytrust.org.uk/resources/publications}{The Equality Trust} (publication syndicator)\crlf
32-36 Loman Street, London SE1 0EH.

\item
{\it Zeitgeist: The Movie} (film)\crlf
Peter Joseph, 122 minutes, GMP LLC, 2007.

\item
{\it Zeitgeist: Addendum} (film)\crlf
Peter Joseph, 123 minutes, GMP LLC, 2008.

\item
{\it Zeitgeist: Moving Forward} (film)\crlf
Peter Joseph, 161 minutes, GMP LLC, 2010.
\stopitemize

