% This is part of the Avaneya Project Crew Handbook.
% Copyright (C) 2010, 2011, 2012
%   Kshatra Corp.
% See the file License for copying conditions.

% Resources For Artists section...
\StartSection{Resources For Artists}
These resources are useful for different kinds of artists, such as 2D artists, audio engineers, cinematic artists, modellers, musicians, and so on.

\startitemize[4]
\item
{\it 3D Computer Graphics, Second Edition}\footnote{This book is listed as a resource for artists because it is a wonderful and concise introduction to the theory of computer graphics, but for non-programmers and non-mathematicians. It is a classic, though long out of print. Nevertheless, it is not hard to find used copies available for sale.} (book)\crlf
Andrew Glassner, Green Editorial, ISBN 9781558213050, 1994.

\item
{\it Beginning GIMP: From Novice to Professional} (book)\crlf
Akkana Peck, Apress, ISBN 9781430210702, 2009.

\item
{\it Blender 2.5 Materials and Textures Cookbook} (book)\crlf
Colin Litster, Packt Publishing, ISBN 9781849512886, 2011.

\item
{\it Blender 3D 2.49 Incredible Machines} (book)\crlf
Allan Brito, Packt Publishing, ISBN 9781847197467, 2009.

\item
{\it Blender 3D Architecture, Buildings, and Scenery: Create photorealistic 3D architectural visualizations of buildings, interiors, and environmental scenery} (book)\crlf
Allan Brito, Packt Publishing, ISBN 9781847193674, 2008.

\item
{\it Blender Studio Projects: Digital Movie-Making} (book)\crlf
Tony Mullen & Claudio Andaur, Sybex, ISBN 9780470543139, 2010.

\item
{\it Introducing Character Animation with Blender} (book)\crlf
Tony Mullen, Sybex, ISBN 9780470102602, 2007.

\item
{\it The Essential Blender: Guide to 3D Creation with the Open Source Suite Blender} (book)\crlf
Roland Hess, No Starch Press, ISBN 9781593271664, 2007.
\stopitemize

