% This is part of the Avaneya Project Crew Handbook.
% Copyright (C) 2010, 2011, 2012
%   Kshatra Corp.
% See the file License for copying conditions.

% Introduction to Launchpad section...
\StartSection{Introduction to Launchpad}

Launchpad is a project started by Canonical, the company behind the Ubuntu project. It is an online service that allows for coordinated software project collaboration. It provides services for source control management, blueprints, support, bug tracking, and more. The Avaneya project, the people behind it, and the source code as it develops can be found there.

If you are a contributor, make sure you remember to add yourself to the Launchpad Avaneya Project Crew.

\blank[2*big]
\startnarrower[3*left]
Project development website:\crlf
\fullahref{https://launchpad.net/avaneya}

\blank[2*big]
Launchpad's Avaneya Project Crew home:\crlf
\fullahref{https://launchpad.net/~avaneya}
\stopnarrower

% Blueprints subsection...
\StartSubSection{Bug Tracker}
Launchpad's bug tracker tracks all issues submitted to it originating from anyone. It is very useful for triaging resources and keeping track of any issues that are still outstanding.

\startnarrower[3*left]
\fullahref{https://bugs.launchpad.net/avaneya}
\stopnarrower

% Blueprints subsection...
\StartSubSection{Blueprints}
Launchpad provides a feature called Blueprints which allows you to track the design process from basic feature specifications all the way to implementation. Our project's currently available blueprints include:

\startitemize[4]
\head {\em avaneya-project-crew-handbook}

What you are reading now, of course.

\head {\em ares-architectural-design}

This document describes the AresEngine's internal architectural design. See the {\tt ReadMe} for details.

\startnarrower[3*left]
\fullahref{https://blueprints.launchpad.net/avaneya/+spec/ares-architectural-design}
\stopnarrower

\head {\em ares-package-design}

The AresPackage manifest file format and documentation is provided. See \in{section}[AresPackages] for details.

\startnarrower[3*left]
\fullahref{https://blueprints.launchpad.net/avaneya/+spec/ares-package-design}
\stopnarrower

\head {\em ares-events-design}

The event type schema and definitions for built-in event types, Events.xsd and Events.xml respectively. See \in{section}[Events] for details.
\stopitemize

