% Core Leitmotifs chapter...
\chapter{Core Leitmotifs}

This is probably the most important chapter in the entire book in helping you decide whether this is a project for you to be involved in or not. 

Avaneya has several recurring ideas or @dfn{leitmotifs} it communicates frequently to its users. The following form the core ideas the project feels are important. As you consider each idea, given in no particular order, you are also evaluating Avaneya itself in the process. 

If you feel something more should be added to the list, consider if it is not already implicated from one of the others.
@sp 1

@enumerate
\item
People are not free if they are not informed.@footnote{See @i{The Equality Trust} in @ref{Resources For Everyone}.}
@sp 1

\item
The basic premises neo-classical economics are built upon, such as an infinite growth model@footnote{See @i{Adbusters), with prejudice for @url{http://www.adbusters.org/cultureshop/backissues/85,Issue #85}.} in @ref{Resources For Everyone}} operating within a finite environment, an absence of true cost accounting,@footnote{See @i{Food, Inc.} in @ref{Resources For Everyone} as well as @ref{Genuine Progress Indicator}.} and so on, are fundamentally fraudulent and have acted as a vehicle for dangerously unchecked assumptions.@footnote{See @url{http://anticap.wordpress.com/2010/10/25/jamming-neoclassical-economics/}.}
@sp 1

\item
As government increases in size, it tends to become more dangerous, less useful, and the people less free.@footnote{See @i{Statistics of Democide: Genocide and Mass Murder since 1900 (Macht Und Gesellschaft, Bd. 2)} in @ref{Resources For Everyone}.}
@sp 1

\item
Most major acts of terrorism are staged by government.@footnote{See @i{Debunking 9/11 Debunking} in @ref{Resources For Everyone}.}
@sp 1

\item
Governments are generally not there to make changes to serve the public interest. They are there to keep things the same.@footnote{See @i{Report From Iron Mountain: On The Possibility And Desirability Of Peace} in @ref{Resources For Everyone}.}
@sp 1

\item
Most of the diseases that afflict mankind are caused by diet.@footnote{See @i{The China Study} in @ref{Resources For Everyone}.}
@sp 1

\item
The majority of the presence of the most dangerous drugs in society are attributable to government, either overtly, as in the case of the most dangerous recreational narcotic, alcohol,@footnote{Nutt, David J., Leslie A. King, and Lawrence D. Phillips. @url{http://www.thelancet.com/journals/lancet/article/PIIS0140-6736(10)61462-6,"Drug Harms in the UK: a Multicriteria Decision Analysis."} @i{The Lancet} 376.9752 (2010): 1558-565. Print.} or covertly, as in the case of substances like cocaine.@footnote{@url{http://afp.google.com/article/ALeqM5j6QonBKKMo2gw1e3ql-xUcQEZbVg,"Mexico Drug Plane Used for CIA ‘rendition’ Flights."} @i{Associated Free Press}. 4 Sep. 2008.}
@sp 1

\item
Agriculture is our principle means of interacting with the planet. Change what you eat and you can solve many problems.@footnote{See @i{Food, Inc.} in @ref{Resources For Everyone}.}
@sp 1

\item
Fractional reserve central banking is built upon fraud.@footnote{See @i{The Creature From Jekyll Island} in @ref{Resources For Everyone}.}
@sp 1

\item
Usury is the most ubiquitous and sophisticated form of modern slavery.@footnote{@i{Ibid.}}
@sp 1

\item
Granting corporations the rights of human beings was a profoundly bad idea.@footnote{See @i{The Corporation} in @ref{Resources For Everyone}.}
@sp 1

\item
There is no such thing as human nature, only human behaviour.@footnote{See @ref{Socioeconomic Modelling}.}
@sp 1

\item
Everything is connected and nothing happens in a vacuum.@footnote{Reasonable to take axiomatically.}
@sp 1
@end enumerate
@sp 1

If you have gotten thus far, and you are still comfortable being involved in this project, then you will probably find it rewarding. Otherwise, there is no sense in being here. No one is forcing you to do anything. This is not a project for everyone and there are countless other community driven projects that could probably use your talents.

But regardless of whatever @i{you} choose to do, creativity is required whenever @i{we} present the aforementioned. We should also always try to require the user to do some thinking of their own to arrive at these conclusions. As the Buddhists say, @i{you cannot teach a man anything. He can only teach himself}.

