% This is part of the Avaneya Project Crew Handbook.
% Copyright (C) 2010-2017 Cartesian Theatre™ <info@cartesiantheatre.com>.
% See the file Copying for details on copying conditions.

% Game Mechanics chapter...
\StartSection{Workforce Change}

%--------------> INFLUENCING WORKFORCE <------------------------------------- 
\StartSubSection{Influencing Workforce}
How people are raised, educated, and conditioned through experience largely determines their consumer patterns and may be reflected in a population's consumer choices. In game, workforce is influenced through players building, inacting ordinance, and other various city policies along with inter-city diplomacy. The method in which various workforce units are shaped and changed through the development of a players city are complex and discussed in depth below.

The main method in which a player can influence his own or someone elses workforce is through economic means such as taxation or ordinance (though not limited to financial insentives but extended to include incentives derrived for economic direction), although other methods may exist such as brute force or through environmental coercion. Additional methods include the use of the media and/or propaganda which can be initiated by the player, another player, or AI representing Earth interests.

The means in which the workforce itself is influenced is through their needs and interests (see section Workforce Units Needs). Workforce units will go about striving to fulfill there needs. This depends on the situation at hand and at the players influence.

In avaneya there will be multiple levels of cultural and economic variance influencing different cultural groups to different degrees causing players to deal with this to keep all aspects of the city functioning while still trying to achieve player goals such as technological innovation and maintain an independent yet peaceful existance with Earth.

One group of humans with similar workforce types might make business deals to get it out of the city or off the planet. Another might petition government to increase militia spending. Another might encourage the people to take to arms and fight. And another might join the militia. Outliers may take to vigilante justice or investigate to discover possible conspiracies. Of course the degree to which these can occur depend largely on previous player input to the population. There is a multitude of means in which entities will respond to player sitations that threaten or improve their circumstances and these are unique to the needs being effected and resource the unit has available to make a change or maintain safety.

When there are multiple threats to the same entity, it will respond in order of importance to which it must meet its needs. Clearly air, then water, then food, then security, then belonging. Entities will not respond to their need to belong when they are dieing from a lack of oxygen.

In its natural state a workforce unit is only trying to satisfy its needs. This may be going to a market, going out to lunch, or stealing from a grocery store. In its default state the unit pursues tasks, or actions, in accordance with its needs, using its interests as a guide.

To determine how a unit acts is a matter or factoring in its various needs at the time along with variables such as workforce, income level, social status, and location. In this way we determine whether an unit goes home to cook food for dinner, or goes out to dinner with friends, or goes home has dinner then goes bowling. Factors such as the units income or money, degree of hunger, location to food, food quality, need to belong, number of friends (or daily interactions), neighbourhood safety, "governmental" dietary guidelines, work security, and time available, all converge to determine a units action. 

Not everything will be required to have an action, breathing and drinking water are just factored into resource use and will not evoke a reaction unless the unit is unable to obtain or is blocked from this resource.

As each workforce unit is represented by an individual unit in game, moving and performing tasks, as this unit moves through game space it is consuming resources such as air, water, food, and minerals and like subtances through consumerable resources. What a unit can consume is based in part by what is available and what its needs and action influences are


%SPREADSHEET containing every possible action for each need, age group when possible, with qualitative and quantitative variables, factors that cause a unit to take a certain action, and factors that certain actions will remedy or inflame, duration of action, action chains, reactions that cancle actions, reactions that change actions, action groups and priorities (ie stealing to get food will not happen before buying food), normal vs starvation mode actions, actions by workforce type
% -> SPREADSHEET -> Example Entity Actions [an (E) signifies emergency/fight flight response]
Oxygen Actions
-breath 
-move to air source (E)
-stay calm (E)
-panic and scramble for air (E)
-murder to lessen burden on air supply (E)
-partition government for cheaper air
-protest for cheaper air
-steal air supply
-riot for cheaper air
-ask for air
-beg for air

Water Actions
-move to water source
-stay calm
-panic and scramble for water (E)
-murder to lessen burden on water supply (E)
-partition government for cheaper water
-protest for cheaper water
-riot for cheaper water
-steal water
-ask for water
-beg for water


Food Actions
-eat food
-cook food
-buy food
-ask for food
-beg for food
-steal food 
-go to resturant/cafe/food court/ etc
-go to resturant/cafe/food court/ etc with friend
-murder for food (E)
-kill animal
-partition government for cheaper food
-protest for cheaper food
-riot for cheaper food (E)


Security Actions
-partition government for: militia; police; emergency services
-partition government for: attack neighbouring city; ethnic cleansing
-protest government for: above
-riot for: above
-buy house
-rent house
-squat
-build house
-partition/protest/riot government for: cheaper housing
-purchase firearm/weapon
-partition/protest/riot government for: access to firearms/weapons
-fitness/martial arts
-vigilante group



Belonging Actions
-
-
-
-
-
-


There are two broad categories which influences a units needs. The first is the units own actions, drinking water, eating food, getting a job, visiting a friend, all reduce the need to obtain these things, while excerising increases the need for food and water, and working long hours in isolation increases the need for belonging. The second is external or environmental effects such as propaganda, lack of clean air, food with poor nutitional content, or the action and views of other units. All these happenings have varying amounts of influence on the units needs which in turn effect the manner in which the unit acts to satisfy these needs.

In order to correctly map the degree of influence various needs have on a unit and the degree to which a units actions relieves or increased its needs, all of the variables effecting the needs of a unit are mapped with a quantitative point system. Some of these influences will be wide reaching, effecting many units, such as chemicals in the cities water supply or television, other will be specific, such as health consequences from working shift work. 


%-> SPREADSHEET -> Influences
-Workforce Types
-Food Types
-Actions
-Information Types
-Economic 
-Drugs
-Social
-Ostracism





%--------------> WORKFORCE CONSUMPTION <---------------------------------------
\StartSubSection{Workforce Consumption}

Units aim to fulfill needs, the method in which this is done is via their interests.

Units undertaking actions, buying goods and services (among other things), preferenced by their variables and preferenced further by the hierarchy of needs.?

see section ... in microeconomics.


\StartSubSection{Workforce Crafting}
Most items are created by combining several other items to create them. Generally this is done by businesses such as factories however sometimes units will conduct crafting as a part of their consumption. Units will consume a number of items to create a final product where the utility is greater. 

An example is cooking, where a unit will need to purchase several items to create a final product in a kitchen.
%Spreadsheet of all crafting items.

The method in which units choose recipes and what items go into recipes is described in the micro_economics: Budget Constraints and Preferences section. This section will discuss how the crafting actually takes place once goods are purchased.

%Method of crafting
% this needs to account for both units buying goods and crafting them straigh away, and crafting from goods that are available or already purchased.
% therefore units will accumulate goods to be crafted over a period of time?

%-items chosen as per Budget Constraints and Preferences section

%-items colated at point
  -how
Units  
  -where
  -when

%-point contains tools
  -what
  -where

%-time
  -when
  -how long

%-item crafted
  -now what 

