% This is part of the Avaneya Project Crew Handbook.
% Copyright (C) 2010, 2011, 2012 Cartesian Theatre <kip@thevertigo.com>.
% See the file Copying for details on copying conditions.

% Game Mechanics chapter...
\StartSection{Starting Locations}
The initial place where players are shown after gameplay loading is complete has a value in story telling and in revealing areas of the map to a player. The start location is also the area from which a player first implements her influence and this being so the start location and the areas surrounding them need to be considered carefully, especially on Solnet.
%Sectors [section 9.1]
%Start locations set on Solnet, made approx equal value. good distance apart for growth.

A starting location represents two things, this is the first area the player will be shown on loading the scenario and it is generally the beginning of the players construction and development.  

In the campaign or scenarios the start location will generally be determined according to the scenario with many buildings already in place, unless choosing a point of beginning construction is a part of the scenario / mission. Start locations on campaign or scenario maps can easily be preset. The aim is to focus a players attention for strategic or story driven reasons. Rarely will the start location impact gameplay beyond these reasons in single player play and rarely will the start locations be non--obvious to creators.

On Solnet, start locations will be preset by map creators and will be balanced taking into account all factors such as distance to other players, varieties of resources and their volume, and terrain (intrusiveness, hostility, hospitality, strategic value). Also to be considered are the options certain starting resources enable and how desirable resources are in certain locations and how far these may be depleted; whether new resources are spawned at player spawn and how much of these; resource renewability; technological progress of surrounding players / cities; protection offered by players; ability to mobilize starting wealth to affect gameplay; and overall how much the game dictates what starting players can do to others and what others can do to them. For instance, if starting has a large sum of money can players use this to screw over nearby players then quit.

A fine line needs to be tread between limiting player choice in order to help players and letting players and diplomacy create events and order out of uncertainty. Essentially maintaining playability and while keeping the integrity of sandbox style gaming.

