% This is part of the Avaneya Project Crew Handbook.
% Copyright (C) 2010-2013 Cartesian Theatre <kip@thevertigo.com>.
% See the file Copying for details on copying conditions.

% Game Mechanics chapter...
%\StartSection{Workforce}

\StartSubSection{Workforce}
%INTRODUCTION INTRODUCTION
What a city can produce, how much it can produce, and what it is capable of producing are largely byproducts of its workforce however in Avaneya workforce also determines things such as consumption patterns and criminal behaviour. This section discusses the variety of workforces, what they represent, how they are created, maintained, and influenced, and what they are capable of. 

A workforce unit is essentially a singular individual of the population, an animal, or a piece of technology that either exists at a certain level of education, training, or lack there of, or is built or bred to a variety of types, scales, and qualities (see section Workforce Types). Workforce includes all labour factors driving the economy and extends to consumers (and even the homeless, criminals, and the unemployed). The workforce of a city then is all units in the city capable of exerting influence on the direction in which resources or money flows. A labour workfoce requires certain factors to bring it into fruition which may include a certain level of education, access to work, and restrictions to certain aspects of these. An intellectual workforce requires different factors to develop then a manual labour workforce. While the two are not mutually exclusive it may or may not be more difficult to create multiple classes in a single city. In adition otherwise different workforces may share types aspects of influence over the economy, known as workforce types. For instance a workforce unit in manufacturing may have the same education level and by the same food as a workforce unit in commerce. These two units share a couple of "workforce types" (qualative education and consumer choices) but have been influenced by other means to different outcomes. A cities workforce is influenced through players building, inacting ordinance, and other various city policies along with diplomacy. The method in which various workforce units are shaped and changed through the development of a players city are complex and discussed in depth through the various subsections of the section.

\StartSubSection{Visual Workforce}
A cities multi faceted workforce is actually made up of varying amounts (many thousands) of individuals, humans, animals, and machines, and are depicted visually by people or objects performing tasks or moving around the city. Each workforce unit is capable of influencing the direction the greater workforce and the city take and each unit is an individual human, animal, or some form of machine involved in labour. Visually, workforce units need to be represented as: humans of varying races, ages, sizes, and wearing various clothing; animals of varying species and breeds; and machines of varying types and degrees of repair. Depending on the zoomable scale all these units must be able to be seen performing actions and influencing there surroundings and interacting with other units as well as their environment. A large variety of animations will be required for the number of potential interactions. 

\StartSubSection{Workforce Movement}
Workforce is a dynamic "resource" and can move city wide (if allowed) and globally (if they are not trapped in the city). If your workforce requires cheap housing and this becomes easier to gain in another city, many may migrate to this city seeking better opportunities. Likewise, they can move within your city to different opportunities and movement both within a city and between cities can be controlled to an extent to enable better or worse access to certain locations, increasing and decreasing travel time, or diverting travel for other means. This is just as true for mechinical workforce which may be moveable on its own, or through human interaction. 

Indeed a lot of a workforce units interaction and influence over the cities economy and general development will have to do with its ability to move. How far a unit can travel while meeting all its needs (discussed below) will determine this influence. Depending on the unit type it may be walking to work littering, going out to lunch, moving between two locations at work (goods transport), walking in a biodome paddock eating grass, or travelling along an assembly line. Also, the unit may be in cohesion with other units performing some of these tasks, for example eating lunch with a group of friends, building a tank, baking bread, or theater or dancing. 

The means in which this movement is and can be performed is based on the units needs and the vicinity to be able to meet the needs. Movement can also be restricted through placing private or of limit areas, constructing walk ways in cetain areas and directions, making some direction longer and others shorter, charging to travel some directions, and creating and limiting private and public transport options. 


\StartSubSection{Workforce Bases}
%Workforce Bases
There are three different fundamental types of workforce. There is a mechanical workforce made up of machines, computers, and other similar automatons. There is a human workforce made up of the biological and social entities which are humans. And there is a animal workforce which is also biological but for the games sake is not influenced by the social fabric of a city. These are known as workforce bases. 

The mechanical workforce requires research and development into avenues such as robotics and mechanical engineering with a social system in place requiring either the production or the import of mechanical systems. The human workforce is made up of human labour (physical, mental, and both) performing not only the tasks required by the society, but also tasks or actions that may detriment the society. As the cities population is influenced by numerous external and internal agencies such as media, education, drugs, and fashionable consumption it is convoluted and multifaceted, and the style of play implemented by the player, other players, and NPCs, will help to determine the quality and quanity of a cities human workforce. The animal workfroce is similar to the human workforce in its being a biological entity, and similar to a mechanical workforce in its one dimensional aspect of usefulness. Cities in Avaneya will generally have a mix of these three base workforces and each of these in term will have many varying types (see subsection on Workforce Types). 

In addition to a city producing its own workforce, it may purchase machinery and animals, attract population, and hire individuals directly from other cities. While the city may purchase certain workforces at some stage or another, purchasing of individuals and machinery generally takes place by businesses. The main method in which cities attract additional population (apart from reproduction) is through the cities governing player creating opportunities which can keep population in the city or attracting population from other cities. 


\StartSubSection{Workforce Units Needs}
As mentioned briefly above a cities workforce is made up of many individual units. Each of these units can be a part of one or more workforce types and require certain conditions or needs met in order to be created, maintained, and have degrees of functionality. The human workforce has needs which differentiates them from the other two workforce bases (animals and machines). While machines require maintainence and energy, and animals require food, water, and shelter, in Avaneya humans interact with the players decisions, political or otherwise. Animals and machines are only effected by ordinance and taxation through humans. 

The needs an entity, or individual human, has determine where it will go, what it will do, and how it will do it. If we take sections [] accuracy into account we can view humans in a largely tabula rasa state, with no ill will except to maintain there own needs. Further to this, it can be seen that the nature in which humans interact and the means in which they satisfy their needs is determined by something outside of there biological state. 

The human needs consist of:
Oxygen: breathable air, a human can survive for less than 3 minutes without suffering potentially serious consequences. 

Water: clean drinking water. The human body can survive upto a week without water (provided good conditions). 

Food: sustainence and nutrtion. It is essential to not only have something to eat but that it contains the various essential nutrients to support human life. A human can last several weeks without food provided they have water.

Security: shelter/housing, feeling safe in surroundings, and feeling secure about the future.

Belonging: feeling of making a contribution, having people to talk to, being a part of something more than hirself.

In light of the above paragraphs, it can be seen that it is quite simple to satisfy a humans security needs. However, continually bombard hir with messages of insecurity and threats despite evidence to the contrary and it will seek ever greater means to satisfy this need, perhaps waging unnecessary wars against otherwise peaceful peoples. This is just an example of how a player can create or solve disonance in human needs and scenarios the game can implement to challenge players.



\StartSubsection{workforce Types}
The dynamic and multifaceted nature of workforce may become more evident when reading about the various workforce types. From the list below it can be seen that workforce comprises what the population is doing in reaction to the cities governance or that of other cities and also what stage of development a unit is at. Some types of workforce are easy to maintain in a state desired by players, such as industrial machinery. This needs only to be created, maintained, and given a source of energy. Other workforce types are more specific such as "the high school teacher" or "motor mechanic". These require specific circumstances surrounding a population, with just the right levels of affluence and education. Whats more is human workforce types also share needs with other human workforce types and are dependant on eachother to provide those needs. This makes for a complex array of interaction between workforces, and when combined with taxation levels, ordinance, and geopolitical factors the complexitity is multiplied. In addition workforce types applied to a workforce unit can overlap. A unit could be tertiary educated but unemployed, or a physical labourer with a secondary education. 

Also individuals may move between workforce types, as in moving through education levels or going from unemployed to employed. Some factors which change workforce types are done through time, others done through city management. The game must track each workforce unit, its movements, its needs, its progress (or degression) through education and work, its workforce types, its use of drugs, or its breaking the law. It must track this and the interaction with other workforce units, and the impact this all has on other units, (either directly through actions or indirectly through policy change or similar) and on the environment and the cities statistics, spending, and policies.


%Workforce Types
-Homeless
-Unemployed
-Physical labour (varying levels/classes, ie production, mining)
	(a physical labour type for each task needed in city) 
-Mental labour (variious)
-Education (primary, secondary, tertiary, post tertiary [not complete, see education types])
-Employment skill (varying types depending on field, ie years of experience)
-Training (varying types, ie separate training that a business or government may provide)
-Criminal (varying types)
-Militia
-Consumer (varying types depending on influences, ie social eco status, cultural)
-Mechanical (varying) 

-Protester
-Animal (varying)
-Age (varying)
-Drug user (various)
-Drug addicted (various)





\StartSubSection{Transitioning Between Workforce Types}
%Unit gets skill and either upgrades workforce type or adds a workforce type.
%How do units get skill - interests, needs, continual work, education, age, handicaps.
%How does a unit move between jobs?
%How does a unit choose an education type?





\StartSubSection{Influencing Human Workforce}
%
How people are raised, educated, and conditioned through experience largely determines their consumer patterns and may be reflected in a population's consumer choices. In game, workforce is influenced through players building, inacting ordinance, and other various city policies along with diplomacy. The method in which various workforce units are shaped and changed through the development of a players city are complex and discussed in depth below.
%

There are various methods in which the human workforces are influenced to take a certain direction. The main method in which a player can influence his own or someone elses workforce is through economic means such as taxation or ordinance, although other methods may exist such as brute force or through environmental coercion. Additional methods include the use of the media and/or propaganda which can be initiated by the player, another player, or AI representing Earth interests.

The means in which the workforce itself is influenced is through their various needs (see section Workforce Units Needs). Human workforce units will go about striving to fulfill there needs by a means which is partially predetermined by default programming and partially determine by the needs needing be met and the degree to which they are wanting. Usually, all things being good, this is peaceful but ultimately depends on the situation at hand and at the players influence. Through enacting ordinance, taxation, and various diplomatic relations the player creates variance in the workforce needs levels, and over time the method in which each workforce unit meets its needs.

An example may be the implementation of a higher tax level for a certain group of people. This may place strain on there income, causing crime, violence, and other problems for the city. However if this higher tax level is offset with social funding for this group, giving them other means to obtain there needs then the previous situation with crime level increase may not occur. 

Another example is allowing television and private broadcast. While cheaper for the government and good for business, this system may become abused with continual broadcasts leaving people feeling that their belongingness need is to be fulfilled by wearing the same clothes as some Earth celebrity. 

The above two examples are simple occurances. In avaneya there will be multiple levels of cultural and economic variance influencing different cultural groups to different degrees causing players to deal with this to keep all aspects of the city functioning while still trying to achieve player goals such as technological innovation and maintain an independent yet peaceful existance with Earth. From this it can be seen the the AI of human workforce will be quite involved.   

%Discuss how effected and how entities will react.
The AI of entities will respond to a variety of inputs from the greater enviroment. The manner in which each entity attempts to reach satisfaction of its needs is determined from this environmental input. To describe the process, an individual entity will go about in its default state provided it has access to all its essentials. When this access is threatened or blocked it will attempt to right the situation by some means. The means in which it can right the situation depends on the enviroment, namely economic, political, and location factors, as well as its education/knowledge and social status, and the need that is being threatened. Further to this, an entity may not even be aware of some changes or threats and therefore might not take action

Lets go back to our example on a threat to security. One group of humans with similar workforce types might make business deals to get it out of the city or off the planet. Another might petition government to increase militia spending. Another might encourage the people to take to arms and fight. And another might join the militia. Outliers may take to vigilante justice or investigate to discover possible conspiracies. Of course the degree to which these can occur depend largely on previous player input to the population. There is a multitude of means in which entities will respond to player sitations that threaten or improve their circumstances and these are unique to the needs being effected and resource the unit has available to make a change or maintain safety.

%Multiple threats to multiple needs
When there are multiple threats to the same entity, it will respond in order of importance. Clearly, air, then water, then food, then security, then belonging. Entities will not respond to their need to belond when they are dieing from lack of oxygen.

%Natural State
In its natural state a workforce unit is only trying to satisfy its needs. The manner in which the unit can do this is chosen by the method in which the city is managed. An unit seeking food could naturally go into the wilderness and find fruit etc., however in Avaneya, on Mars, this is not possible and the unit therefore finds other means of getting food. This may be going to a market, going out to lunch, or stealing from a grocery store. In its default state the unit pursues tasks, or actions, in accordance with its needs.

To determine the type of action an unit will undertake is complex. It is a matter or factoring in its various needs at the time along with its workforce type and other variables such as income level, social status, and location. In this way we determine whether an unit goes home to cook food for dinner, or goes out to dinner with friends, or goes home has dinner then goes bowling. Factors such as the units income or money, degree of hunger, location to food, food quality, need to belong, number of friends (or daily interactions), neighbourhood safety, "governmental" dietary guidelines, work security, and time available, all converge to create the units action. These are automatically calculated by the program and are known as actions or specifically as unit actions. The actions taken by units are classed into categories based on needs. 

Not everything will be required to have an action, breathing and drinking water are just factored into resource use and will not evoke a reaction unless the unit is unable to obtain this resource. The main use for entity ations is to document visually and statistically the units input into the economy and to create dynamic, broad, and creative unit responses to player input.

%Example Entity Actions [an (E) signifies emergency/fight flight response]
Oxygen Actions
-breath 
-move to air source (E)
-stay calm (E)
-panic and scramble for air (E)
-murder to lessen burden on air supply (E)
-partition government for cheaper air
-protest for cheaper air
-steal air supply
-riot for cheaper air
-ask for air
-beg for air

Water Actions
-move to water source
-stay calm
-panic and scramble for water (E)
-murder to lessen burden on water supply (E)
-partition government for cheaper water
-protest for cheaper water
-riot for cheaper water
-steal water
-ask for water
-beg for water


Food Actions
-eat food
-cook food
-buy food
-ask for food
-beg for food
-steal food 
-go to resturant/cafe/food court/ etc
-go to resturant/cafe/food court/ etc with friend
-murder for food (E)
-kill animal
-partition government for cheaper food
-protest for cheaper food
-riot for cheaper food (E)


Security Actions
-partition government for: militia; police; emergency services
-partition government for: attack neighbouring city; ethnic cleansing
-protest government for: above
-riot for: above
-buy house
-rent house
-squat
-build house
-partition/protest/riot government for: cheaper housing
-purchase firearm/weapon
-partition/protest/riot government for: access to firearms/weapons
-fitness/martial arts
-vigilante group


Belonging Actions
-
-
-
-
-
-





\StartSubSection{Workforce Unit Variables}


%Variables which uniquely identify the unit and determine its capabilites.

Unique ID
[Unique alpha-numeric key](unique to each unit)
[Date and time stamp of %birth](auto generates secondary age variable)[>>][Age]
-These above two generate a unique unit, allowing for keys to be recycled, but date/time stamp will change.

%[First Name and Last Name](may become unique to a cities cultural influences)
[Race/Ancestry]
[Income](fluctuates)
[Workforce Types](multiple, fluctuates)
%[Health Effects](multiple)
[Social Category](?)

%SAMPLE UNIT VARIABLES
Here is a sample list of a units variables:
[UNIQUE ID][DATE/TIME O.B][LATITUDE:LONGITUDE]
[CURRENT ACTION][EMPLOYMENT][OXYGEN][WATER][FOOD][SECURITY][BELONGING]
[NAME][FATHERS RACE][MOTHERS RACE][HEALTH EFFECT][HEALTH EFFECT][EDUCATION TYPE][EDUCATION TYPE][WORKFORCE TYPE][WORKFORCE TYPE]
[WORKFORCE TYPE][WORKFORCE TYPE][INCOME][SOCIAL INTEREST][SOCIAL INTEREST][SOCIAL INTEREST][EDUCATION INTEREST][CONSUMPTION INTEREST][CONSUMPTION INTEREST][CONSUMPTION INTEREST][CONSUMPTION INTEREST][CONSUMPTION INTEREST][CONSUMPTION INTEREST][NEGATIVE]

And here is a sample list of an average labour worker within sub-par city with several Earth influences filling those boxes:
[A245JK89DJ102SK][2013/17/03:2146.46.12][-37.048601,143.739624][MARCHING IN PROTEST][C+R PTY WELDING LINES][O:9][W:4][F:5][S:4][B:8]
[ABUDL JHAKA][MIDDLE EAST][EURASIAN][OVERWEIGHT][UNFIT][PRIMARY][SECONDARY][MANUAL LABOUR][MANAGING][PROTESTER][TAXI DRIVER][31000][COFFEE][CINEMA][RESTURANT][JOURNALISM][FAST FOOD][MOVIES][BANANAS][WATCHES][DESIGNER BOOTS][RIFLES][ASTHEMA]



In this example the unique ID identifies the unit along with its game time date and time stamp. The games longitutude and latitude place and track where the unit is, current action notes what the unit is involved in and the employment variable shows where he is working.
Next we can see the unit has plenty of air (9/10), is a little thirsty, will soon need food, is not feeling secure which is probably why he is protesting, and is fitting in with the crowd with a beloning level of 8. Optional for the game is the tracing of the units bloodlines, showing mothers and fathers ancestry. Next are its health effects which reflect its consumption choices and how the unit may eventually being spending it money (ie on healthcare or medication) or how it will eventually die. Next we can see this unit has obtained general primary and secondary education types which influences the basic skill set it has and therefore determines the work it can obtain. Next we have the workforce types it is involved in which denotes where is has been working, where it is working, so its skills, along with any other non-social activity, in this case he is actively protesting. We can then see he is currently earning the equivalent of 31,000 a year in wages and prefers to socialise by having coffee and going to movies resturants. He also has an education preference for journalism which means if it was available and he could still meet his needs he would be studying journalism instead of what he is currently doing. He has 
a consumption preference for fast food, movies, bananas, watches, designer boots, and rifles; this means he will spend his disposable income on these goods in no particular order. Note that in this above exmaple, education types does not accurately depict a units education, education types will actually be made up of many types forming an education as chosen by the player. Primary, secondary, etc are just the games default educational system in place without any player management.
	

\StartSubSection{Generation of Unit Variables}
Of the above variables, some are automatically asigned to the unit, such as ID, name, date of birth, and location, however others are generated and changed throughout the game. Need variables are continually changing as influences from the environment affect the unit and may go from being low when the need has not been met for a while to being high when the unit has just had the need satisfied. In addition some variables may come and go completely, such consumption or social interests. 

Consumption interests are variables that determine where a unit will spend money and/or time when this is available, it also determines preferences when there are multiple options available. The consumption interest variable is specifically created for a unit when it is exposed to media (television or similar, public advertising, radio) or social interaction, both of which are generated when the unit does not have consumption interests already in place. 

Each unit has an amount of consumption interests determined by the age variable of the unit, this is to limit a toddler having a complete set of consumption variables too early. The variables are continually changing throughout the units lifespan and are based on the highest averages of external influences which cause consumption interest generation. So if a unit has 5 consumption variables at age 25, these will be the 5 variables that have had the most influence over the units environment. It's highest consumption varibale might be accredited organic food, the next might be "clown foods" inc, then shot guns, pure water, and finally fast cars. In addition as each in-game second (or minute or whatever period of time is deemed appropriate) passes the units average yearly ratios of influence are being calculated, with the variables changing places or being replaced completely. While it is possible that this could cause rapid change of variables, the gradual nature in which policy and building changes, an these changes in turn influece units should then be causing the influence ratios to be changing just as gradually.   


[Age][>>>][Amount of consumption interests] 




\StartSubSection{Needs and Actions Variance}
There are two broad categories which influences a units needs. The first is the units own actions, drinking water, eating food, getting a job, visiting a friend, all reduce the need to obtain these things, while excerising increases the need for food and water, and working long hours in isolation increases the need for belonging. The second is external or environmental effects such as propaganda, lack of clean air, food with poor nutitional content, or the action and views of other units. All these happenings have varying amounts of influence on the units needs which in turn effect the manner in which the unit acts to satisfy these needs.

In order to correctly map the degree of influence various needs have on a unit and the degree to which a units actions relieves or increased its needs, all of the variables effecting the needs of a unit are mapped with a point system quantively describing the degree to which needs effect the unit and the degree to which an action or event influences a units needs. Some of these will be wide reaching, effecting many units, such as chemicals in the cities water supply or television, other will be specific, such as health consequences from working shift work. In turn all these effect unit needs which cause unit actions, and may also create or change unit variables (mentioned above). 



Influence:
-Workforce Types
-Food Types
-Actions
-Information Types
-Economic 
-Drugs
-Social
-Ostracism



%SAMPLE  (Before) / (after)
%Input		Effects
%		security	food		belongingness				
%Crack Cocaine	+2 / -10	+10 / -5	+10 / -10			 
%
%Running	+1		-2		
%
%
%Enemployed	-3				-2*
%
%
%
%
%*depends on social system.





\StartSubSection{Workforce Consumption}
As discussed, in Avaneya a workforce unit is not just a unit producing goods and services, it is also a unit having any influence over the direction of the economy. This includes its patters of consumption. As each workforce unit is represented by an individual unit in game, moving and performing tasks, as this unit moves through game space it is consuming resources such as air, water, food, and minerals and like subtances through consumerable resources. What a unit can consume is based in part by what is available and what its needs and action influences are.



 



%DISCARDED NOTES:
There will be underlying conditions that will determine what a cities workforce is comprised of and players will have certain degress of influence over the direction the workforce takes but if certain conditions aren't met the citizenry may not accept or appreciate the current workforce type with the consequences of innaction by the player being varried. How players make these conditions and the direction they take with there workforce is a choice made in the sandbox style of play, it is a creation. They can create anything from human/animal labour pool to complete nanocracies with robotic workforces and humans indulging in deccandent pleasure palaces. The player can manage his city through inacting certain ordinance, directing funds, and other general city management processes including placement of civic buildings and diplomacy. 






=======
% This is part of the Avaneya Project Crew Handbook.
% Copyright (C) 2010-2013 Cartesian Theatre <info@cartesiantheatre.com>.
% See the file Copying for details on copying conditions.

% Game Mechanics chapter...
%\StartSection{Workforce}

\StartSubSection{Workforce}
%INTRODUCTION INTRODUCTION
What a city can produce, how much it can produce, and what it is capable of producing are largely byproducts of its workforce however in Avaneya workforce also determines things such as consumption patterns and criminal behaviour. This section discusses the variety of workforces, what they represent, how they are created, maintained, and influenced, and what they are capable of. 

A workforce unit is essentially a singular individual of the population, an animal, or a piece of technology that either exists at a certain level of education, training, or lack there of, or is built or bred to a variety of types, scales, and qualities (see section Workforce Types). Workforce includes all labour factors driving the economy and extends to consumers (and even the homeless, criminals, and the unemployed). The workforce of a city then is all units in the city capable of exerting influence on the direction in which resources or money flows. A labour workfoce requires certain factors to bring it into fruition which may include a certain level of education, access to work, and restrictions to certain aspects of these. An intellectual workforce requires different factors to develop then a manual labour workforce. While the two are not mutually exclusive it may or may not be more difficult to create multiple classes in a single city. In adition otherwise different workforces may share types aspects of influence over the economy, known as workforce types. For instance a workforce unit in manufacturing may have the same education level and by the same food as a workforce unit in commerce. These two units share a couple of "workforce types" (qualative education and consumer choices) but have been influenced by other means to different outcomes. A cities workforce is influenced through players building, inacting ordinance, and other various city policies along with diplomacy. The method in which various workforce units are shaped and changed through the development of a players city are complex and discussed in depth through the various subsections of the section.

\StartSubSection{Visual Workforce}
A cities multi faceted workforce is actually made up of varying amounts (many thousands) of individuals, humans, animals, and machines, and are depicted visually by people or objects performing tasks or moving around the city. Each workforce unit is capable of influencing the direction the greater workforce and the city take and each unit is an individual human, animal, or some form of machine involved in labour. Visually, workforce units need to be represented as: humans of varying races, ages, sizes, and wearing various clothing; animals of varying species and breeds; and machines of varying types and degrees of repair. Depending on the zoomable scale all these units must be able to be seen performing actions and influencing there surroundings and interacting with other units as well as their environment. A large variety of animations will be required for the number of potential interactions. 

\StartSubSection{Workforce Movement}
Workforce is a dynamic "resource" and can move city wide (if allowed) and globally (if they are not trapped in the city). If your workforce requires cheap housing and this becomes easier to gain in another city, many may migrate to this city seeking better opportunities. Likewise, they can move within your city to different opportunities and movement both within a city and between cities can be controlled to an extent to enable better or worse access to certain locations, increasing and decreasing travel time, or diverting travel for other means. This is just as true for mechinical workforce which may be moveable on its own, or through human interaction. 

Indeed a lot of a workforce units interaction and influence over the cities economy and general development will have to do with its ability to move. How far a unit can travel while meeting all its needs (discussed below) will determine this influence. Depending on the unit type it may be walking to work littering, going out to lunch, moving between two locations at work (goods transport), walking in a biodome paddock eating grass, or travelling along an assembly line. Also, the unit may be in cohesion with other units performing some of these tasks, for example eating lunch with a group of friends, building a tank, baking bread, or theater or dancing. 

The means in which this movement is and can be performed is based on the units needs and the vicinity to be able to meet the needs. Movement can also be restricted through placing private or of limit areas, constructing walk ways in cetain areas and directions, making some direction longer and others shorter, charging to travel some directions, and creating and limiting private and public transport options. 

\StartSubSection{Workforce Consumption}
As discussed, in Avaneya a workforce unit is not just a unit producing goods and services, it is also a unit having any influence over the direction of the economy. This includes its patters of consumption


\StartSubSection{Workforce Bases}
%Workforce Bases
There are three different fundamental types of workforce. There is a mechanical workforce made up of machines, computers, and other similar automatons. There is a human workforce made up of the biological and social entities which are humans. And there is a animal workforce which is also biological but for the games sake is not influenced by the social fabric of a city. These are known as workforce bases. 

The mechanical workforce requires research and development into avenues such as robotics and mechanical engineering with a social system in place requiring either the production or the import of mechanical systems. The human workforce is made up of human labour (physical, mental, and both) performing not only the tasks required by the society, but also tasks or actions that may detriment the society. As the cities population is influenced by numerous external and internal agencies such as media, education, drugs, and fashionable consumption it is convoluted and multifaceted, and the style of play implemented by the player, other players, and NPCs, will help to determine the quality and quanity of a cities human workforce. The animal workfroce is similar to the human workforce in its being a biological entity, and similar to a mechanical workforce in its one dimensional aspect of usefulness. Cities in Avaneya will generally have a mix of these three base workforces and each of these in term will have many varying types (see subsection on Workforce Types). 

In addition to a city producing its own workforce, it may purchase machinery and animals, attract population, and hire individuals directly from other cities. While the city may purchase certain workforces at some stage or another, purchasing of individuals and machinery generally takes place by businesses. The main method in which cities attract additional population (apart from reproduction) is through the cities governing player creating opportunities which can keep population in the city or attracting population from other cities. 


\StartSubSection{Workforce Units Needs}
As mentioned briefly above a cities workforce is made up of many individual units. Each of these units can be a part of one or more workforce types and require certain conditions or needs met in order to be created, maintained, and have degrees of functionality. The human workforce has needs which differentiates them from the other two workforce bases (animals and machines). While machines require maintainence and energy, and animals require food, water, and shelter, in Avaneya humans interact with the players decisions, political or otherwise. Animals and machines are only effected by ordinance and taxation through humans. 

The needs an entity, or individual human, has determine where it will go, what it will do, and how it will do it. If we take sections [] accuracy into account we can view humans in a largely tabula rasa state, with no ill will except to maintain there own needs. Further to this, it can be seen that the nature in which humans interact and the means in which they satisfy their needs is determined by something outside of there biological state. 

The human needs consist of:
Oxygen: breathable air, a human can survive for less than 3 minutes without suffering potentially serious consequences. 

Water: clean drinking water. The human body can survive upto a week without water (provided good conditions). 

Food: sustainence and nutrtion. It is essential to not only have something to eat but that it contains the various essential nutrients to support human life. A human can last several weeks without food provided they have water.

Security: shelter/housing, feeling safe in surroundings, and feeling secure about the future.

Belonging: feeling of making a contribution, having people to talk to, being a part of something more than hirself.

In light of the above paragraphs, it can be seen that it is quite simple to satisfy a humans security needs. However, continually bombard hir with messages of insecurity and threats despite evidence to the contrary and it will seek ever greater means to satisfy this need, perhaps waging unnecessary wars against otherwise peaceful peoples. This is just an example of how a player can create or solve disonance in human needs and scenarios the game can implement to challenge players.



\StartSubsection{workforce Types}
The dynamic and multifaceted nature of workforce may become more evident when reading about the various workforce types. From the list below it can be seen that workforce comprises what the population is doing in reaction to the cities governance or that of other cities and also what stage of development a unit is at. Some types of workforce are easy to maintain in a state desired by players, such as industrial machinery. This needs only to be created, maintained, and given a source of energy. Other workforce types are more specific such as "the high school teacher" or "motor mechanic". These require specific circumstances surrounding a population, with just the right levels of affluence and education. Whats more is human workforce types also share needs with other human workforce types and are dependant on eachother to provide those needs. This makes for a complex array of interaction between workforces, and when combined with taxation levels, ordinance, and geopolitical factors the complexitity is multiplied. In addition workforce types applied to a workforce unit can overlap. A unit could be tertiary educated but unemployed, or a physical labourer with a secondary education. 

Also individuals may move between workforce types, as in moving through education levels or going from unemployed to employed. Some factors which change workforce types are done through time, others done through city management. The game must track each workforce unit, its movements, its needs, its progress (or degression) through education and work, its workforce types, its use of drugs, or its breaking the law. It must track this and the interaction with other workforce units, and the impact this all has on other units, (either directly through actions or indirectly through policy change or similar) and on the environment and the cities statistics, spending, and policies.


%Workforce Types
-Homeless
-Unemployed
-Physical labour (varying levels/classes, ie production, mining)
-Mental labour (variious)
-Education (primary, secondary, tertiary, post tertiary [not complete, see education types])
-Employment skill (varying types depending on field, ie years of experience)
-Training (varying types, ie separate training that a business or government may provide)
-Criminal (varying types)
-Militia
-Consumer (varying types depending on influences, ie social eco status, cultural)
-Mechanical (varying) 

-Protester
-Animal (varying)
-Age (varying)
-Drug user (various)
-Drug addicted (various)





\StartSubSection{Transitioning Between Workforce Types}


\StartSubSection{Influencing Human Workforce}
%
How people are raised, educated, and conditioned through experience largely determines their consumer patterns and may be reflected in a population's consumer choices. In game, workforce is influenced through players building, inacting ordinance, and other various city policies along with diplomacy. The method in which various workforce units are shaped and changed through the development of a players city are complex and discussed in depth below.
%

There are various methods in which the human workforces are influenced to take a certain direction. The main method in which a player can influence his own or someone elses workforce is through economic means such as taxation or ordinance, although other methods may exist such as brute force or through environmental coercion. Additional methods include the use of the media and/or propaganda which can be initiated by the player, another player, or AI representing Earth interests.

The means in which the workforce itself is influenced is through their various needs (see section Workforce Units Needs). Human workforce units will go about striving to fulfill there needs by a means which is partially predetermined by default programming and partially determine by the needs needing be met and the degree to which they are wanting. Usually, all things being good, this is peaceful but ultimately depends on the situation at hand and at the players influence. Through enacting ordinance, taxation, and various diplomatic relations the player creates variance in the workforce needs levels, and over time the method in which each workforce unit meets its needs.

An example may be the implementation of a higher tax level for a certain group of people. This may place strain on there income, causing crime, violence, and other problems for the city. However if this higher tax level is offset with social funding for this group, giving them other means to obtain there needs then the previous situation with crime level increase may not occur. 

Another example is allowing television and private broadcast. While cheaper for the government and good for business, this system may become abused with continual broadcasts leaving people feeling that their belongingness need is to be fulfilled by wearing the same clothes as some Earth celebrity. 

The above two examples are simple occurances. In avaneya there will be multiple levels of cultural and economic variance influencing different cultural groups to different degrees causing players to deal with this to keep all aspects of the city functioning while still trying to achieve player goals such as technological innovation and maintain an independent yet peaceful existance with Earth. From this it can be seen the the AI of human workforce will be quite involved.   

%Discuss how effected and how entities will react.
The AI of entities will respond to a variety of inputs from the greater enviroment. The manner in which each entity attempts to reach satisfaction of its needs is determined from this environmental input. To describe the process, an individual entity will go about in its default state provided it has access to all its essentials. When this access is threatened or blocked it will attempt to right the situation by some means. The means in which it can right the situation depends on the enviroment, namely economic, political, and location factors, as well as its education/knowledge and social status, and the need that is being threatened. Further to this, an entity may not even be aware of some changes or threats and therefore might not take action

Lets go back to our example on a threat to security. One group of humans with similar workforce types might make business deals to get it out of the city or off the planet. Another might petition government to increase militia spending. Another might encourage the people to take to arms and fight. And another might join the militia. Outliers may take to vigilante justice or investigate to discover possible conspiracies. Of course the degree to which these can occur depend largely on previous player input to the population. There is a multitude of means in which entities will respond to player sitations that threaten or improve their circumstances and these are unique to the needs being effected and resource the unit has available to make a change or maintain safety.

%multiple threats to multiple needs
When there are multiple threats to the same entity, it will respond in order of importance. Clearly, air, then water, then food, then security, then belonging. Entities will not respond to their need to belond when they are dieing from lack of oxygen.

%Natural State
In its natural state a workforce unit is only trying to satisfy its needs. The manner in which the unit can do this is chosen by the method in which the city is managed. An unit seeking food could naturally go into the wilderness and find fruit etc., however in Avaneya, on Mars, this is not possible and the unit therefore finds other means of getting food. This may be going to a market, going out to lunch, or stealing from a grocery store. In its default state the unit pursues tasks, or actions, in accordance with its needs.

To determine the type of action an unit will undertake is complex. It is a matter or factoring in its various needs at the time along with its workforce type and other variables such as income level, social status, and location. In this way we determine whether an unit goes home to cook food for dinner, or goes out to dinner with friends, or goes home has dinner then goes bowling. Factors such as the units income or money, degree of hunger, location to food, food quality, need to belong, number of friends (or daily interactions), neighbourhood safety, "governmental" dietary guidelines, work security, and time available, all converge to create the units action. These are automatically calculated by the program and are known as actions or specifically as unit actions. The actions taken by units are classed into categories based on needs. 

Not everything will be required to have an action, breathing and drinking water are just factored into resource use and will not evoke a reaction unless the unit is unable to obtain this resource. The main use for entity ations is to document visually and statistically the units input into the economy and to create dynamic, broad, and creative unit responses to player input.

%Example Entity Actions
Oxygen Actions
-breath 
-move to air source
-stay calm
-panic and scramble for air
-murder to lessen burden on air supply
-partition government for cheaper air

Water Actions
-move to water source
-stay calm
-panic and scramble for water
-murder to lessen burden on water supply
-partition government for cheaper water


Food Actions



Security Actions



Belonging Actions

\StartSubSection{Workforce Unit Variables}


%Variables which uniquely identify the unit and determine its capabilites.

Unique ID
[Unique alpha-numeric key](unique to each unit)
[Date and time stamp of %birth](auto generates secondary age variable)[>>][Age]
-These above two generate a unique unit, allowing for keys to be recycled, but date/time stamp will change.

%[First Name and Last Name](may become unique to a cities cultural influences)
[Race/Ancestry]
[Income](fluctuates)
[Workforce Types](multiple, fluctuates)
%[Health Effects](multiple)
[Social Category](?)

%SAMPLE UNIT VARIABLES
Here is a sample list of a units variables:
[UNIQUE ID][DATE/TIME O.B][LATITUDE:LONGITUDE]
[CURRENT ACTION][EMPLOYMENT][OXYGEN][WATER][FOOD][SECURITY][BELONGING]
[NAME][FATHERS RACE][MOTHERS RACE][HEALTH EFFECT][HEALTH EFFECT][EDUCATION TYPE][EDUCATION TYPE][WORKFORCE TYPE][WORKFORCE TYPE]
[WORKFORCE TYPE][WORKFORCE TYPE][INCOME][SOCIAL INTEREST][SOCIAL INTEREST][SOCIAL INTEREST][EDUCATION INTEREST][CONSUMPTION INTEREST][CONSUMPTION INTEREST][CONSUMPTION INTEREST][CONSUMPTION INTEREST][CONSUMPTION INTEREST][CONSUMPTION INTEREST]

And here is a sample list of an average labour worker within sub-par city with several Earth influences filling those boxes:
[A245JK89DJ102SK][2013/17/03:2146.46.12][-37.048601,143.739624][MARCHING IN PROTEST][C+R PTY WELDING LINES][O:9][W:4][F:5][S:4][B:8]
[ABUDL JHAKA][MIDDLE EAST][EURASIAN][OVERWEIGHT][UNFIT][PRIMARY][SECONDARY][MANUAL LABOUR][MANAGING][PROTESTER][TAXI DRIVER][31000][COFFEE][CINEMA][RESTURANT][JOURNALISM][FAST FOOD][MOVIES][BANANAS][WATCHES][DESIGNER BOOTS][RIFLES]



In this example, the unique ID identifies the unit along with its game time date and time stamp. The games longitutude and latitude place and track where the unit is, current action notes what the unit is involved in and the employment variable shows where he is working.
 Next we can see the unit has plenty of air (9/10), is a little thirsty, will soon need food, is not feeling secure which is probably why he is protesting, and is fitting in with the crowd with a beloning level of 8. Optional for the game is the tracing of the units bloodlines, showing mothers and fathers ancestry. Next are its health effects which reflect its consumption choices and how the unit may eventually being spending it money (ie on healthcare or medication) or how it will eventually die. Next we can see this unit has obtained general primary and secondary education types which influences the basic skill set it has and therefore determines the work it can obtain. Next we have the workforce types it is involved in which denotes where is has been working, where it is working, so its skills, along with any other non-social activity, in this case he is actively protesting. We can then see he is currently earning the equivalent of 31,000 a year in wages and prefers to socialise by having coffee and going to movies resturants. He also has an education preference for journalism which means if it was available and he could still meet his needs he would be studying journalism instead of what he is currently doing. He has 
a consumption preference for fast food, movies, bananas, watches, designer boots, and rifles; this means he will spend his disposable income on these goods in no particular order. Note that in this above exmaple, education types does not accurately depict a units education, education types will actually be made up of many types forming an education as chosen by the player. Primary, secondary, etc are just the games default educational system in place without any player management.
	

\StartSubSection{Generation of Unit Variables}
Of the above variables, some are automatically asigned to the unit, such as ID, name, date of birth, and location, however others are generated and changed throughout the game. Need variables are continually changing as influences from the environment affect the unit and may go from being low when the need has not been met for a while to being high when the unit has just had the need satisfied. In addition some variables may come and go completely, such consumption or social interests. 

Consumption interests are variables that determine where a unit will spend money and/or time when this is available, it also determines preferences when there are multiple options available. The consumption interest variable is specifically created for a unit when it is exposed to media (television or similar, public advertising, radio) or social interaction, both of which are generated when the unit does not have consumption interests already in place. 

Each unit has an amount of consumption interests determined by the age variable of the unit, this is to limit a toddler having a complete set of consumption variables too early. The variables are continually changing throughout the units lifespan and are based on the highest averages of external influences which cause consumption interest generation. So if a unit has 5 consumption variables at age 25, these will be the 5 variables that have had the most influence over the units environment. It's highest consumption varibale might be accredited organic food, the next might be "clown foods" inc, then shot guns, pure water, and finally fast cars. In addition as each in-game second (or minute or whatever period of time is deemed appropriate) passes the units average yearly ratios of influence are being calculated, with the variables changing places or being replaced completely. While it is possible that this could cause rapid change of variables, the gradual nature in which policy and building changes, an these changes in turn influece units should then be causing the influence ratios to be changing just as gradually.   


[Age][>>>][Amount of consumption interests] 




\StartSubSection{Needs and Actions Variance}
There are two broad categories which influences a units needs. The first is the units own actions, drinking water, eating food, getting a job, visiting a friend, all reduce the need to obtain these things, while excerising increases the need for food and water, and working long hours in isolation increases the need for belonging. The second is external or environmental effects such as propaganda, lack of clean air, food with poor nutitional content, or the action and views of other units. All these happenings have varying amounts of influence on the units needs which in turn effect the manner in which the unit acts to satisfy these needs.

In order to correctly map the degree of influence various needs have on a unit and the degree to which a units actions relieves or increased its needs, all of the variables effecting the needs of a unit are mapped with a point system quantively describing the degree to which needs effect the unit and the degree to which an action or event influences a units needs. Some of these will be wide reaching, effecting many units, such as chemicals in the cities water supply or television, other will be specific, such as health consequences from working shift work. In turn all these effect unit needs which cause unit actions, and may also create or change unit variables (mentioned above). 



Influence:
-Workforce Types
-Food Types
-Actions
-Information Types
-Economic 
-Drugs
-Social
-Ostracism



%SAMPLE  (Before) / (after)
%Input		Effects
%		security	food		belongingness				
%Crack Cocaine	+2 / -10	+10 / -5	+10 / -10			 
%
%Running	+1		-2		
%
%
%Enemployed	-3				-2*
%
%
%
%
%*depends on social system.









 



%DISCARDED NOTES:
There will be underlying conditions that will determine what a cities workforce is comprised of and players will have certain degress of influence over the direction the workforce takes but if certain conditions aren't met the citizenry may not accept or appreciate the current workforce type with the consequences of innaction by the player being varried. How players make these conditions and the direction they take with there workforce is a choice made in the sandbox style of play, it is a creation. They can create anything from human/animal labour pool to complete nanocracies with robotic workforces and humans indulging in deccandent pleasure palaces. The player can manage his city through inacting certain ordinance, directing funds, and other general city management processes including placement of civic buildings and diplomacy. 






=======
% This is part of the Avaneya Project Crew Handbook.
% Copyright (C) 2010-2013 Cartesian Theatre <kip@thevertigo.com>.
% See the file Copying for details on copying conditions.

% Game Mechanics chapter...
%\StartSection{Workforce}

\StartSubSection{Workforce}
%INTRODUCTION INTRODUCTION
What a city can produce, how much it can produce, and what it is capable of producing are largely byproducts of its workforce however in Avaneya workforce also determines things such as consumption patterns and criminal behaviour. This section discusses the variety of workforces, what they represent, how they are created, maintained, and influenced, and what they are capable of. 

A workforce unit is essentially a singular individual of the population, an animal, or a piece of technology that either exists at a certain level of education, training, or lack there of, or is built or bred to a variety of types, scales, and qualities (see section Workforce Types). Workforce includes all labour factors driving the economy and extends to consumers (and even the homeless, criminals, and the unemployed). The workforce of a city then is all units in the city capable of exerting influence on the direction in which resources or money flows. A labour workfoce requires certain factors to bring it into fruition which may include a certain level of education, access to work, and restrictions to certain aspects of these. An intellectual workforce requires different factors to develop then a manual labour workforce. While the two are not mutually exclusive it may or may not be more difficult to create multiple classes in a single city. In adition otherwise different workforces may share types aspects of influence over the economy, known as workforce types. For instance a workforce unit in manufacturing may have the same education level and by the same food as a workforce unit in commerce. These two units share a couple of "workforce types" (qualative education and consumer choices) but have been influenced by other means to different outcomes. A cities workforce is influenced through players building, inacting ordinance, and other various city policies along with diplomacy. The method in which various workforce units are shaped and changed through the development of a players city are complex and discussed in depth through the various subsections of the section.

\StartSubSection{Visual Workforce}
A cities multi faceted workforce is actually made up of varying amounts (many thousands) of individuals, humans, animals, and machines, and are depicted visually by people or objects performing tasks or moving around the city. Each workforce unit is capable of influencing the direction the greater workforce and the city take and each unit is an individual human, animal, or some form of machine involved in labour. Visually, workforce units need to be represented as: humans of varying races, ages, sizes, and wearing various clothing; animals of varying species and breeds; and machines of varying types and degrees of repair. Depending on the zoomable scale all these units must be able to be seen performing actions and influencing there surroundings and interacting with other units as well as their environment. A large variety of animations will be required for the number of potential interactions. 

\StartSubSection{Workforce Movement}
Workforce is a dynamic "resource" and can move city wide (if allowed) and globally (if they are not trapped in the city). If your workforce requires cheap housing and this becomes easier to gain in another city, many may migrate to this city seeking better opportunities. Likewise, they can move within your city to different opportunities and movement both within a city and between cities can be controlled to an extent to enable better or worse access to certain locations, increasing and decreasing travel time, or diverting travel for other means. This is just as true for mechinical workforce which may be moveable on its own, or through human interaction. 

Indeed a lot of a workforce units interaction and influence over the cities economy and general development will have to do with its ability to move. How far a unit can travel while meeting all its needs (discussed below) will determine this influence. Depending on the unit type it may be walking to work littering, going out to lunch, moving between two locations at work (goods transport), walking in a biodome paddock eating grass, or travelling along an assembly line. Also, the unit may be in cohesion with other units performing some of these tasks, for example eating lunch with a group of friends, building a tank, baking bread, or theater or dancing. 

The means in which this movement is and can be performed is based on the units needs and the vicinity to be able to meet the needs. Movement can also be restricted through placing private or of limit areas, constructing walk ways in cetain areas and directions, making some direction longer and others shorter, charging to travel some directions, and creating and limiting private and public transport options. 


\StartSubSection{Workforce Bases}
%Workforce Bases
There are three different fundamental types of workforce. There is a mechanical workforce made up of machines, computers, and other similar automatons. There is a human workforce made up of the biological and social entities which are humans. And there is a animal workforce which is also biological but for the games sake is not influenced by the social fabric of a city. These are known as workforce bases. 

The mechanical workforce requires research and development into avenues such as robotics and mechanical engineering with a social system in place requiring either the production or the import of mechanical systems. The human workforce is made up of human labour (physical, mental, and both) performing not only the tasks required by the society, but also tasks or actions that may detriment the society. As the cities population is influenced by numerous external and internal agencies such as media, education, drugs, and fashionable consumption it is convoluted and multifaceted, and the style of play implemented by the player, other players, and NPCs, will help to determine the quality and quanity of a cities human workforce. The animal workfroce is similar to the human workforce in its being a biological entity, and similar to a mechanical workforce in its one dimensional aspect of usefulness. Cities in Avaneya will generally have a mix of these three base workforces and each of these in term will have many varying types (see subsection on Workforce Types). 

In addition to a city producing its own workforce, it may purchase machinery and animals, attract population, and hire individuals directly from other cities. While the city may purchase certain workforces at some stage or another, purchasing of individuals and machinery generally takes place by businesses. The main method in which cities attract additional population (apart from reproduction) is through the cities governing player creating opportunities which can keep population in the city or attracting population from other cities. 


\StartSubSection{Workforce Units Needs}
As mentioned briefly above a cities workforce is made up of many individual units. Each of these units can be a part of one or more workforce types and require certain conditions or needs met in order to be created, maintained, and have degrees of functionality. The human workforce has needs which differentiates them from the other two workforce bases (animals and machines). While machines require maintainence and energy, and animals require food, water, and shelter, in Avaneya humans interact with the players decisions, political or otherwise. Animals and machines are only effected by ordinance and taxation through humans. 

The needs an entity, or individual human, has determine where it will go, what it will do, and how it will do it. If we take sections [] accuracy into account we can view humans in a largely tabula rasa state, with no ill will except to maintain there own needs. Further to this, it can be seen that the nature in which humans interact and the means in which they satisfy their needs is determined by something outside of there biological state. 

The human needs consist of:
Oxygen: breathable air, a human can survive for less than 3 minutes without suffering potentially serious consequences. 

Water: clean drinking water. The human body can survive upto a week without water (provided good conditions). 

Food: sustainence and nutrtion. It is essential to not only have something to eat but that it contains the various essential nutrients to support human life. A human can last several weeks without food provided they have water.

Security: shelter/housing, feeling safe in surroundings, and feeling secure about the future.

Belonging: feeling of making a contribution, having people to talk to, being a part of something more than hirself.

In light of the above paragraphs, it can be seen that it is quite simple to satisfy a humans security needs. However, continually bombard hir with messages of insecurity and threats despite evidence to the contrary and it will seek ever greater means to satisfy this need, perhaps waging unnecessary wars against otherwise peaceful peoples. This is just an example of how a player can create or solve disonance in human needs and scenarios the game can implement to challenge players.



\StartSubsection{workforce Types}
The dynamic and multifaceted nature of workforce may become more evident when reading about the various workforce types. From the list below it can be seen that workforce comprises what the population is doing in reaction to the cities governance or that of other cities and also what stage of development a unit is at. Some types of workforce are easy to maintain in a state desired by players, such as industrial machinery. This needs only to be created, maintained, and given a source of energy. Other workforce types are more specific such as "the high school teacher" or "motor mechanic". These require specific circumstances surrounding a population, with just the right levels of affluence and education. Whats more is human workforce types also share needs with other human workforce types and are dependant on eachother to provide those needs. This makes for a complex array of interaction between workforces, and when combined with taxation levels, ordinance, and geopolitical factors the complexitity is multiplied. In addition workforce types applied to a workforce unit can overlap. A unit could be tertiary educated but unemployed, or a physical labourer with a secondary education. 

Also individuals may move between workforce types, as in moving through education levels or going from unemployed to employed. Some factors which change workforce types are done through time, others done through city management. The game must track each workforce unit, its movements, its needs, its progress (or degression) through education and work, its workforce types, its use of drugs, or its breaking the law. It must track this and the interaction with other workforce units, and the impact this all has on other units, (either directly through actions or indirectly through policy change or similar) and on the environment and the cities statistics, spending, and policies.


%Workforce Types
-Homeless
-Unemployed
-Physical labour (varying levels/classes, ie production, mining)
	(a physical labour type for each task needed in city) 
-Mental labour (variious)
-Education (primary, secondary, tertiary, post tertiary [not complete, see education types])
-Employment skill (varying types depending on field, ie years of experience)
-Training (varying types, ie separate training that a business or government may provide)
-Criminal (varying types)
-Militia
-Consumer (varying types depending on influences, ie social eco status, cultural)
-Mechanical (varying) 

-Protester
-Animal (varying)
-Age (varying)
-Drug user (various)
-Drug addicted (various)





\StartSubSection{Transitioning Between Workforce Types}
%Unit gets skill and either upgrades workforce type or adds a workforce type.
%Hows does units get skill - interests - needs - continual work - education





\StartSubSection{Influencing Human Workforce}
%
How people are raised, educated, and conditioned through experience largely determines their consumer patterns and may be reflected in a population's consumer choices. In game, workforce is influenced through players building, inacting ordinance, and other various city policies along with diplomacy. The method in which various workforce units are shaped and changed through the development of a players city are complex and discussed in depth below.
%

There are various methods in which the human workforces are influenced to take a certain direction. The main method in which a player can influence his own or someone elses workforce is through economic means such as taxation or ordinance, although other methods may exist such as brute force or through environmental coercion. Additional methods include the use of the media and/or propaganda which can be initiated by the player, another player, or AI representing Earth interests.

The means in which the workforce itself is influenced is through their various needs (see section Workforce Units Needs). Human workforce units will go about striving to fulfill there needs by a means which is partially predetermined by default programming and partially determine by the needs needing be met and the degree to which they are wanting. Usually, all things being good, this is peaceful but ultimately depends on the situation at hand and at the players influence. Through enacting ordinance, taxation, and various diplomatic relations the player creates variance in the workforce needs levels, and over time the method in which each workforce unit meets its needs.

An example may be the implementation of a higher tax level for a certain group of people. This may place strain on there income, causing crime, violence, and other problems for the city. However if this higher tax level is offset with social funding for this group, giving them other means to obtain there needs then the previous situation with crime level increase may not occur. 

Another example is allowing television and private broadcast. While cheaper for the government and good for business, this system may become abused with continual broadcasts leaving people feeling that their belongingness need is to be fulfilled by wearing the same clothes as some Earth celebrity. 

The above two examples are simple occurances. In avaneya there will be multiple levels of cultural and economic variance influencing different cultural groups to different degrees causing players to deal with this to keep all aspects of the city functioning while still trying to achieve player goals such as technological innovation and maintain an independent yet peaceful existance with Earth. From this it can be seen the the AI of human workforce will be quite involved.   

%Discuss how effected and how entities will react.
The AI of entities will respond to a variety of inputs from the greater enviroment. The manner in which each entity attempts to reach satisfaction of its needs is determined from this environmental input. To describe the process, an individual entity will go about in its default state provided it has access to all its essentials. When this access is threatened or blocked it will attempt to right the situation by some means. The means in which it can right the situation depends on the enviroment, namely economic, political, and location factors, as well as its education/knowledge and social status, and the need that is being threatened. Further to this, an entity may not even be aware of some changes or threats and therefore might not take action

Lets go back to our example on a threat to security. One group of humans with similar workforce types might make business deals to get it out of the city or off the planet. Another might petition government to increase militia spending. Another might encourage the people to take to arms and fight. And another might join the militia. Outliers may take to vigilante justice or investigate to discover possible conspiracies. Of course the degree to which these can occur depend largely on previous player input to the population. There is a multitude of means in which entities will respond to player sitations that threaten or improve their circumstances and these are unique to the needs being effected and resource the unit has available to make a change or maintain safety.

%Multiple threats to multiple needs
When there are multiple threats to the same entity, it will respond in order of importance. Clearly, air, then water, then food, then security, then belonging. Entities will not respond to their need to belond when they are dieing from lack of oxygen.

%Natural State
In its natural state a workforce unit is only trying to satisfy its needs. The manner in which the unit can do this is chosen by the method in which the city is managed. An unit seeking food could naturally go into the wilderness and find fruit etc., however in Avaneya, on Mars, this is not possible and the unit therefore finds other means of getting food. This may be going to a market, going out to lunch, or stealing from a grocery store. In its default state the unit pursues tasks, or actions, in accordance with its needs.

To determine the type of action an unit will undertake is complex. It is a matter or factoring in its various needs at the time along with its workforce type and other variables such as income level, social status, and location. In this way we determine whether an unit goes home to cook food for dinner, or goes out to dinner with friends, or goes home has dinner then goes bowling. Factors such as the units income or money, degree of hunger, location to food, food quality, need to belong, number of friends (or daily interactions), neighbourhood safety, "governmental" dietary guidelines, work security, and time available, all converge to create the units action. These are automatically calculated by the program and are known as actions or specifically as unit actions. The actions taken by units are classed into categories based on needs. 

Not everything will be required to have an action, breathing and drinking water are just factored into resource use and will not evoke a reaction unless the unit is unable to obtain this resource. The main use for entity ations is to document visually and statistically the units input into the economy and to create dynamic, broad, and creative unit responses to player input.

%Example Entity Actions
Oxygen Actions
-breath 
-move to air source
-stay calm
-panic and scramble for air
-murder to lessen burden on air supply
-partition government for cheaper air

Water Actions
-move to water source
-stay calm
-panic and scramble for water
-murder to lessen burden on water supply
-partition government for cheaper water


Food Actions



Security Actions



Belonging Actions

\StartSubSection{Workforce Unit Variables}


%Variables which uniquely identify the unit and determine its capabilites.

Unique ID
[Unique alpha-numeric key](unique to each unit)
[Date and time stamp of %birth](auto generates secondary age variable)[>>][Age]
-These above two generate a unique unit, allowing for keys to be recycled, but date/time stamp will change.

%[First Name and Last Name](may become unique to a cities cultural influences)
[Race/Ancestry]
[Income](fluctuates)
[Workforce Types](multiple, fluctuates)
%[Health Effects](multiple)
[Social Category](?)

%SAMPLE UNIT VARIABLES
Here is a sample list of a units variables:
[UNIQUE ID][DATE/TIME O.B][LATITUDE:LONGITUDE]
[CURRENT ACTION][EMPLOYMENT][OXYGEN][WATER][FOOD][SECURITY][BELONGING]
[NAME][FATHERS RACE][MOTHERS RACE][HEALTH EFFECT][HEALTH EFFECT][EDUCATION TYPE][EDUCATION TYPE][WORKFORCE TYPE][WORKFORCE TYPE]
[WORKFORCE TYPE][WORKFORCE TYPE][INCOME][SOCIAL INTEREST][SOCIAL INTEREST][SOCIAL INTEREST][EDUCATION INTEREST][CONSUMPTION INTEREST][CONSUMPTION INTEREST][CONSUMPTION INTEREST][CONSUMPTION INTEREST][CONSUMPTION INTEREST][CONSUMPTION INTEREST]

And here is a sample list of an average labour worker within sub-par city with several Earth influences filling those boxes:
[A245JK89DJ102SK][2013/17/03:2146.46.12][-37.048601,143.739624][MARCHING IN PROTEST][C+R PTY WELDING LINES][O:9][W:4][F:5][S:4][B:8]
[ABUDL JHAKA][MIDDLE EAST][EURASIAN][OVERWEIGHT][UNFIT][PRIMARY][SECONDARY][MANUAL LABOUR][MANAGING][PROTESTER][TAXI DRIVER][31000][COFFEE][CINEMA][RESTURANT][JOURNALISM][FAST FOOD][MOVIES][BANANAS][WATCHES][DESIGNER BOOTS][RIFLES]



In this example the unique ID identifies the unit along with its game time date and time stamp. The games longitutude and latitude place and track where the unit is, current action notes what the unit is involved in and the employment variable shows where he is working.
Next we can see the unit has plenty of air (9/10), is a little thirsty, will soon need food, is not feeling secure which is probably why he is protesting, and is fitting in with the crowd with a beloning level of 8. Optional for the game is the tracing of the units bloodlines, showing mothers and fathers ancestry. Next are its health effects which reflect its consumption choices and how the unit may eventually being spending it money (ie on healthcare or medication) or how it will eventually die. Next we can see this unit has obtained general primary and secondary education types which influences the basic skill set it has and therefore determines the work it can obtain. Next we have the workforce types it is involved in which denotes where is has been working, where it is working, so its skills, along with any other non-social activity, in this case he is actively protesting. We can then see he is currently earning the equivalent of 31,000 a year in wages and prefers to socialise by having coffee and going to movies resturants. He also has an education preference for journalism which means if it was available and he could still meet his needs he would be studying journalism instead of what he is currently doing. He has 
a consumption preference for fast food, movies, bananas, watches, designer boots, and rifles; this means he will spend his disposable income on these goods in no particular order. Note that in this above exmaple, education types does not accurately depict a units education, education types will actually be made up of many types forming an education as chosen by the player. Primary, secondary, etc are just the games default educational system in place without any player management.
	

\StartSubSection{Generation of Unit Variables}
Of the above variables, some are automatically asigned to the unit, such as ID, name, date of birth, and location, however others are generated and changed throughout the game. Need variables are continually changing as influences from the environment affect the unit and may go from being low when the need has not been met for a while to being high when the unit has just had the need satisfied. In addition some variables may come and go completely, such consumption or social interests. 

Consumption interests are variables that determine where a unit will spend money and/or time when this is available, it also determines preferences when there are multiple options available. The consumption interest variable is specifically created for a unit when it is exposed to media (television or similar, public advertising, radio) or social interaction, both of which are generated when the unit does not have consumption interests already in place. 

Each unit has an amount of consumption interests determined by the age variable of the unit, this is to limit a toddler having a complete set of consumption variables too early. The variables are continually changing throughout the units lifespan and are based on the highest averages of external influences which cause consumption interest generation. So if a unit has 5 consumption variables at age 25, these will be the 5 variables that have had the most influence over the units environment. It's highest consumption varibale might be accredited organic food, the next might be "clown foods" inc, then shot guns, pure water, and finally fast cars. In addition as each in-game second (or minute or whatever period of time is deemed appropriate) passes the units average yearly ratios of influence are being calculated, with the variables changing places or being replaced completely. While it is possible that this could cause rapid change of variables, the gradual nature in which policy and building changes, an these changes in turn influece units should then be causing the influence ratios to be changing just as gradually.   


[Age][>>>][Amount of consumption interests] 




\StartSubSection{Needs and Actions Variance}
There are two broad categories which influences a units needs. The first is the units own actions, drinking water, eating food, getting a job, visiting a friend, all reduce the need to obtain these things, while excerising increases the need for food and water, and working long hours in isolation increases the need for belonging. The second is external or environmental effects such as propaganda, lack of clean air, food with poor nutitional content, or the action and views of other units. All these happenings have varying amounts of influence on the units needs which in turn effect the manner in which the unit acts to satisfy these needs.

In order to correctly map the degree of influence various needs have on a unit and the degree to which a units actions relieves or increased its needs, all of the variables effecting the needs of a unit are mapped with a point system quantively describing the degree to which needs effect the unit and the degree to which an action or event influences a units needs. Some of these will be wide reaching, effecting many units, such as chemicals in the cities water supply or television, other will be specific, such as health consequences from working shift work. In turn all these effect unit needs which cause unit actions, and may also create or change unit variables (mentioned above). 



Influence:
-Workforce Types
-Food Types
-Actions
-Information Types
-Economic 
-Drugs
-Social
-Ostracism



%SAMPLE  (Before) / (after)
%Input		Effects
%		security	food		belongingness				
%Crack Cocaine	+2 / -10	+10 / -5	+10 / -10			 
%
%Running	+1		-2		
%
%
%Enemployed	-3				-2*
%
%
%
%
%*depends on social system.





\StartSubSection{Workforce Consumption}
As discussed, in Avaneya a workforce unit is not just a unit producing goods and services, it is also a unit having any influence over the direction of the economy. This includes its patters of consumption. As each workforce unit is represented by an individual unit in game, moving and performing tasks, as this unit moves through game space it is consuming resources such as air, water, food, and minerals and like subtances through consumerable resources. What a unit can consume is based in part by what is available and what its needs and action influences are.



 



%DISCARDED NOTES:
There will be underlying conditions that will determine what a cities workforce is comprised of and players will have certain degress of influence over the direction the workforce takes but if certain conditions aren't met the citizenry may not accept or appreciate the current workforce type with the consequences of innaction by the player being varried. How players make these conditions and the direction they take with there workforce is a choice made in the sandbox style of play, it is a creation. They can create anything from human/animal labour pool to complete nanocracies with robotic workforces and humans indulging in deccandent pleasure palaces. The player can manage his city through inacting certain ordinance, directing funds, and other general city management processes including placement of civic buildings and diplomacy. 







>>>>>>> MERGE-SOURCE
