% This is part of the Avaneya Project Crew Handbook.
% Copyright (C) 2010, 2011, 2012
%   Kshatra Corp.
% See the file License for copying conditions.

% Goals section...
\StartSection{Goals}

If scenarios are molecules, goals might be the atoms they are made from. Let us take a look now at some of the goals a scenario can be constructed with, beginning with the exception. Bear in mind that this is only a subset of all the possibilities, since new goals can be defined by simply extending the scripts that drive the game engine.

\startitemize[4]
\setupwhitespace[big]
\head {\em No Goal:} 

A scenario may be goalless. In this case, the user erects their city and is responsible for its management without thought of a terminal condition for success, such as a time limit. But as with any other scenarios, the user can still strive to maximize their Genuine Progress Indicator (GPI). The GPI is described further in \in{section}[Terms & Concepts].

\head {\em Achieve Minimal GPI:}

The user must achieve a minimal GPI. This may or may not involve a time constraint.

\head {\em Natural Disasters Recovery:}

The user must rebuild their city in response to damage sustained from solar flares, micro meteorite strikes, land slides, marsquakes, or other natural disasters. Yes, Mars experiences all of the aforementioned.

    - Depleted aquifer -> collapsing city

\head {\em Protect Another City:}

The user must come to the assistance of another Arcadian city in repelling an attack of a foreign aggressor. The city has come under siege at the hands of Terran influences using the United Nations. Constraints may vary, such preventing the city from exceeding a threshold of damage, protecting a mineral extraction site, and so on.

\head {\em Protect Human Rights:}

Articles XI --- XXVI of the Rubicon Act described in \in{section}[Civil Disobedience & False Flags] enumerate the fundamental freedoms and human rights Arcadians enjoy with respect to their new state. The user is responsible for ensuring that they are upheld.

\head {\em Reduce Crime:}

The user must reduce crime by addressing factors, such as education, affordable living, health, and other fundamental human needs and environmental conditions that give rise to crime.

\head {\em Improve Transportation Network:}

The user's city is congested and requires that its transportation network is made more efficient.

\head {\em Improve Public Health:}



\head {\em Protect From Corporate Media:}

\head {\em Recover from GMO terminator gene}

\head {\em Prevent staged terror}
    - Red Unionists (backed contras)
    - NAU claims sabotage of mineral extraction sites as justification for troop deployment, though no sites exist.

\stopitemize
