% This is part of the Avaneya Project Crew Handbook.
% Copyright (C) 2010, 2011, 2012
%   Kshatra Corp.
% See the file License for copying conditions.

% Single & Multiplayer section...
\StartSection{Single & Multiplayer}

Let us start with an overview of the different modalities of play. These can be either single or multiplayer. Single player, as the name suggest, the user plays alone. Well, actually, with the computer and the challenges it presents. This includes a tutorial, campaign, and custom mode. The user can restore from a saved game in all three modes.

In tutorial mode, the user is guided through interaction with the elements of the game and its user interface by Khalid Zafar, one of the leading characters described on \at{page}[Khalid Zafar] of \in{chapter}[Leading Characters]. The aim is to train the user to the point where they are confident in taking on the campaign mode or multiplayer.

In campaign mode, the user assumes the role of Arda Baştürk, or any of the other leading characters, as they journey through the campaign which takes place near and following the end of the fictional timeline described in \in{section}[Aftermath & Diaspora].

The custom mode allows users to specifically select which scenarios they would like to play without thought of commitment to a full campaign. This is useful for those who would like to explore nonlinear gameplay.We will discuss scenarios further in \in{section}[Scenarios]. 

In multiplayer, the user can play over the local area network (LAN) with their friends in any of the available scenarios that allow multiplayer use. As in the case of the single player campaign, the user again has the option of loading from a previous saved game. See \in{section}[Networking] if you are curious what technologies will be used to provide this functionality.

Another multiplayer mode is {\it Solnet}\index{Solnet}. Solnet is the online service that allows users to join a much larger group for cooperative play over the internet. Users, having purchased a reasonable monthly subscription, gain access to the official Solnet server which hosts at least the Arcadia Planitia region of Mars following the fictional timeline's aftermath. 

But playing over Solnet offers another important advantage that makes it distinct from playing over the LAN. Consider the problem of what to do with the world the user plays in after they have signed off from a multiplayer session. This is not a problem for a game hosted on the LAN because it is expected that no one cares after you and your friends pack up and have had enough. That logical cluster of space, objects, states, and so on, is destroyed and no machine need have it persist and updated. But in the case of playing over Solnet, that world must remain. It is persistent because you are sharing the same world with potentially many others who would not appreciate it if it was abruptly destroyed simply because some of its users went to bed. Thus, the world hosted on Solnet carries on with a life of its own - even in your temporary absence.

The Solnet subscriptions are critical for ensuring the official server runs on reliable high quality hardware that can handle large concurrent loads made accessible through sometimes costly, albeit necessary, hosting. This is also vital for promoting additional development of the game's engine, artwork, music, and other graphical media that continue to make the game better and sustainable.

The following then is a summary of the different modes of play we have just discussed.

\startitemize[4]
    \item Single Player
        \startitemize[4]
        \item Tutorial
        \item Campaign
        \item Custom
        \stopitemize

    \item Multiplayer
        \startitemize[4]
        \item Local Area Network (LAN)
        \item Solnet
        \stopitemize
\stopitemize

That concludes a general overview of who can play with who and by what medium, but we must also discuss what it is that the user does either by themselves or with others. Let us now take a look at the concept of scenarios.

