% This is part of the Avaneya Project Crew Handbook.
% Copyright (C) 2010, 2011, 2012
%   Kshatra Corp.
% See the file License for copying conditions.

% Units & Controllers section...
\StartSection{Units & Controllers}

A unit is any vehicle, building, or other special object that needs to have its own unique properties and behaviours. It will always have a {\it controller}. This could be a human, either the user or someone else during a multiplayer session, or the computer's artificial intelligence.

It is the project's desire for an extensive unit tree. Although the user may direct the locations and activities of units under their control, not all units in the unit tree they will have access to -- nor will they necessarily even know about or want them. Some units, such as certain types of vehicles and buildings, only the computer can create when appropriate. We refer to these as non-playable units\index{Non-playable units} (NPUs)\index{NPUs}. At other times still, the user will be able to create certain units in specific scenarios that in others they would not.

As an example, you can and should build a greenhouse. But other influences in the game world might have the Clown Food\index{Clown Food} franchise erect a restaurant without having asked if the circumstance was right. Nevertheless, NPUs can usually be influenced by the decisions that you make, even if you are not directly in control of them. Nearly everything influences everything in the game.

But most units can and will change with time depending on various factors that take place in the game world. For example, an area zoned for residential usage may start with only a few pressurized yurts, but after time, it may develop into an extravagant, full fledged, pressurized biodome. With further time yet, it may fall into a state of disrepair with the underprivileged seeking shelter within its ruins - should the social, economic, and other conditions be met.

We should now take a look at some of the different categories that form the basic unit tree available to the user. This is a very modest subset of all units the user can possibly create, but does not illustrate the variety each unit can express. For buildings these categories are {\it Agricultural}; {\it Construction}; {\it Economic}; {\it Education and Research}; {\it Energy}; {\it Health}; {\it Recreational and Social}; {\it Justice}; {\it Militia}; {\it Production}; {\it Transportation}; and {\it Zones} for commercial or residential use. Because these are high level categories, they can sometimes be further subdivided. For example, {\it Health, Recreational, and Social}, contains a {\it Health} category, which, in turn, contains an Addiction Resource Centre unit. We will denote categories italicized and buildings non-italicized.

Beginning with the user's agricultural tree as shown below, we can see the basic units necessary that provide the city with food. Biophotoreactor plants produce kelp that is used as a fertilizer in the greenhouses. The greenhouses vary based on their pressure rating, as described in \in{table}[table:Pressure Ratings]. The mushroom ranches recover up to 70\% of organic plant waste into high quality sources of protein.
\placetable[here,left,hang][table:Agricultural]{}
{
    \bTABLE[split=repeat]
    \setupTABLE[row][1][align=yes]
    \setupTABLE[row][each][align=center]
    \bTABLEhead
    \bTR
      \bTH Agricultural \eTH
    \eTR
    \eTABLEhead
    \bTABLEbody
    \bTR
      \bTC Biophotoreactor Plant \eTC
    \eTR
    \bTR
      \bTC Greenhouse (68 mb) \eTC
    \eTR
    \bTR
      \bTC Greenhouse (340 mb) \eTC
    \eTR
    \bTR
      \bTC Mushroom Ranch \eTC
    \eTR
    \bTR
      \bTC Seed Store \eTC
    \eTR
\eTABLEbody
\eTABLE
}

\in{Table}[table:Agricultural] shows the user's basic {\it Economy and Commerce} tree. Everything appropriate that has to do with commerce, money supply, trade and exchange, we group into this category.

