% This is part of the Avaneya Project Crew Handbook.
% Copyright (C) 2010, 2011, 2012
%   Kshatra Corp.
% See the file License for copying conditions.

% How To Read This Book chapter...
\StartChapter{How To Read This Book}

Be honest. Large books are intimidating. Fortunately, at least in the case of this one, most people need not read it in its entirety. Having said that, project contributors certainly have nothing to lose in doing so. We tried to organize it in such a way that readers benefit the most when its chapters are explored in the order presented within the book.

Depending on the audience, different parts of this book will appeal more so to some than others. We describe who {\it you} might be in the context of a member of our open community in \in{chapter}[Specialities].

If you are a passive, casual, or non-technical reader, simply wanting the Coles Notes on the project, we encourage you to check out chapters \in[Leitmotifs], \in[The Game], \in[Leading Characters], and \in[Timeline] at a minimum. This should get you up to speed in as little time as possible.

If you already are, or would like to consider becoming a member of our community, you would probably do well to read most of the book. One noteworthy exception is \in{chapter}[Engineer Contributors], unless you would like to get involved in an engineering capacity, or you are simply curious of the types of problems the engineers will have to deal with.

One last thing. We do not recommend viewing this book with any of the non-free programs available from Adobe. Adobe Reader, and possible other Adobe programs, appears to contain a bug that results in this document not being rendered correctly.\cite[miklavec2012_broken_adobe] Unfortunately Adobe does not permit the community to repair these defects.\cite[extras={ See section 4.5.}][adobe_shitty_eula] Regardless, there are many {\it libre} viewers available that do not have these issues, such as any of the default viewers that ship with Ubuntu.

With all of that out of the way, on we go!

\StopChapter

