% This is part of the Avaneya Project Crew Handbook.
% Copyright (C) 2010, 2011, 2012
%   Kshatra Corp.
% See the file License for copying conditions.

% Leitmotifs chapter...
\StartChapter{Leitmotifs}

This is probably the most important chapter in the entire book in helping you to decide whether this is a project for you to be involved in or not. Avaneya has several recurring ideas or {\it leitmotifs}\index{leitmotif} within it that will help you determine whether it resonates with you.

We would like to keep the leitmotif set as minimal as reasonably possible, with everything else that we felt relevant, it hopefully already being possible to derive when obvious from that which is already present. However, if you still feel the set should be modified in some way, please bring it up on any of the project crew communication methods related in \in{section}[Communication]. We are open minded but, like us, you must also be prepared to reasonably substantiate your argument.

The following form the core leitmotif set the project felt were important, given in no particular order.

\startitemize[4]
\item
Most illegal narcotics on the street originate from government.\footnote{\href{http://afp.google.com/article/ALeqM5j6QonBKKMo2gw1e3ql-xUcQEZbVg}{Mexico Drug Plane Used for CIA rendition Flights. }{\it Associated Free Press}. 4 Sep. 2008.}\footnote{Watson, Paul J. \href{http://www.infowars.com/top-mexican-drug-lord-i-trafficked-cocaine-for-the-u-s-government/}{Top Mexican Drug Lord: I Trafficked Cocaine For The U.S. Government}. Infowars.com, 27 Apr. 2011. Web.}\footnote{Watson, Paul J. \href{http://www.infowars.com/ny-times-afghan-opium-kingpin-on-cia-payroll/}{NY Times: Afghan Opium Kingpin On CIA Payroll}. Prison Planet.com, 28 Oct. 2009. Web. 09 Apr. 2012.}\footnote{Attias, Bernardo. \href{http://www.csun.edu/~hfspc002/news/cia.drug.html}{The U.S. Government Is the Biggest Drug Dealer in the World!} 12 Oct. 1996. Web.}\footnote{\href{http://narcosphere.narconews.com/userfiles/70/Zambada.Notice.Rule12.3.pdf}{United States of America V. Jesus Vicente Zambada-Niebla}. United States of America Northern District of Illinois Eastern Division. 15 Mar. 2011. NarcoNews.com. Web.}\footnote{\href{http://www.youtube.com/watch?v=UT5MY3C86bk}{Former LA Police Officer Mike Ruppert Confronts CIA Director John Deutch on Drug Trafficking}. {\it YouTube}. 15 Nov. 1996.}

\item
People are never free. To be free is to be independent of influence, which no one is.\footnote{See {\it The Equality Trust} in \in{chapter}[Resources For Everyone].} We are all subject to influence, otherwise there would be no such thing as advertising. Everything is connected and nothing happens in a vacuum.

\item
All of the major axioms of neoclassical economics, the backbone of modern economic theory, are dangerously wrong. People do not always have rational preference, do not always seek self maximization, and do not always act on the basis of full and relevant information. Markets are not natural systems comparable to thermodynamics, do not always correct themselves, and their growth is not always good. The field is a pseudo-science.\footnote{See {\it Adbuster's} \href{http://anticap.wordpress.com/2010/10/25/jamming-neoclassical-economics/}{Jamming neoclassical economics} campaign, with prejudice for \href{http://www.adbusters.org/cultureshop/backissues/85}{Issue \#85}.}

\item
Government is among the leading causes of death.\footnote{Rummel, R. J. {\it Statistics of Democide: Genocide and Mass Murder since 1900 (Macht Und Gesellschaft)}. Münster: Lit Verlag, 1998. Print.}

\item
Most acts of terrorism are staged by government.\footnote{Griffin, David Ray. {Debunking 9/11 Debunking: An Answer to Popular Mechanics and Other Defenders of the Official Conspiracy Theory}. Northampton, MA: Olive Branch, 2007. Print.} \footnote{Warner, Kip. \href{http://www.thecanadiancharger.com/page.php?id=5}{An Open Letter to Minister of Public Safety Regarding 9/11}. Canadian Charger, 23 Aug. 2011.}

\item
Government is generally there to protect the interests of a few unelected principle benefactors, not society.\footnote{Doe, John. \href{http://www.teachpeace.com/Report_from_Iron_Mountain.pdf}{Report from Iron Mountain on the Possibility and Desirability of Peace}. Ed. Leonard C. Lewin. New York: Dial, 1967. Print.}

\item
Most health related deaths are rooted in diet.\footnote{Campbell, T. Colin, and Thomas M. Campbell. {The China Study}. First Paperback Edition ed. Dallas, TX: BenBella, 2005. Print.}\footnote{Canada. Statistics Canada. \href{http://www.statcan.gc.ca/tables-tableaux/sum-som/l01/cst01/hlth36a-eng.htm}{Leading Causes of Death, by Sex}. 2008. Web. 
According to {\it The China Study}, the top three causes are entirely dietary.}

\item
Agriculture is our principle means of interacting with the planet. Changing what we eat solves many global problems.\footnote{{\it Food Inc.} Dir. Robert Kenner. Perf. Eric Schlosser and Richard Lobb. Magnolia, 2008. DVD.}

\item
Usury, including fractional reserve banking, is slavery.\footnote{Griffin, G. Edward. {\it The Creature from Jekyll Island: A Second Look at the Federal Reserve}. 5th ed. Westlake Village, CA: American Media, 2010. Print.}

\item
Giving corporations the rights of human beings was catastrophic.\footnote{{\it The Corporation}. Dir. Jennifer Abbott and Mark Achbar. Perf. Mikela J. Mikael and Rob Beckwermert. Zeitgeist Films, 2004. DVD.}

\item
Very little of what we call human nature is actually genetic. Genes specify human needs, not behaviour.\footnote{See \in{section}[Socioeconomic Simulation].}
\stopitemize

If you have gotten thus far, and you are still comfortable being involved in this project, then you will probably find it rewarding. Otherwise, there is no sense in being here. There is no requirement that you believe anything or share our views. No one is forcing you to do anything. This is not a project for everyone and there are countless other software {\it libre} projects that could use the talents of those with reservations.

But regardless of whatever our readers choose to do, creativity is required whenever we present the aforementioned. We always try to encourage the users to do some thinking of their own. As Socrates once said, {\it I cannot teach anyone anything. I can only make them think}.

\StopChapter

