% This is part of the Avaneya Project Crew Handbook.
% Copyright (C) 2010-2015 Cartesian Theatre <info@cartesiantheatre.com>.
% See the file Copying for details on copying conditions.

% Game Mechanics chapter...


\StartSection{Companies}
In Avaneya there are really two kinds of units that contribute to the macro economy, the workforce unit who is like a person, and the company. The company acts like a type of workforce unit to simplyfy a multitude of things and to calculate when to be created and satiate demand for certain goods and create demand for other goods. 


\StartSubSection{Creating Companies}
Referring to seciton [...] on demand curves, when workforce units are consuming something in particular, be it from a category like food (or one if its subcategories like vegetables), this creates a demand and things will get purchased at a certain price. Also as mentioned in section [...] in some cities this price will be predetermined by the player in other cities this price will be determine by the market, or what workforce units are willing to spend on something (refer to section [...micro_economics.tex] for more on workforce units role in the economy).

Companies are created naturally - by the game - when the price of producing each unit of the specific good they will produce is less than what they will sell each unit to the population for.
%Comapnies Calculating goods, costs and purchasing means of production.   
-Based on what it costs to purchase and resell a good/service the game will calculate this vs what a new company will make for selling it. If the purchase cost is less than breaking even cost of the selling the goods, a company has the potenital to be made given the appropriate access to the goods/services as well as workforce units to fill the appropriate labor roles at various levels.

-Different goods require differnt inputs to recieve said goods as an output. The production of an apple is very different to the production of a car but both require input of various qualities and quantites to action output. In Avaneya, some goods like the apple have a simple level of required inputs. Apples require a seed/tree, which requires time to grow, nutrients (such as soil and water, or hydroponics), and finally a unit to pick the apple. These three factors add up to create a final cost of producing the good to a sellable point. In contrast a car requires raw materials and metals, a production line of various parts with a workforce appropriately skilled, then a final production line to put the car together. 

When calculating whether to enter production of a goods or service, these factors of the supply chain and calculated to determine the feasibility of an enterprise.

In some instances, a company can have control over the entire production process, in others a final company will need to purchase goods from other companies to bring a good to the final sellable point. As in the example above, it is relatively easy for a farmer to have complete control over all inputs in his apple orchard but a car manufacturer will most certainly require to purchase parts from other companies and then assemble to a final product.

It can be seen that supply of all inputs for output may become a problem. The game calculates how much access a company has to the required goods to justify starting production in the current situation. If a firm has access to enough parts to make a single car, it will not purchase an entire factory for this. Therefore the game also calculates the costs of the firm purchasing land and buildings and hiring a workforce into this decision.

If: Sellable price  > Cost of parts + Cost of production + Cost of land and buildings
A firm will attempt to start production.        


Firms will attempt to purchase various goods from other firms, transferring the appropriate amount at the market level. The goods will then be transported to the purchaser on the appropriate route. If transport is haulted due to inadequate facilities no purchase will take place, and if the firm has not yet established, no calculation to determine its creation (as stated above) need take place.





%Calculating access to goods and zoning.
Companies also require adequate zoning, both sizes and locations of zoning affects the ability of a company to function 







Companies can also be created by the player .




%Creation

%Change


%Conglomorates, Multinationals.

%Companies influencing quality of product.
%Employess influencing company quality.
%Companies hiring employees.

%Companies expanding and diversifying.

%Companies and stock and bond options.
