% This is part of the Avaneya Project Crew Handbook.
% Copyright (C) 2010-2014 Cartesian Theatre <info@cartesiantheatre.com>.
% See the file Copying for details on copying conditions.

% Game Mechanics chapter...
%\StartSection{Workforce}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Section Contents:
% 
%  Introduction 
% 
%  Visual Representation of the Workforce
% 
%  Workforce Unit Creation
%
%  Workforce Movement
% 
%  Needs, Interests, Types, and Variables
%
%  Workforce Unit Needs
% 
%  Workforce Interests
% 
%  Workforce Interest Creation
% 
%  Workforce Types
% 
%  Transitioning Between Workforce Types
% 
%  Workforce Variables
% 
%  Influencing Workforce
% 
%  Workforce Consumption
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%5


%-----------> WORKFORCE <---------------------------------------------
\StartSection{Workforce}

What a city can produce, how much it can produce, and what it is capable of producing are largely byproducts of its workforce, that is the human population interacting, with eachother and technology, to create goods and services. Workforce also determines things such as consumption patterns and criminal behaviour, it is humanness. This section discusses and describe what are the immense variety of workforces, what they represent, how they are created, maintained, and influenced, and what they are capable of. 

A workforce unit is a single entity of the human population which influences economic input and output. This unit exists at a certain level of education, training (or lack there of), life experience and may increase in complexity over its lifespan. The overall workforce is all productivity and idleness factors in measureing the economy, extending to consumers (and even the homeless, criminals, and the unemployed). The workforce of a city is all units in the city capable of exerting influence on the direction in which resources or money flow. So the workforce is the primary means by which labor and skills are utilized and from which productivity is tracked.

Certain workforce units will have vastly greater levels of complexity and therefore overhead than others. For instance a highly educated member of government as a lot more complexity than a basic line worker at a basic factory. The member of government will have gone through multiple levels of education and various types of employment to reach its current state, in contrast the factory worker only requires rudimentary education and could walk into his job without skills. While this is a simple example it serves a point to illustrate that some of the workforce will have more variables than others and be more difficult to create or maintain.

A labour workforce requires certain factors to bring it into fruition which may include a certain level of education, access to work, and even restrictions to certain aspects of these or other factors. An intellectual workforce requires different factors to develop than a manual labour workforce. While the two are not mutually exclusive certain combinations and mixes of labour will be more difficult to create in a single city. In adition, otherwise different workforces may share aspects of influence over the economy, known as workforce types. For instance a workforce unit in manufacturing may have some of the same education level as a workforce unit in commerce. These two units share a couple of "workforce types" (being qualative education) but have been influenced by other means to different outcomes. 

The workforce performs not only the tasks required by the society, but also tasks or actions that may detriment the society. The aim is to keep the style of play open by not limiting or directing players down paths of right and wrong but rather show an outcome of workforce actions and then challenge the player to deal with this in constructive even creative ways; casting influences over the population and controlling the creation of the workforces variables. 

A cities workforce is influenced through players building, inacting ordinance, and various other city policies and diplomacy and raised through education, attracted via work opportunities from other cities, and hired directly. It is also influenced by numerous external and internal agencies such as media, education, drugs, and fashionable consumption. Through this process the workforce becomes increasingly convoluted and multifaceted, and the style of play implemented by the player, other players, and influences from NPCs, will help to determine the quality and quanity of a city's workforce. 

%--------------> VISUAL REPRESENTATION OF THE WORKFORCE <--------------------------
\StartSubSection{Visual Representation of the Workforce}
A cities multi faceted workforce is actually made up of varying amounts (many thousands) of individual humans and are depicted visually as people, each an individual person, performing tasks or moving around the city.  

Visually, the workforce units need to be represented as humans of varying skin tones, ages, sizes, and wearing various clothing and apparel. Depending on the zoomable scale all these units must be able to be seen performing actions and influencing there surroundings and interacting with other units as well as their environment and objects in this. 

The depiction of skin tone will be generated from the units parents (see subsection workforce unit creation for more detail on unit creation). 
%Move to unit creation section?
Each unit will have a skin tone variable from 0 to 6 with 0 being the whitest, 6 being the blackest, and anything in between being degrees between these two. 

At unit creation the skin tone value will be calculated from the values of the units parents, creating a random value for the newly created unit within limits set by the units parents skin tone value. Refer to table ... 
%Table showing skin tone examples.
   M 
F  0 1 2 3 4 5 6
0  0 1 2 3 4 5 - 
1  1 2 3 4 5 6 6
2  2 3 4 5 6 6 6
3  3 4 5 6 6 6 6
4  4 5 6 6 6 6 6
5  5 6 6 6 6 6 6
6  - 6 6 6 6 6 6 

This table shows for each parent the skin tone value that each newly created unit will fall between at random with it being increasly more common for a darker value to be predominant as the skin tone value increases. This needs to be reflected in the random generation, the increased proability of darker tones in new offspring in darker couples.

The only expcetion to this is where one parent has a skin tone value 0 and the other a skin tone value 6. In this case the unit will default to a value of 3.

While this method does not completely represent polygenic inheritance, its works well as a model for the game.  
%(source 1 - Biology, Campbell Reece Meyers - pg 263)
%(source 2 - http://multiracial.com/site/index.php/2000/10/01/how-do-we-inherit-our-skin-color/)


Initially the depiction of clothing will be based how characters are shown in the concept art. Eventually clothing will be culturally influenced and the clothes units are depcited in will depend on unit variables and its access to certain clothes. Units with certain interests in fashion may only purchase certain types of clothing, as will units exposed to certain clothing advertisements. The depiction of units in clothes needs to be done in a manner that allows them to change over time, for weather, times of day, trends, and culture.  

The depiction of size is less important. Units of will vary in size noticeably between age groups, though some age groups will show no variation in size such as the various stages from young adult to late middle age where little change will occur (refer to section ... on age variables for further). 

The depcition of age will be based entirely around which age bracket the unit falls within and this will change as the unit moves through various age brackets similar to unit size(refer to section ... on age variables for further). Some age bracket changes will be drastic others will be minor changes.

In addition a large variety of animations will be required for the number of potential interactions with an aim of avoiding repetitiveness. These animations will include things such as walking, running, eating, talking, picking things up, etc. Refer to table ... for a complete list of all actions required. This table takes into consideration all events that a unit may end up doing, from basic everyday tasks like thos previously mentioned to specific tasks for specific jobs. 
%-> SPREADSHEET -> animations

%--------------> WORKFORCE UNIT CREATION <----------------------------------------
\StartSubSection{Workforce Unit Creation}
Workforce units are created in two ways, the first when a two units interact socially, the second is through technology such as invetro fertilization. In the first instance, where new units are created as a result of the love between a mommy workforce unit and a daddy workforce unit, the social situation will allow for enough privacy so that the actual act is not depicted. After this event there is a probability of the unit falling pregnant depending on both the unit's age brackets, refer to table .. below on pregnancy probability for further. 

  M			        Key:
F 1  2  3  4  5  6  7  8  9     1 - new born
1 -  -  -  -  -  -  -  -  -     2 - toddler
2 -  -  -  -  -  -  -  -  -     3 - child
3 -  -  -  -  -  -  -  -  -     4 - adolescent
4 -  -  - 10  -  -  -  -  -     5 - young adult
5 -  -  -  -  9  8  5  3  -     6 - middle aged
6 -  -  -  -  8  6  4  1  -     7 - late middle aged
7 -  -  -  -  4  3  2  1  -     8 - elderly 
8 -  -  -  -  1  1  1  1  -     9 - elderly infirmed
9 -  -  -  -  -  -  -  -  -     

Table depicting probability of falling pregnant depending on age
Note: Probability out of 10 (0 to 10)

%Units created through sex
Workforce unit creation falls within the belonging need of the workforce unit but may also fulfil an interest. In determining if unit creation process is implemented two workforce units either want sex to fullful an interest, or want sex to reach higher tier belonging need of having own family (or one of each); and are at or will access a private location.

There are designated areas in which workforce unit have sex. These will be obvious places such as a bedroom but also places such as private toilets, and other secluded locations. Such secluded locations will be generated through various means as a city is created. When units are at a stage where they want to have sex they will go to a private location. Two examples: two units in a house with other units in it will move to a bedroom; units in an office may move to a restroom. 
% -> SPREADSHEET -> Locations of "private settings"

If access is gained then the probability of unit creation is referred to table ... (above).

Sometimes units will be made through reaching higher tier belonging needs, sometimes they will be created through units fulfilling interests. This creates situations in which unit creation will be planned and unplanned, will be created in a variety of social classes as well as help create social classes.

%How frequently sex/baby making happens? What determines frequency?
Frequency will be dependant on unit variables that factor in things such as income and intelligence.
Generally a unit will have social and belonging factors met then income factors met also. 
Units wanting to keep living standard is one way to slow reproduction.

%Abandonned Units and Adoption


%Units created through other artificial means
Units who reach the belonging need tier in which they create their own family but cannot or do not will try either adoption (see above) or technology such as invetro fertilization. 

Through the probability methods listed in table ... above, there will be some units who will not end up with children, especially at older ages. Provided these units still try to meet their beloning needs they will eventually (define eventually) attempt artificial means of child creation. This is provided they have access to this, be it in their city or another's, and they can afford this method.

Through in vetro the female workforce units age is the sole factor after a successful conception. Table ... shows the probabilty of success for in ventro.

			        Key:
F 1  2  3  4  5  6  7  8  9     1 - new born
1 -  -  -  -  5  4  3  -  -     2 - toddler
				3 - child
				4 - adolescent
				5 - young adult
				6 - middle aged
				7 - late middle aged
				8 - elderly 
				9 - elderly infirmed

Table depicting probability of successful in vetro depending on female unit's age
Note: Probability out of 10 (0 to 10

Table ... above dipicts a base probability rate. Other factors may highten or lessen these probabilities such as technological complexity and the operator units expertise (level in appropriate workforce type).

In vitro will be attempted by units, provided they still seak to fulfil the beloning need, as frequently as is allowed with what is games equivalent of a 6 month refractory period between attempts.

Same sex units in vetro who use a sperm donor, the unit within the couple that chooses to be the genetic parent is determined at random on a 50/50 split.


%Unit variables created at unit creation and what influence same.
While most unit variables are determined from the environmental experiences of the unit, some will be determined at the creation of the workforce unit, and other will be created from experiences during pregnancy.

Skin color, as mentioned above in the section of the visual representation of the workforce, will be determined by factors from the the parent units.

Sexuality is another unit variable that is determined at birth. Each unit has a sexuality variable that represents what would be the genetic component of a unit's sexuality, ultimately determining what gender of unit it will date and have a family with.

The level of this variable is determined by the level present in both parent units and interacts with environmental influences (see section ... influences) at (... age ranges etc...?) to determine the degree to which a unit is homosexual on a spectrum that includes bisexuality. 

Table ... below indicates that degree of genes passed on to new workforce units from the male and female workforce units.

   M			        
F  0  1  2  3  4  5  6  7  8  9     
0  0  1  1  1  2  2  3  3  4  4
1  1  1  1  2  2  3  3  4  4  4     
2  1  1  1  1  3  4  4  5  5  6     
3  1  1  2  2  2  3  3  5  5  6     
4  2  2  3  3  4  4  5  6  6  7     
5  3  3  4  4  5  6  6  7  7  8     
6  3  3  4  4  5  6  6  7  7  8     
7  4  4  5  5  6  7  7  8  8  9      
8  4  4  5  5  6  7  7  8  8  9       
9  5  5  6  6  7  8  8  9  9  9

Table depicting degree of which units sexuality variable compounds with homosexual environmental influences to determine sexuality.
Note: scale ranges from 0 to 10.


This table (...) represents the degree to which an environmental influence that determines homosexuality will be compounded by. People with a higher level of genetic homosexuality will have this influence increased at a quicker rate, filling the scale to homosexuality quicker, therefore be more likely to be gay.

Units will keep the same genetic sexuality variable on the scale from 0 - 9, this variable does not change as the unit is exposed to more or less gay influences. The environmental influences determine where the unit falls within a straight, bisexual, or homosexual [STR/BI/GAY] spectrum.



%units being born with probability of being gay. 
%set probabilty higher and rely on environment to regulate.
%Units gayness from mum and dad.
%List of environmental influences that help form homosexuality.




Other unit variables may be created from experiences during pregnancy
%Effects of pregnancy (FAS etc)


%(source 1 - Life-Span Human Development, Sigelman & Rider - pp 69 - 70)



%--------------> WORKFORCE MOVEMENT <----------------------------------------
\StartSubSection{Workforce Movement}
Workforce is a dynamic "resource" and can move city wide (if allowed) and globally (if they are not trapped in the city and have the appropriate travel documents, etc). If your workforce requires cheap housing and this becomes easier to gain in another city, many may migrate to this city seeking better opportunities given that chance. Likewise they can move within your city to different opportunities and movement both within a city and between cities can be controlled to an extent to enable better or worse access to certain locations, increasing and decreasing travel time.

Indeed a lot of a workforce units interaction and influence over the cities economy and general development will have to do with its ability to move. How far a unit can travel while meeting all its needs (discussed below) will determine this influence. Depending on the unit type it may be walking to work littering, going out to lunch, moving between two locations at work (such as goods transport). It may be in cohesion with other units performing some of these tasks, for example eating lunch with a group of friends, building a vehicle, baking bread, or performing theater or dance. 

The means in which unit movement is performed is based on the units needs and the environs which allow it to meet certain needs. Movement can also be restricted through placing private or off limit areas, constructing walk ways, allocating cetain areas and directions, making some directions longer and others shorter, charging to travel some routes, and creating and limiting private and public transport options. 

%Expand.

%--------------> NEEDS, INTERESTS, TYPES AND VARIABLES <---------------------
\StartSubSection{Needs, Interests, Types and Variables} %Variables vs Needs vs Interests vs Types



%--------------> WORKFORCE UNITS NEEDS <-------------------------------------
\StartSubSection{Workforce Units Needs}
As mentioned briefly above, a cities workforce is made up of many individual units. Each of these units can be a part of one or more workforce types. These workforce types require certain conditions or needs met in order to be created, maintained, and have degrees of functionality. 

The needs a unit has determine where it will go, what it will do, and how it will do it. If we take section [Maslow's] accuracy into account we can view humans in a somewhat tabula rasa state, with no ill will except to maintain there own needs. Further to this, it can be seen that the nature in which humans interact and the means in which they satisfy their needs is frequently determined by something outside of there biological state. 

The human needs consist of oxygen, water, food, security, and belonging.



Oxygen: breathable air, a human can survive for less than 3 minutes without suffering potentially serious consequences. 
%Expand and example.

Water: clean drinking water. The human body can survive upto a week without water (provided good conditions). 
%Expand and example.

Food: sustainence and nutrtion. It is essential to not only have something to eat but that it contains the various essential nutrients to support human life. A human can last several weeks without food provided they have water.
%Expand and example.

Security: shelter/housing, feeling safe in surroundings, and feeling secure about the future.
%Expand and example.

Belonging: feeling of making a contribution, having people to talk to, being a part of something more than itself, sex/workforce unit creation.
%Belonging- family; friends; partner; own family; nation/state. (hierarchy or random?)
%Expand and example.



In light of the above paragraphs, it can be seen that it is relatively simple to satisfy a humans security needs. However, continually bombard it with messages of insecurity and threats despite evidence to the contrary and it will seek ever greater means to satisfy this need, perhaps waging unnecessary wars against otherwise peaceful peoples. This is just an example of how a player can create or solve disonance in human needs and scenarios the game can implement to challenge players.










%--------------> WORKFORCE INTERESTS <-------------------------------------------------
\StartSubSection{Workforce Interests}%Interests interacting with needs.
Each human workforce unit has a variety of interests which dictates to the simulation its consumption pattern, or how the unit will react in certain situations. This will determine the direction it will take in its spare time as well as the way in which it satisfies its needs. 

The majority of workforce unit changes occur at the unit level with each individual unit deciding which direction it will take in furthering (or lessening) its development. The primary way this is undertaking is the unit seeking to meet its units which are its primary goal. Some of its interests are redimentary and based around being more complex way of meeting its needs, such as eating at fine resturants with friends, other interests will be education or career specific. 

Interests are based around the units needs. There are interests for food, shelter, belonging, and less so for water and oxygen.

Units will have a set number of open slots that can acquire interests based on: (1) Their environmental exposure (friends, family, community, etc); (2) Previous variables (education, other interests, age).

%Example: having certain worforce types will open up influences/interests. In this light not all external influences will influence the interests of all workforce units. 

As the unit goes through age brackets which change it, these slots will become open and then filled with new or old interests. This may occur either immediately and gradually depending on the strength/frequency of the signal that the unit is receiving. In some cases this will grow over time as the unit engages in the influence. The stronger the strength of the interest the harder it is to be replaced.

Once gained interests occupying a slot in the unit will be given a strength number, level, or similar indicator to determine how involved, addicted, interested, needy, etc, the unit is of that interest. 

Various influences will determine this strength. Length of exposure, exposure by family while growing up, exposure by peers, acceptance in society, are examples of stronger influences that create an interest quicker. In addition the interest itself may have a degree of strength imposing itself onto the unit, such as an addictive drug. Different influences affect their allocated interest at different rates. For example the food need influences of junk food (food high in fats and sugars) increase at a rate quicker than interests of healthy food. Influences created with social groups also increase quicker, such as sharing meals with family, or seeing movies with friends.


Replacing an interest occurs through exposure to other interests. Lower level, seasonal, or trend based interest's will be replaced relatively easily compared to interests which are higher order. This represents interests that are ingrained both individually and socially. 
%Expand.

In the simulation this represents changes in trends, the availability of items, and with the units internal interests, such as education, the units ability to "self examine" and view needs to change. With this in mind there will be a link between many interests or interests will be in categories, allowing units to jump from interest to interest as part of its modeled "internal dialogue".
%Expand and explain.

To aid in interest generation through the ages, units will have slots of the next age open in a hidden manner, the slots are fed information to help determine what interests will fill those slots when they become open. In this way we can model the influence of domestic violence, household drug use, etc, as well as positive influence's such as parents who play instruments, exercise, etc, and even neutral influences, eg food consumption, clothes shopping.

Some interests will only be available or last during certain age groups. A child who liked "Random Childhood Show" and moves up into an age bracket which the show doesn't represent will have the interest slot that this show represents re-opened. 

Figure [Workforce_Interests.gv] displays the method in which units gain and replace interests. From the graph we can see that there are specific external variables which cause the unit to generate interests. Primarily there are various media forms as well as other workforce units performing actions. Media will largely be in the form of advertisements or programs with "content". Units, who perform behaviours will be "demonstrating" behaviours that influence other units. All behaviours, media, and externalities based around interest generation will have an interest category, a broad interest variable such as "Consumption Interest" or "Social Interest". Within each of these categories are content types which is the specific interest that is being displayed or is to be generated for the perceiving unit. If the interests category is full then the unit can still generate an interest if this new interest is able to replace an old one. If the interests category is not full a new interest will be generated with the unit under that category (e.g Consumption Interest > Rap Songs) provided the unit has perceived the externality for the required duration for that interest.
%Example: influences -> interests -> needs -> unit.
%Example: social influence "A" -> social interest "1" -> belonging need -> unit.

Interest Externalities and Perception Times
%Expand.

%->SPREADSHEET -> List of interest by category with times required to be viewed.


As a unit acts, moves, etc, it will have a record of all influences on it. If a unit has an interest slot open each time it comes across an influence associated with that same interest, than the influence is recorded and summed over time until the interest is finally generated or filled by another influence (if it is easier for processing speed these lists can be terminated after a certain amount of time of no increase in influence).

Each unit has an amount of consumption interests determined by the age variable of the unit, this is to limit a toddler having a complete set of consumption interests too early. The interests are continually changing throughout the units lifespan and are based on the highest averages of external influences which cause consumption interest generation. A units strongest consumption interest might be accredited organic food, the next might be "clown foods" inc, then shot guns, pure water, and finally fast cars. In addition as each in-game second passes the units average yearly ratios of influence are being calculated, with the variables changing places or being replaced completely.   
 

%--------------> WORKFORCE INTEREST CREATION <---------------------------------------
\StartSubSection{Workforce Interest Creation}
Figure [... (picture and flowchart of unit interacting with environmeny at various stages, underneath is an example representation of influences impacting the unit and how the simulation is tracking them)]
%Expand and explain.

%Example 1
A unit will have a variety of influences impacting it. As it walks down the street it is impacted by a billboard advertisement, a buscar, and a group of people eating (having a good time).

%Example 2
The unit arrives at work and is influenced by a group of people drinking coffee, a person eating, and a radio advertisement. Notice that different level of influence the lone person eating has in this image compared to that in the previous image of a group of people eating and enjoying themselves.

%Example 3
Next the unit engages in its employed office based work. In this example the unit has no influences impacting it while working. This is not universal but in this case there are no influences.



Units with interests in areas impact other units.
A unit with an interest will act on that interest therefore influencing other units.
%Expand and explain.
%Example.



%--------------> WORKFORCE TYPES <------------------------------------------------
\StartSubsection{Workforce Types}
The dynamic and multifaceted nature of workforce may become more evident when reading about the various workforce types. From the list below it can be seen that workforce comprises what the population is doing in reaction to the cities governance or that of other cities and also what stage of development a unit is at. Some types of workforce are easy to create and maintain in a state desired by players. Others are more specific and complex requiring specific circumstances surrounding a population such as the right levels of affluence and education. In addition workforce types applied to a workforce unit can overlap. A unit could be tertiary educated but unemployed, or a physical labourer with a secondary education. 

Units may move between workforce types, as in moving through education levels or going from unemployed to employed. Some factors which change workforce types are done through time, others done through city management.


% -> SPREADSHEET -> Workforce Types 
-Homeless
-Unemployed
-Physical labour (varying levels/classes, ie production, mining)
	(a physical labour type for each task needed in a city) 
-Mental labour (various)(as above)
-Education (primary, secondary, tertiary, post tertiary [not complete, see education types])
-Employment skill (varying types depending on field, ie years of experience)
-Training (varying types, ie separate training that a business or government may provide)
-Criminal (varying types)
-Militia
-Consumer (varying types depending on influences, ie social eco status, cultural)
-Mechanical (varying)(only applies to machines)

-Protester (various depending on cause)
-Animal (varying)(only applies to animals)
-Age (varying)(only one, changes)
-Drug user (various)
-Drug addicted (various)
-Rioter (various depending on cause)



%--------------> TRANSITIONING BETWEEN WORKFORCE TYPES <---------------------------
\StartSubSection{Transitioning Between Workforce Types}
Unit gets skill and either upgrades workforce type or adds a workforce type.

Workforce units can have multiple workforce types noting the variety of abilities they have. The methods in which the units gain these varies from default/automatic changes (such as units changing types as they move through different age brackets giving them certain effects, abilities, needs, and wants that is particular to that age bracket and to a particular cities style) through to workforce types specifically chosen by the unit such as further education or changing jobs. 

Figure [Workforce_Types_Changes.gv] dicpicts the change of a workforce unit over time. We can see in this figure that certain aspects of the unit remain at certain stages of its lifespan. In the first period of "time lapse" in the example the unit moves from secondary education into tertiary education while still working in a factory (as a line worker). During this period the unit has aged (age group 2 has changed to 3), gained experience in line working (Line Worker 1 has changed to 2, that is moving from level 1 to 2) and has completed 2 units of tertiary education giving it the extra workforce types Economics 1 and Literature 1. We can also see that during the first time lapse the unit has not lost his secondary education workforce type even though he is no longer undertaking this education type as this represents skills and knowledge the unit has gained. 

More time has now passed and the unit will have aged but not so much so as to have moved into a new age group. Therefor it has retained age group 4. The unit has also studied further at a tertiary level gaining economics 4 and literature 2. Note that these have replaced economics 4 and literature 2 as they represent levels or improved skill in these areas, while they appear to replace the previous economics 1 and literate 1, they are upgrades of these with economics 4 having all that was in economics 1 and more (the same goes for literature 2). This is similar to how the unit gained further experience as a line worker in the previous example. The unit has also started work as an economist. The "Economist 1" represents that the unit is level 1 at being a working economist. Also note that although the unit is currently working as an economist at level 1, he has retained the line worker level 2 workforce type as this is a skill the unit will retain, similar to his secondary education, as well as his education of economics 4 and literature 2. Given more time this unit may gain levels in Economist and will also change age groups. 

From this example we can see that workforce units are gaining experience over time in a variety of areas depending on what they are doing over time and that each type of experience is categorised as a workforce type which can either change, improve, or increase in number over time. 

A workforce type does not note when the unit is currently working only what skills he/she has. The unit variables (see[subsection]), discussed later, contains the information of where the unit is currently working, among other things.

How units get skills - interests, needs, continual work, education, age, handicaps.
%Expand and explain.
%Example.

How units move between jobs?
%Expand and explain.
%Example.

How units choose an education type?
%Expand and explain.
%Example.

%--------------> WORKFORCE VARIABLES <---------------------------------
\StartSubSection{Workforce Variables}

Age:
Workforce units change age types as the unit ages, being the changes in the length of time since it the unit was created. 

As mentioned above, workforce units will change one of there types as the unit ages. The subsection on unit variables notes that all units have a time stamp noting there creation, from this time stamp the game will calculate what age this unit is and therefore what age bracket it falls within. The age bracket is just another workforce type which allows the game to give attributes or apply conditions to the unit. New borns have different needs and will effect the world differently than middle aged units which in turn will be different from the elderly. 

Specifically there will be 9 age groups being: the age groups will be new born, toddler, child, adolescent, young adult, middle aged, late middle aged, elderly, elderly infirmed. A unit may not reach all of these age groups but those it does reach it must pass through sequentially.


Calamity:
Units will come across certain situations which will cause them injury and/or disabilty which could be temporary or permanent. This change in workforce is not specifically chosen but an outcome of unit performing certain actions or encountering certain things. These units will automatically be given this type which will be listed on its variable.  


%Variables which uniquely identify the unit and determine its capabilites.

Unique ID
[Unique alpha-numeric key](unique to each unit)
[Date and time stamp of birth](auto generates secondary age variable)[>>][Age]
-These above two generate a unique unit, allowing for keys to be recycled, but date/time stamp will change.

%[First Name and Last Name](may become unique to a cities cultural influences)
[Race/Ancestry]
[Income](fluctuates)
[Workforce Types](multiple, fluctuates)
%[Health Effects](multiple)
[Social Category](?)

%[Political](?)maybe just have complex interaction of existing interests with politcal ideals.

%SAMPLE UNIT VARIABLES
Here is a sample list of a units variables:
[UNIQUE ID][DATE/TIME O.B][LATITUDE:LONGITUDE]
[CURRENT ACTION][EMPLOYMENT][OXYGEN][WATER][FOOD][SECURITY][BELONGING]
[NAME][FATHERS RACE][MOTHERS RACE][SEXUALITY][STR/BI/GAY][HEALTH][HEALTH][EDUCATION TYPE][EDUCATION TYPE][WORKFORCE TYPE][WORKFORCE TYPE]
[WORKFORCE TYPE][WORKFORCE TYPE][WEALTH][SOCIAL INTEREST][SOCIAL INTEREST][SOCIAL INTEREST][WORK INTEREST][CONSUMPTION INTEREST][CONSUMPTION INTEREST][CONSUMPTION INTEREST][CONSUMPTION INTEREST][CONSUMPTION INTEREST][CONSUMPTION INTEREST]

And here is a sample list of an average labour worker within sub-par city with several Earth influences filling those boxes:
[A245JK89DJ102SK][2013/17/03:2146.46.12][-37.048601,143.739624][MARCHING IN PROTEST][C&R PTY WELDING LINES][O:9][W:4][F:5][S:4][B:8]
[ABUDL JHAKA][MIDDLE EAST][EURASIAN][OVERWEIGHT][UNFIT][PRIMARY][SECONDARY][MANUAL LABOUR][MANAGING][PROTESTER][TAXI DRIVER][31000][COFFEE][CINEMA][RESTURANT][JOURNALISM][FAST FOOD][MOVIES][BANANAS][WATCHES][DESIGNER BOOTS][RIFLES]



In this example the unique ID identifies the unit along with its game time date and time stamp. The games longitutude and latitude place and track where the unit is, current action notes what the unit is involved in and the employment variable shows where he is working.
Next we can see the unit has plenty of air (9/10), is a little thirsty, will soon need food, is not feeling secure which is probably why he is protesting, and is fitting in with the crowd with a beloning level of 8. Optional for the game is the tracing of the units bloodlines, showing mothers and fathers ancestry. Next are its health effects which reflect its consumption choices and how the unit may eventually being spending it money (ie on healthcare or medication) or how it will eventually die. Next we can see this unit has obtained general primary and secondary education types which influences the basic skill set it has and therefore determines the work it can obtain. Next we have the workforce types it is involved in which denotes where is has been working, where it is working, so its skills, along with any other non-social activity, in this case he is actively protesting. We can then see he is currently earning the equivalent of 31,000 a year in wages and prefers to socialise by having coffee and going to movies resturants. He also has an education preference for journalism which means if it was available and he could still meet his needs he would be studying journalism instead of what he is currently doing. He has 
a consumption preference for fast food, movies, bananas, watches, designer boots, and rifles; this means he will spend his disposable income on these goods in no particular order. Note that in this above exmaple, education types does not accurately depict a units education, education types will actually be made up of many types forming an education as chosen by the player. Primary, secondary, etc are just the games default educational system in place without any player management.

Of the above variables, some are automatically asigned to the unit, such as ID, name, date of birth, and location, however others are generated and changed throughout the game. 

%Expand and explain.
%Examples.

%--------------> INFLUENCING WORKFORCE <-------------------------------------
\StartSubSection{Influencing Workforce}
How people are raised, educated, and conditioned through experience largely determines their consumer patterns and may be reflected in a population's consumer choices. In game, workforce is influenced through players building, inacting ordinance, and other various city policies along with inter-city diplomacy. The method in which various workforce units are shaped and changed through the development of a players city are complex and discussed in depth below.

The main method in which a player can influence his own or someone elses workforce is through economic means such as taxation or ordinance (though not limited to financial insentives but extended to include incentives derrived for economic direction), although other methods may exist such as brute force or through environmental coercion. Additional methods include the use of the media and/or propaganda which can be initiated by the player, another player, or AI representing Earth interests.

The means in which the workforce itself is influenced is through their needs and interests (see section Workforce Units Needs). Workforce units will go about striving to fulfill there needs. This depends on the situation at hand and at the players influence.

In avaneya there will be multiple levels of cultural and economic variance influencing different cultural groups to different degrees causing players to deal with this to keep all aspects of the city functioning while still trying to achieve player goals such as technological innovation and maintain an independent yet peaceful existance with Earth.

One group of humans with similar workforce types might make business deals to get it out of the city or off the planet. Another might petition government to increase militia spending. Another might encourage the people to take to arms and fight. And another might join the militia. Outliers may take to vigilante justice or investigate to discover possible conspiracies. Of course the degree to which these can occur depend largely on previous player input to the population. There is a multitude of means in which entities will respond to player sitations that threaten or improve their circumstances and these are unique to the needs being effected and resource the unit has available to make a change or maintain safety.

When there are multiple threats to the same entity, it will respond in order of importance to which it must meet its needs. Clearly air, then water, then food, then security, then belonging. Entities will not respond to their need to belong when they are dieing from a lack of oxygen.

In its natural state a workforce unit is only trying to satisfy its needs. This may be going to a market, going out to lunch, or stealing from a grocery store. In its default state the unit pursues tasks, or actions, in accordance with its needs, using its interests as a guide.

To determine how a unit acts is a matter or factoring in its various needs at the time along with variables such as workforce, income level, social status, and location. In this way we determine whether an unit goes home to cook food for dinner, or goes out to dinner with friends, or goes home has dinner then goes bowling. Factors such as the units income or money, degree of hunger, location to food, food quality, need to belong, number of friends (or daily interactions), neighbourhood safety, "governmental" dietary guidelines, work security, and time available, all converge to determine a units action. 

Not everything will be required to have an action, breathing and drinking water are just factored into resource use and will not evoke a reaction unless the unit is unable to obtain or is blocked from this resource.

As each workforce unit is represented by an individual unit in game, moving and performing tasks, as this unit moves through game space it is consuming resources such as air, water, food, and minerals and like subtances through consumerable resources. What a unit can consume is based in part by what is available and what its needs and action influences are

% -> SPREADSHEET -> Example Entity Actions [an (E) signifies emergency/fight flight response]
Oxygen Actions
-breath 
-move to air source (E)
-stay calm (E)
-panic and scramble for air (E)
-murder to lessen burden on air supply (E)
-partition government for cheaper air
-protest for cheaper air
-steal air supply
-riot for cheaper air
-ask for air
-beg for air

Water Actions
-move to water source
-stay calm
-panic and scramble for water (E)
-murder to lessen burden on water supply (E)
-partition government for cheaper water
-protest for cheaper water
-riot for cheaper water
-steal water
-ask for water
-beg for water


Food Actions
-eat food
-cook food
-buy food
-ask for food
-beg for food
-steal food 
-go to resturant/cafe/food court/ etc
-go to resturant/cafe/food court/ etc with friend
-murder for food (E)
-kill animal
-partition government for cheaper food
-protest for cheaper food
-riot for cheaper food (E)


Security Actions
-partition government for: militia; police; emergency services
-partition government for: attack neighbouring city; ethnic cleansing
-protest government for: above
-riot for: above
-buy house
-rent house
-squat
-build house
-partition/protest/riot government for: cheaper housing
-purchase firearm/weapon
-partition/protest/riot government for: access to firearms/weapons
-fitness/martial arts
-vigilante group


Belonging Actions
-
-
-
-
-
-


There are two broad categories which influences a units needs. The first is the units own actions, drinking water, eating food, getting a job, visiting a friend, all reduce the need to obtain these things, while excerising increases the need for food and water, and working long hours in isolation increases the need for belonging. The second is external or environmental effects such as propaganda, lack of clean air, food with poor nutitional content, or the action and views of other units. All these happenings have varying amounts of influence on the units needs which in turn effect the manner in which the unit acts to satisfy these needs.

In order to correctly map the degree of influence various needs have on a unit and the degree to which a units actions relieves or increased its needs, all of the variables effecting the needs of a unit are mapped with a quantitative point system. Some of these influences will be wide reaching, effecting many units, such as chemicals in the cities water supply or television, other will be specific, such as health consequences from working shift work. 


%-> SPREADSHEET -> Influences
-Workforce Types
-Food Types
-Actions
-Information Types
-Economic 
-Drugs
-Social
-Ostracism



%SAMPLE  (Before) / (after)
%Input		Effects
%		security	food		belongingness				
%Crack Cocaine	+2 / -10	+10 / -5	+10 / -10			 
%
%Running	+1		-2		
%
%
%Enemployed	-3				-2*
%
%
%
%
%*depends on social system.




%--------------> WORKFORCE CONSUMPTION <---------------------------------------
\StartSubSection{Workforce Consumption}

Units aim to fulfill needs, the method in which this is done is via their interests.

Units undertaking actions, buying goods and services (among other things), preferenced by their variables and preferenced further by the hierarchy of needs.?



 



%DISCARDED NOTES:
There will be underlying conditions that will determine what a cities workforce is comprised of and players will have certain degress of influence over the direction the workforce takes but if certain conditions aren't met the citizenry may not accept or appreciate the current workforce type with the consequences of innaction by the player being varried. How players make these conditions and the direction they take with there workforce is a choice made in the sandbox style of play, it is a creation. They can create anything from human/animal labour pool to complete nanocracies with robotic workforces and humans indulging in deccandent pleasure palaces. The player can manage his city through inacting certain ordinance, directing funds, and other general city management processes including placement of civic buildings and diplomacy. 
