% This is part of the Avaneya Project Crew Handbook.
% Copyright (C) 2010-2016 Cartesian Theatre <info@cartesiantheatre.com>.
% See the file Copying for details on copying conditions.

% Bootstrap Teams section...
\StartSection{Bootstrap Teams}

The initial fictional bootstrap crew carried by the {\it Avaneya} spacecraft is organized into teams. Some of the teams are temporary whereas others are intended to address recurring problems. Some team members may be members of more than one team, and some team leaders may be leaders of more than one team.

The \index{construction team}construction team consists of specialists with training in a variety of disciplines. This includes cold weather construction experts and structural engineers. They are responsible for building erection in the early world described in the fictional timeline of \in{chapter}[Timeline].

The \index{experimental team}experimental team consists of specialists that conduct experiments both {\it en route} and {\it in situ}. This includes a xenobiologist. \goto{Henrik Nørgaard}[Henrik Nørgaard] is the team lead.

The \index{greenhouse team}greenhouse team is responsible for providing the population with a stable food supply. This includes biogeochemists, algaculturalists, and geologists. Some also function as medical assistants to Henrik. \goto{Nayana Rai}[Nayana Rai] is the team lead.

The \index{mechanics team}mechanics team is responsible for ensuring that all machinery is properly functioning and carrying out any repairs that may be needed. It includes at least mechanical engineers and mechanics skilled at troubleshooting problems that may arise in the field. \goto{Senka Rukavina}[Senka Rukavina] is the team lead.

The \index{recovery team}recovery team is a temporary team responsible for tracking and recovering all cargo jettisoned from {\it Avaneya} upon its arrival.

The \index{survey team}survey team is responsible for assessing various locations on their merits of habitability, accessibility, mineralogy, other available resources, such as water ice, and so on. The team includes at least biogeochemical and geological specialists. \goto{Senka Rukavina}[Senka Rukavina] is the team lead.

The \index{systems team}systems team is responsible for all communications and other electronic equipment. They include specialists in artificial intelligence, knowledge engineers, cerebral mechanics, cyberneticists, and communications. \goto{Khalad Zafar}[Khalid Zafar] is the team lead.

The \index{terraformation team}terraformation team is responsible for surveying and reporting back to UNSA on the terraformation efforts. It includes at least astrogeophysicists and areobotanists. \goto{Nayana Rai}[Nayana Rai] is the team lead.

