% This is part of the Avaneya Project Crew Handbook.
% Copyright (C) 2010, 2011, 2012
%   Kshatra Corp.
% See the file License for copying conditions.

% Euphemisms section...
\StartSection{Euphemisms}

% At 800 new arrivals every year, each Aravan capable of transporting 400 
% persons twice every year, this would take about ten Terran years or 5.6 
% MYrs...
\StartTimelineDate{14 Scorpius, 3 B.R.}
The Arcadian Settlement's annual census notes a population now in excess of 8,000 colonists and growing.
\StopTimelineDate

\StartTimelineDate{22 Scorpius, 3 B.R.}
Originally published by the {\it RANND Corporation}, Edward Geisler's seminal work, {\it Transplanetary Commerce with Mars: The Red Horn of Plenty}, is hailed by industry leaders for the influence and achievement the white paper, long esteemed as a classic of great notoriety, culminates in as the first of a series of private launches en route to Mars departs Earth. Beginning from Kiruna, European Union, the heavy lift spacecraft {\it Arrogo} bears a payload of both personnel and cargo to be received at the starport in the Arcadian Settlement. It is already validated with an approved UNSA application.

The launch is celebrated by industry leaders as a landmark transformation as the first transnational corporations promote themselves to the first transplanetaries. The mission stakeholders consist of an international private consortium from a variety of sectors ranging from soft drink vendors to agricultural biotechnology.

Geisler's paper discussed the means and opportunities off-world commerce with Mars theoretically would provide and draws historical analogies with the European expansion into the \quote{new world} of the Americas and the great wealth that was tapped. The paper is identified as the original driving force behind the industry's efforts that succeed in the day's launch.
\StopTimelineDate

\StartTimelineDate{1 Sagittarius, 3 B.R.}
Arda receives a UNSA directive to terminate an orthomolecular and several other alternative medical research programs occupying settlement laboratory space. UNSA intends to lease the vacated facilities to the pharmaceutical company GlaxoKlineSmith.
\StopTimelineDate

\StartTimelineDate{41 Sagittarius, 3 B.R.}
Arda receives a UNSA directive to terminate the settlement's kombucha brewery research program. All equipment and most other material assets are to remain behind in the vacated facilities. 

Although consumption of the \quote{Arcadian Manowar}\index{Arcadian Manowar}\footnote{The Portuguese man o' war is a species of jellyfish ({\it Physalia physalis}) from which the beverage draws its name. It is said to resemble the \quote{mushroom} used in the drink's fermentation.} beverage had been a cultural practise in the settlement for years, its widespread adoption had been preceded with the discovery that those exposed to higher than average cosmic radiation, which is to be expected during extended EVAs for instance, had on average benefited from its consumption. Research conducted within settlement laboratories using {\it in vitro} trials provided compelling evidence that the fermented tea aided in the repair of radiation-induced cellular DNA damage. 

Until now, the beverage had been a minor export back to Earth where it had long been prized, widely sought after, and considered superior to any Terran variant. The biochemistry of the age old assortment of bacteria and yeast that acts as a symbiotic colony for its fermentation, as used for centuries on Earth, but metabolizing now on Mars under a lower than Earth gravitational field, has been implicated in popular Terran urban legends as the cause for its unique and unrivalled quality. 

The difficulty in obtaining a genuine bottle from a very limited stock, drawn from a marketplace already saturated with Asian Union counterfeits, has always ensured it remained an expensive novelty on Earth. This had been the case ever since the early railgunned bottled imports had been received on Earth and celebrated as the first consumables ever imported from off world. Since then, its possession and consumption became symbolic of personal wealth. With the termination of the settlement's program, the diminishing modest stocks remaining on Earth are expected to be auctioned off at great expense to private collectors.
\StopTimelineDate

\StartTimelineDate{22 Capricorn, 3 B.R.}
Arda receives a UNSA directive to reduce greenhouse water usage. The savings are to support a recently opened Coke bottling facility on the site of the settlement's former kombucha brewery.
\StopTimelineDate

\StartTimelineDate{47 Capricorn, 3 B.R.}
Arda receives a UNSA directive to terminate all double blind herbal research studies being conducted by settlement's laboratories hosting alternative medical research. UNSA cites the NAU Congress and EU Parliament having recently passed legislation listing herbs currently under settlement research, such as St. John's Wort, echinacea, and a number of others, as contraband. 

Arda pleads unsuccessfully with UNSA to ban GlaxoKlineSmith and all other lobbyists from influencing UNSA policy.
\StopTimelineDate

\StartTimelineDate{30 Aquarius, 3 B.R.}
Arda receives a UNSA directive to provide logistical support to recently landed Clown Food contractors requiring assistance in erecting several kilometers of facilities. The facilities are for the production of genetically modified beef, pork, and potatoes to provide for a newly expanded franchise with restaurants across the settlement and its neighbouring starport.

Arda beckons UNSA to alter its decision. He claims that the tillable land required for animals is not found on Mars but must be created through a process which is resource intensive. He claims further that most of the energy used in the raising of warm-blooded herbivores for human consumption is inefficiently radiated off into space as heat long before it can be consumed, making veganism not simply an arbitrary dietary preference, but a practical operational requirement. In a world where resource management is critical and the margin for error much smaller than on Earth, Arda concludes that Clown Food's requirements will burden the Arcadian Settlement unnecessarily. 

UNSA does not respond to the concerns Arda raises. Arda attempts to revoke the contractors' visas. UNSA overrides the revocation, reprimands Arda, and reasserts the original directive.
\StopTimelineDate

\StartTimelineDate{4 Pisces, 3 B.R.}
Arda receives a UNSA directive to terminate several mushroom ranches. The ranches were used to recycle plant waste into high quality sources of edible protein. The energy used in maintaining ranch temperature is reallocated for railgun use. All plant waste destined for the ranches, or as composted greenhouse fertilizer, must now be transported to newly created outdoor garbage fills across the planitia. Outdoor atmospheric conditions and cosmic radiation ensure the organic matter will neither degrade nor become reusable.
\StopTimelineDate

\StartTimelineDate{19 Aries, 3 B.R.}
Arda receives a UNSA directive to secure all new settlement energy requirements from private energy contractor Haliburton. Haliburton will provide UNSA with nuclear reactors that will superannuate the settlement's standard practise of expanding primarily through new geothermal wells and secondarily through additional photovoltaic panels. The directive specifies that these purchases are to be absorbed by the settlement's Terran sustainability research budget.
\StopTimelineDate

\StartTimelineDate{51 Taurus, 3 B.R.}
Arda receives a UNSA directive to terminate several settlement research labs currently investigating Terran sustainability. Some of these include key biophotoreactor plants that produce kelp fertilizer for greenhouse use. The resources freed are to be redirected to increased deuterium production for Terran export and its requisite energy demanding railgun use. Arda reluctantly complies with the directive. UNSA cites technical reasons while Arda cites settlement debt.
\StopTimelineDate

