% This is part of the Avaneya Project Crew Handbook.
% Copyright (C) 2010, 2011, 2012
%   Kshatra Corp.
% See the file License for copying conditions.

% Civil Disobedience and False Flags section...
\StartSection{Civil Disobedience and False Flags}

\StartTimelineDate{1 Gemini, 2 B.R.}
The beginning of a new year is marked with the first of a long series of partisan meetings for settlement independence held in secret. Arda is in attendance. The underground effort is discreetly coined {\it Free Arcadia}\index{Free Arcadia}.
\StopTimelineDate

\StartTimelineDate{4 Pisces, 1 B.R.}
Arda wins a landslide election and accepts executive office as President of a newly created interim settlement government of limited self-imposed independence. Arda and his staff will operate from within the newly erected {\it Arcadia Hall}. UNSA neither authorizes the interim government nor acknowledges the occurrence of an election.

Kissinger strongly cautions Arda against {\it "the formation of redundant administrative offices inconsistent with UNSA interests"}. On Earth, the Security Council holds an emergency meeting in New York on the settlement situation.
\StopTimelineDate

\StartTimelineDate{7 Pisces, 1 B.R.}
Promoted from the position of United Nations Deputy Secretary-General, Security Council Resolution 12661 is adopted appointing Kissinger to the newly created {\it Office of the Governor of the Arcadian Settlement} in response to the settlement's self initiated election. Though UNSA does not actually acknowledge the election in the wording of the resolution, the appointment asserts the Governor's seniority to any of the settlement's self elected public officers. Arcadia Hall does not issue a statement in response to the Security Council resolution.
\StopTimelineDate

\StartTimelineDate{50 Taurus, 1 B.R.}
Arda puts the issue of settlement independence to a publicly held vote of 15,343 colonists. The vote is carried at 98 \%.
\StopTimelineDate

\StartTimelineDate{51 Taurus, 1 B.R.}
The asteroid {\it 52048 Varuna} aerobreaks into the Martian atmosphere, disintegrating as predicted. Meteor showers are reported impacting the south pole's dry ice sheets by the Terraformation Team.
\StopTimelineDate

\StartTimelineDate{55 Taurus, 1 B.R.}
Arcadia Hall issues a public advisory to alert the settlement of Arda having gone missing. The advisory is recalled when he is spotted engaged in private reflection, drafting a document under the light of a midnight desert aurora.
\StopTimelineDate

\midaligned{\bf \star \hspace[star_hspace] \star \hspace[star_hspace] \star}

% Year zero...
\StartTimelineDate{1 Gemini, 0 A.R. (Year Zero)}
Intending to reach both planets, Arda formally addresses the Secretary-General of the United Nations through an open letter transmitted over the solnet intended for public consumption. He reflects on the predicament faced by the settlers.

\startTimelineCorrespondenceDocument
    \input Source/Timeline/Arda_Independence_Letter.tex
\stopTimelineCorrespondenceDocument

Arda's publicly accessible transmission over the solnet is accompanied with copies of all settlement research and literature to date, including a copy of the {\it Rubicon Act}. 

The {\it Rubicon Act}, also known more formally as the {\it Constitution of the Republic of Arcadia Planitia}, is the first of the settlement's self-declared legislation and aims to initiate and secure its independence. The document was the product of years of drafting through covertly held partisan meetings on Mars and, if successful, would beget the first off-world autonomous city-state in human history. The cultural significance to Arcadians is such that their calendar posits its year zero relative to it (see \in{subsection}[Arcadian Calendar] for more detail). Notable portions summarized follow:

\startTimelineGeneralDocument
    \input Source/Timeline/Rubicon_Act_Summary.tex
\stopTimelineGeneralDocument

A failed denial-of-service attack lodged against the Terran solnet downlink station at UNSA's Jet Propulsion Laboratory on Earth attempts to disrupt reception of Arda's transmission. Station technicians determine the attack originated from within the station's own local intranet from behind a restricted access zone using several zero day vulnerabilities in the station's transceiver software.

Arda's independence letter propagates rapidly over peer to peer networks. Responses on Earth among Terrans following Martian events through independent media are predominantly in favour of the settlement's commitment to independence. Mainstream media sensationalize and relate the situation as a costly mission in crisis with a Mission Commander of now questionable intentions and character.
\StopTimelineDate

\StartTimelineDate{2 Gemini, 0 A.R.}
Arda arrives by dirigible at {\it Valles Marineris}, a 4,000 km long canyon reaching up to 7 km in depth. He spends the sol in solitary reflection.

The former NASA had long ago salvaged aluminium recovered in 2001 A.D. from the ruins of the former World Trade Centers in lower Manhattan for reuse in their Mars rover {\it Spirit's} cable shield. To remind future Arcadians of the reasons for independence, Arda has the {\it Constitution} engraved in aluminium recovered from the derelict craft as it came to finally rest at its UNESCO World Heritage Site on Mars. The plate is kept at Arcadia Hall on public display where it becomes known as the {\it Lamothe Plate}. This is in reference, and out of reverence, for Dr Richard Lamothe, renowned Nobel prize winning French astrophysicist and areologist. Lamothe passed several decades ago, but volumes of all of his works are kept amongst Arda's extensive personal library.

NASA had originally selected the metal for reuse with {\it Spirit} as a symbol of perseverance in the face of evil. Arda, convinced both the metal and the message it embodied was wanting of further refinement, selected it for reuse in the Lamothe Plate as a symbol of perseverance in the face of government.
\StopTimelineDate

\StartTimelineDate{14 Gemini, 0 A.R.}
A Survey Team is deployed on a routine assignment using ground based shallow radar to scout for artesian aquifers. Travelling solo, Arda departs Arcadia to rendezvous later with the team as previously scheduled. Midway he performs an EVA to dismount the rover and check one of its tires. While outside, he is knocked from his feet as an explosion occurs on the other side of the rover from within the cabin. Although his forearm is shattered, he sustains no life threatening injuries. With the rover's first aid kit destroyed, he improvises a make shift caste around the suit's elasticized mesh by applying liquid water to solidify coatings of supportive regolith around his forearm.

Arda then inspects the rover to find the cabin fully depressurized and the airlock's seal too badly damaged to maintain positive pressure. The rover's electronic controls are non-responsive, rendering the machine derelict in the middle of a Martian desert. The nearest rescue effort that acknowledges his emergency VHF transponder is an aquifer drilling rig crew at least 42 hours away. With only 35 hours of liquid oxygen available for his respirator, he extends it in time by syphoning cryogenic liquid oxygen from the rover's propulsion system.

The damaged rover is later towed back to a service bay in the city. Mechanics find fragments of a container with a gaseous methane / oxygen (\chemical{CH_4 / O_2}) signature that had been under 35,000 kPa. The mechanics rule out the rover's own equipment since its fuel cells are methanol based (\chemical{CH_3OH}) and no other known onboard equipment requires methane. In an effort to determine who last serviced the vehicle, the mechanics note that the electronic maintenance log is found to be missing. Remnants in the cabin of what appear to be fragments of a timer based detonator are recovered with its components not matching any of the rover's.
\StopTimelineDate

\StartTimelineDate{10 Cancer, 0 A.R.}
Arcadia Hall amends the Constitution by passing the {\it Humanoid Act}, stripping corporate legal fictions of the rights of human beings\footnote{See \in{section}[Corporate Personhood] for more on the concept of corporate personhood.} and denying money as a form of free speech.\footnote{See \in{section}[Money As Free Speech] for an explanation on how money has been considered a form of free speech.} Transplanetaries that had their registered offices within Arcadia's territorial region do so now without the safety of corporate personhood.
\StopTimelineDate

% Coke pulled out of Vietnam the year before the United States military did...
\StartTimelineDate{42 Cancer, 0 A.R.}
{\it Coke} quietly closes down its production and bottling facilities. Several other transplanetaries, including the successful fast food franchise {\it Clown Food}, are seen at Starport Arcadia making preparations to return to Earth. None of the transplanetaries preparing to return to Earth issue public statements.
\StopTimelineDate

\StartTimelineDate{62 Cancer, 0 A.R.}
No. \#2 Railgun is taken offline for unscheduled rail and insulator surface repair by Terran UNSA contractor, {\it Solar Urban Moving Systems}.
\StopTimelineDate

\StartTimelineDate{20 Leo, 0 A.R.}
No. \type{#}2 Railgun's blockhouse reports an explosion that has taken out its compulsator power supply. The single largest loss of life ever to occur on Mars unfolds as two-hundred and fifty three people are killed and seven more injured as an adjacent residential district's inflatable dome in Arcadia's Fourth Precinct depressurizes from shrapnel.

Arda convenes an emergency meeting with all state secretaries present at Arcadia Hall. Forensic evidence of a cryobomb's (\chemical{CH_4/O_2}) chemical residue is presented. Chemical analysis suggests that the recovered components are of probable non-Martian origin. 

Before the meeting adjourns, Terran media implicate the alleged identities of the \quote{railgun bombers} as Free Arcadia partisan extremists with ties to Arda.

Shortly after the meeting adjourns, Arda begins reading a presidential address from Arcadia Hall covering his staff's preliminary findings. However, the live broadcast is cutoff prematurely from reaching a Terran audience due to technical difficulties with the {\it Mars Enhanced Telecommunications Orbiter}. The satellite used to broker nearly all communication between the two planets is remotely taken offline for an unscheduled systems diagnostic by {\it Bronfman-Murdoch Aerospace} contractors at UNSA's Jet Propulsion Laboratory on Earth. Terran mainstream media networks reassure Arcadia Hall's Office of the President that Arda's presidential address will be rescheduled for replay later, simultaneous the Terran Superbowl.

Kissinger meets with Arda privately and advises him to step down from his political position of President in light of the optics of the situation amidst growing pressure from UNSA back on Earth. However, he maintains that UNSA still supports him in the predominantly technical capacity of Mission Commander due to the influence and respect wielded across the colony, as well as to oversee the immediate resumption of stable deuterium exports. The Governor, as a final suggestion, advises Arda to \quote{dismantle all indigenous self-governance apparatus immediately due to its redundancy with UNSA and inability to guarantee the safety of the colonists}.
\StopTimelineDate

\StartTimelineDate{21 Leo, 0 A.R.}
Arcadia Hall instructs Starport Arcadia to begin fuelling {\it ERV Bravo}, Earth-bound, to take advantage of a minimum-energy launch window. They authorize the confiscation, revocation, and destruction of Kissinger's visa. A security detail takes him into custody where the irate Governor is escorted to the starport. An automated solo departure lifting off later that evening will return him to Earth.
\StopTimelineDate

\StartTimelineDate{31 Leo, 0 A.R.}
The Terraformation Team at the south pole report findings of dry ice sublimation in response to the asteroid {\it 52048 Varuna's} impact.
\StopTimelineDate

