% This is part of the Avaneya Project Crew Handbook.
% Copyright (C) 2010, 2011, 2012
%   Kshatra Corp.
% See the file License for copying conditions.

% Countdown section...
\StartSection{Countdown}

\StartTimelineDate{38 Leo, 0 A.R.}
The United Nations Security Council passes resolution 12664 where it is adopted unanimously. Announcement of the resolution is not yet made public, but deferred. The resolution begins with:

\startTimelineGeneralDocument
The Security Council,

Reaffirming the principles and purposes of the Charter of the United Nations,

Determined to combat by all means threats to transplanetary peace and security caused by terrorist acts,

Deploring the gross and systematic violation of human rights, including the repression of peaceful demonstrators, expressing deep concern at the deaths of civilians, and rejecting unequivocally the incitement to hostility and violence against the civilian population and infrastructure excited through {\it in situ} administrative elements of the United Nations Arcadian Settlement,

Recognizing the inherent right of individual or collective self-defence in accordance with the Charter,

\startitemize[R]
\setupwhitespace[big]
\item Unequivocally condemns in the strongest terms the horrifying terrorist attacks which took place recently at the site of the No. \type{#}2 Railgun on Mars in the United Nations Arcadian Settlement and regards such acts, like any act of international terrorism on Earth, as a threat to transplanetary peace and security;

\item Expresses its deepest sympathy and condolences to the victims, their families, and of the United Nations Arcadian Settlement;

\item Calls on all Mars Treaty signatories to work together urgently to bring to justice the perpetrators, organizers and sponsors of these terrorist attacks and stresses that those responsible for aiding, supporting or harbouring the perpetrators, organizers and sponsors of these acts will be held accountable;

\item Calls also on all Mars Treaty signatories to redouble their efforts to prevent and suppress terrorist acts including by increased cooperation and full implementation of the relevant anti-terrorist conventions and Security Council resolutions;

\item Expresses its readiness to take all necessary steps to respond to the terrorist attacks of the Arcadian Settlement railgun bombing, and to combat all forms of terrorism, in accordance with its responsibilities under the Charter of the United Nations;

\item Authorizes Mars Treaty signatories that have notified the Secretary-General, acting nationally or through regional organizations or arrangements, and acting in cooperation with the Secretary-General and the Office of the President of the Space Agency, to take all necessary measures to protect civilians, civilian populated areas, and assets under threat of attack in the United Nations Arcadian Settlement, and requests those signatories concerned to inform the Secretary-General immediately of the measures they take pursuant to the authorization conferred by this paragraph which shall be immediately reported to the Security Council;

\item Decides to remain actively seized of the matter.\footnote{Most Security Council resolutions end with this phrase. Article 12 of the United Nations Charter states that while it \quote{\it is exercising in respect of any dispute or situation...the General Assembly shall not make any recommendation...unless the Security Council so requests.} In other words, the General Assembly of many is being instructed by the Security Council of few to not interfere in an affair the latter still wishes to handle exclusively for the time being.}
\stopitemize
\stopTimelineGeneralDocument

The resolution recommends that the {\it United Nations Department of Peacekeeping Operations (UNDPKO)} immediately place the multinational contingents under its permanently standing {\it Rapid Reaction Force} on high alert. This is authorized under Article 51\footnote{Article 51 of the Charter of the United Nations asserts \quote{\it the inherent right of individual or collective self-defence if an armed attack occurs against a Member of the United Nations.}. It is frequently cited as the legal basis for the use of force.} of the Charter.
\StopTimelineDate

\StartTimelineDate{44 Leo, 0 A.R.}
The {\it Non-Aligned Movement}\index{Non-Aligned Movement} issues a {\it Statement to the United Nations Security Council on the Situation Between the United Nations and the Republic of Arcadia Planitia}. The association reaffirms the principles of non-intervention, non-interference, the Arcadian sovereignty, and its right of self-determination.

The organization is an association of Terran states advocating the doctrine of non-alignment. This is defined as respect for independence, sovereignty, territorial integrity, and security of states as they strive for freedom from imperialism, all forms of colonialism, racism, foreign aggression, occupation, domination, and interference. Its constituents include the states of Venezuela, Iran, Cuba, and others.
\StopTimelineDate

\StartTimelineDate{2 Virgo, 0 A.R.}
The United Nations Department of Peacekeeping Operations' {\it Joint Chiefs of Staff} assemble in Brussels, European Union. There they liaise with government officials and other high level military UN servants for the first day of a three Terran month summit. The summit is conducted under maximum security. Though its existence is made public, the exercise of a high level of discretion in all of its activities ensures that its nature is not. This draws public criticism due to the presence of a number of high level state officials from across the planet and the lack of transparency.

Rumours spread rapidly over the solnet purporting the purpose of the summit as one of planning the first off world military occupation of another planet in history. Leaked documents suggest the summits agenda is to assess the feasibility, followed by addressing the preparation of all strategic, intelligence gathering, logistical, personnel, and financial planning necessary to conduct such an operation. Known high level UNSA liaisons, including UNSA's President Dr. Samuel Lieberman, are captured on camera being hustled in and out of meetings within the compound. Public speculation is further compounded when representatives of major international financial institutions and the NAU-CIA director are also captured on camera by independent media.
\StopTimelineDate

\StartTimelineDate{27 Libra, 0 A.R.}
UNDPKO's summit in Brussels concludes with the establishment of the newly formed {\it Mars Command (MARSCOM)}. It is established under the civilian administered UNDPKO and is now responsible for overseeing all of the latter's peacekeeping operations at the strategic level, in the route to, in the vicinity of, and on the Red Planet. The headquarters is located at Peterson Air Force Base, North American Union, and is jointly staffed by an international consortium of military personnel from all nations signatory to the Mars Treaty.

A large volume of requests for quotation for the procurement of military equipment are disseminated to a multitude of eager defence contractors.
\StopTimelineDate

\StartTimelineDate{34 Libra, 0 A.R.}
INTERPOL issues an international arrest warrant in the form of a \quote{red notice} for Arda Baştürk.
\StopTimelineDate

\StartTimelineDate{36 Libra, 0 A.R.}
Lockheed Martin-Boeing's Skunk Works advanced research facilities begins refitting UNSA's {\it Tarikin III} personnel ferry into the {\it Kali}, a militarized, remotely piloted, spacecraft carrying a payload of personnel and materials. This is in response to a UNSA issued contract for a classified mission payload specification issued by MARSCOM's Joint Chiefs of Staff.
\StopTimelineDate

\StartTimelineDate{22 Scorpius, 0 A.R.}
Selected Rapid Reaction Force personnel are assigned to UNSA training facilities in Antarctica and Huelva, Spain, now administered by MARSCOM.
\StopTimelineDate

\StartTimelineDate{23 Scorpius, 0 A.R.}
Arda receives an asymmetrically encrypted message warning of an allegedly imminent plan that was drafted by a consortium of Terran financial interests. The information was allegedly drawn from leaked source material made possible through software vulnerabilities in the proprietary firewall protecting UNSA intranets. The anonymous tip-off claims these interests have influenced the United Nations such that it is prepared to organize and land a brigade of peacekeepers to Martian theatre. The brigade will operate under UNSA's newly formed MARSCOM, which itself operates as an organ of the United Nations, which in turn allegedly answers to the Terran central banking cartels. 

The field commanders on the ground would be tasked with several prioritized strategic objectives while in theatre. 

\startTimelineCorrespondenceDocument
TOP SECRET
DISTRIBUTION RESTRICTED / FOR NAMED RECIPIENTS ONLY
MARSCOM JOINT CHIEFS OF STAFF
\color[colour_item]{MEMORANDUM CONCERNING PRIORITIZED STRATEGIC OBJECTIVES FOR ARCADIAN SETTLEMENT}
\startitemize[4]
%\setupwhitespace[big]
\item \color[colour_item]{ALPHA} - KILL OR CAPTURE ATATÜRK\footnote{The NAU-CIA attaché to MARSCOM use this cryptonym for Arda.}

    \startitemize[n]
    \item IF CAPTURED, PERFORM IMMEDIATE EXTRACTION TO KALI, AREOSTATIONARY ORBIT. HOLD IN CUSTODY. AWAIT EXTRAJUDICIAL RENDITION EARTHSIDE.
    \item IF KILLED, IMPLICATE 'RED UNIONISTS'. MEDIA PREBRIEFED. BHADRA 2 COMMSAT WILL STANDBY TO BROADCAST NEWS EARTHSIDE.
    \stopitemize

\item \color[colour_item]{BRAVO} - CAPTURE ARCADIA
    \startitemize[n]
    \item INSERT NEW PLAINCLOTHES NAU-CIA, SPEC FORCE, AND BLACKWATER ASSETS INTO ARCADIA. WILL PROVIDE INTEL.
    \item RULES OF ENGAGEMENT AUTHORIZE NEUTRALIZATION OF ANY LOCAL RESISTANCE, W/ NO QUARTER. EXPECT MINIMAL, IF ANY, FIGHTING IN BUILT UP AREAS.
    \item DISARM SIERRA.\footnote{MARSCOM's phonetic alphabet designation for a settler.} CONFISCATE ALL SMALL ARMS.
    \item ENHANCED INTERROGATION OF ALL SUSPECTED FAP\footnote{Free Arcadia partisans.} CAPTURED TO LOCATE SIERRA'S WEAPON'S CACHES AND OTHER INTEL. PERFORM EXTRAJUDICIAL DETENTION, TRIAL, AND EXECUTION AS NECESSARY. ATATÜRK EXCEPTION, VIZ OBJECTIVE ALPHA(ONE).
    \item MAINTAIN STARPORT ARCADIA OPERATIONAL.
    \item SEIZE CONTROL OF MAX NO OF MATERIAL ASSETS INTACT.
    \item RESTORE UNSA ADMIN, BROKERED THROUGH MARSCOM'S INITIAL COMMAND AND CONTROL UNTIL STABLE.
    \stopitemize

\item \color[colour_item]{CHARLIE} - EXTRACT ALREADY DEPLOYED INTEL ASSETS
    \startitemize[n]
    \item EXTRACT AND DEBRIEF NAU-CIA ASSETS ALREADY DEPLOYED IN THEATRE SINCE LANDING 28 MONTHS PRIOR (TARIKIN III).
    \stopitemize

\item \color[colour_item]{DELTA} - ECONOMIC REFORMS
    \startitemize[n]
    \item REINSTATEMENT OF SETTLEMENT LABOUR AND KEY PERSONNEL NECESSARY FOR RESUMPTION OF STABLE DEUTERIUM AND OTHER KEY EXPORTS.
    \item PROTECT COMBAT SUPPORT SQUADRON'S ERECTION OF 'FIRST CENTRAL BANK OF THE UNITED NATIONS ARCADIAN SETTLEMENT'.
    \item REVERT ARCADIAN JENYA BACK TO TERRAN BANCOR.
    \item CAPTURE ALL RHODIUM BACKED JENYAS HELD IN RESERVE.
    \stopitemize

\item \color[colour_item]{ECHO} - POLITICAL REFORMS
    \startitemize[n]
    \item IMPLEMENT MARTIAL LAW OF SUFFICIENT TIME TO ACHIEVE ALL STRATEGIC OBJECTIVES.
    \item ASSIST NAU-CIA ASSETS RIG ARCADIA HALL ELECTION FOR OFFICE OF THE PRESIDENT. ALL BALLOT OPTIONS COMPROMISED. CANDIDATES SYMPATHETIC TO WHITE LISTED TERRAN INDUSTRIAL AND FINANCIAL INTERESTS PROVIDED IN SCHEDULE D.
    \stopitemize
\stopitemize

IN UNLIKELY EVENT MARSBRIG UNABLE TO HOLD ARCADIA, EMERGENCY CONTINGENCY PLAN'S STRATEGIC OBJECTIVES FOLLOW:
\startitemize[4]
\item \color[colour_item]{FOXTROT} - INSERT AND ACTIVATE 'FIFTH COLUMN'

    \startitemize[n]
    \item ALL ASSETS, EXCEPT OBJECTIVE FOXTROT(TWO)'S, FALLBACK TO RENDEZVOUS TANGO. AWAIT EXTRACTION TO KALI, AREOSTATIONARY ORBIT. ORDERS WILL FOLLOW.

    \item IF OBJECTIVE BRAVO(ONE) OK, THESE ASSETS REMAIN BEHIND IN SIERRA SAFE HOUSES AS RESIDENT FIFTH COLUMN CONTRAS. IF POSSIBLE W/O COMPROMISING COVER, FIFTH COLUMN WILL PROTECT RETREATING MARSCOM ASSETS.

    \item FIFTH COLUMN CONDUCTS INSURGENCY OPS, EXECUTING FALSE FLAGS MASQUERADING AS 'RED UNIONISTS'. CAN CONDUCT OPS INDEPENDENTLY W/O NEED FOR COMM CONTACT IF BROKEN. PROVIDES MARSCOM W/ CLANDESTINE UNIT BEHIND ARCADIAN DEMARCATION LINE FOLLOWING STRATEGIC OR TACTICAL RETREAT.

    \item IF OBJECTIVE DELTA(THREE) FAILED, FIFTH COLUMN MUST RAISE FUNDS IN LOCAL JENYAS TO SUPPORT INDEPENDENT OPS VIA NARCOTRAFFICKING. SIERRA'S SEIZED LABS POSSIBLY CONDUCIVE TO DOMESTIC SYNTHETIC PRODUCTION. SYNTHETIC METHOD PROVIDES PLAUSIBLE DENIABILITY, GREATER RELIABILITY, AND W/O NEED FOR TERRAN IMPORT.
    \stopitemize
\stopitemize
\stopTimelineCorrespondenceDocument

The anonymous informant provides detailed information on troop movement to Terran MARSCOM administered training facilities, size of units, weapons and tactics to be employed, probable location of encampment, and route of approach. The brigade will be deployed to Martian theatre under the banner of {\it United Nations Mars Expeditionary Peacekeeping Force (UNMEPF)}. The force draws elements mostly from North American Union and European Union resources specializing in airborne and mechanized infantry, armoured reconnaissance, tactical UAVs, self-propelled artillery, and logistical support.

The message is encrypted using a public key of Arda's not used since his long departed days of a military posting with the European Union's Mountain and Commando Brigade in Hakkari. Khalid examines the digital signature and finds it authentic. Arda and his remaining staff deliberate before deeming the message credible.
\StopTimelineDate

\StartTimelineDate{24 Scorpius, 0 A.R.}
Khalid advises Arda that he can take out UNSA's orbital surveillance capabilities by destroying their {\it Mars Enhanced Telecommunications Orbiter} while the settlement still has access. Arda lets him proceed. UNSA detects the settlement login attempt and transmits a command sequence instructing the orbiter to revoke Arcadia's keypair. They are several minutes too late. The orbiter burns up its entire hydrazine fuel reserves used for attitude control, deorbiting itself and incinerating during atmospheric entry.

All solnet communications between Earth and Mars that had been brokered by METO are now effectively disabled. However, the orbiter had also been used for reliable Arcadian intercommunication. Khalid adapts in its absence by redesigning key settlement communication equipment to combine line-of-sight UHF and shortwave communication, requiring no satellite.

UNSA has its JPL lock down all four {\it Mars Positioning Satellites} and place them in safe mode in expectation of an RAP denial-of-service attack, though there is none. Arda's staff reason the satellites have no use as UNSA surveillance assets and are still useful for their own navigation on the Red Planet.
\StopTimelineDate

\StartTimelineDate{28 Scorpius, 0 A.R.}
The United Nations Security Council makes its having previously passed resolution 12664 public for the first time. The timing coincides with the saturation of key Terran broadcast networks airing a recent adultery scandal of a prominent Terran celebrity.
\StopTimelineDate

\StartTimelineDate{29 Scorpius, 0 A.R.}
Arda holds an emergency cabinet meeting at {\it Arcadia Hall} in secrecy. Several hours into the meeting, Arda summons a group of three of his best geologists and three of his best civil engineers. They are tasked in private before they depart, led by Senka, on rovers to fulfil their assignment.
\StopTimelineDate

% Expediency is plausible if partisan meetings had been held at such a location where surveying had already been done in the past...
\StartTimelineDate{30 Sagittarius, 0 A.R.}
Arcadia Hall issues a city-wide emergency evacuation order\index{evacuation order} that is carried out over the next two weeks. All non-essential personnel are transported to an unmarked installation constructed within a subterranean lava tube. The installation is heated and powered by a geothermal power station with liquid water directly available from an artesian aquifer that supplies it.
\StopTimelineDate

% If Hohmann Type II, assume 8.5 Terran months since he left Mars...
\StartTimelineDate{48 Sagittarius, 0 A.R.}
{\it ERV Bravo} carrying Leonard lands at Edwards Air Force Base, North American Union. The spacecraft is seized as a UNSA asset. Leonard leaves for the City of London\index{City of London} the following day.
\StopTimelineDate

\StartTimelineDate{50 Sagittarius, 0 A.R.}
Arcadia completes its evacuation of all non-essential personnel. The movement of non-essential materials and Arcadia Hall's administrative personnel out of the city to the underground installation commences immediately following. A make-shift facility for the staff and functions of Arcadia Hall is erected within the shelter and given the same name.

Excavation begins for the construction of unpressurized underground passageways linking the shelter which is outside of the city with entrances within it. Usage of tunnel entrances within the city help to ensure that vehicle tracks spotted from orbital telemetry cannot be used to pinpoint the location of the underground installation outside of the city. 

Senka predicts the tracks that are already present on the surface will be erased through surface wind deposition and erosion during the seasonal dust storm, assuming one should occur. This normally occurs during the orbital perihelion when the planet is at its closest to the sun.

Within the underground installation, and with most of the city evacuated, Arda gathers the settlers to provide an explanation for the recent commotion surrounding their evacuation and the dangerous predicament they now face. Influenced by the 18\high{th} century A.D. works of Vattel's {\it Law of Nations}, Arda presents a compelling argument that it is within Arcadians natural rights the option of collective self-defence. In addition, he reminds them that their Constitution already legally permits the raising of militias. On this latter point, an election is held on the raising of militia units to provide protection for both Arcadia's inhabitants and the city itself. The Arcadians are reluctant, but see it as a last resort. The vote carries with a majority.

Arda reasons that since prior military service is not unusual in the field of manned space exploration, putting out a call for all such settlers would be a reasonable place to start. A list is assembled with more than a hundred personnel. Approximately half of that having extensive experience, Arda included.
\StopTimelineDate

\StartTimelineDate{2 Capricorn, 0 A.R.}
Relying on a combination of military training manuals remaining on Arcadian local solnet caches, their own personal experiences, and knowledge of the current state of Arcadia's industrial infrastructure, such as milling and robotic assembly machines, Arda spends the morning reflecting on a measured response to an expected landing of a UN military force. 

The sol also marks the first of a series of meetings held over the next several months on the recruitment, organization, and equipping of an indigenous militia. The meetings are held by an {\it ad hoc} group called the {\it City Security Committee}. It consists of settlement personnel with experience ranging from industrial processes, former military experience, mechanics, leadership and management, chemists, engineers, and others. Acting as a think tank, they are formed to provide potential solutions to the problem of an armed confrontation with UN forces.

Arda reasons that Arcadia has a minimum of nine months to prepare for the arrival of United Nations forces, assuming the latter experiences no delays in the equipping, training, launch, and delivery of their brigade. If they take longer, the militia can expand its capabilities as much as time permits. To console doubts, Arda further reasons that the United Nations would not assume the risk of deploying a military force across the solar system, with all of the costs and political liabilities that accompany it, unless they were confident the settlers would not be capable of defending themselves. He notes this overconfidence as their Achilles heel.

The {\it Arcadian Militia} is established on paper, but its creation does not extend beyond that. Arda assumes responsibility over the null portfolio as commander-in-chief.
\StopTimelineDate

\StartTimelineDate{3 Capricorn, 0 A.R.}
With settlement approval, Arda authorizes the use of subterranean areas of Arcadia for the garrison of several militia units yet to be created. These would include everything from drill halls and supply depots to training facilities. The units planned are few in number, but reflect the settlement's current capabilities within the minimum time remaining before the expected landing of the United Nations brigade.

\startitemize[4]
\item {\it 1\high{st} People's Own Arcadian Mechanized Infantry}. An infantry regiment to be equipped with armoured personnel carriers.

\item {\it 2\high{nd} People's Own Arcadian Dragoons}. An armoured regiment that will be capable of providing armoured reconnaissance services.

\item {\it 3\high{rd} People's Own Arcadian Combat Support Engineers}. An engineering battalion to provide heavy equipment, signals (communication), and logistical support.
\stopitemize
\StopTimelineDate

\StartTimelineDate{12 Capricorn, 0 A.R.}
Plans for a feasible battle rifle capable of operating under the extreme environmental conditions Mars presents are proposed before the {\it City Security Committee} where they are well received. The R1A1 battle rifle is a gas-operated, self-loading, air-cooled, magazine-fed, shoulder-fired, kinetic energy weapon. It is a modernized adaptation of the former Terran Coldwar era FN FAL. Everything required for its manufacture is already available at Arcadia's disposal.

The receiver and heavy barrel are made of an alloyed magnesium, though the original design called for steel which is heavier. These metal components are treated with a protective finish that matches in colour the sands of {\it Arcadia Planitia}. Even with the original weapon weighing only a third on Mars what it would on Earth, the design modifications assure the weapon is made lighter yet.

The stock is modified to use a lightweight composite hemp plastic. The material is stiff, has a high heat tolerance, flame retardant, and is biodegradable.

The weapon features both magnesium alloy sights and an ROA1 detachable optical scope for target acquisition. The front and rear sights bear small sealed tritium markings that provide a small glow to aid in night sighting. The optical scope has night vision, thermal imaging, and laser range finder capabilities.

The receiver is chambered to the already well field tested Terran 7.62x51mm round. A magnesium alloy is selected for the cartridge casing. The ejection port is modified with a cover which the original design did not have. This reduces the likelihood of very fine Martian iron oxide dust as small as 1.5μm from entering the breech and fouling the action.

The gas regulator also requires departure from the original design. This is to manage the greater gas differential of the high chamber pressure contrasted with the near vacuum of Mars at the end of the muzzle.

An RGA1 personnel grenade launcher is optionally mounted underneath and can chamber a variety of cartridges. These include magnesium flares, smoke grenades, or \quote{cryonades}. Cryonades are small grenades that contain cryogenic liquid oxygen and methane. Armourers are instructed that these latter explosive cartridges must be stored when not in use at cryogenic temperatures to prevent them from becoming inert through degassing.

Although the original design that forms the basis for a contemporary Martian adaptation is antiquated, the weapon is relatively simple to manufacture and already well proven in extreme arctic and high altitude climates. The committee notes that this is important since these climates share the closest environmental similarities with Mars. The weapon should be able to operate in temperatures as low as \math{-90^{\circ}}C. The committee suggests that any icing issues that may arise will have to be addressed later with field trials.
\StopTimelineDate

% They need a strategic transport...
\StartTimelineDate{33 Capricorn, 0 A.R.}
{\it Kali's} launch window is missed due to low-Earth orbit having become dangerously saturated with space debris\index{space debris}. UNSA engineers fear they may be approaching the Kessler effect\index{Kessler effect} of causing a runaway chain reaction which would reduce all objects in orbit. Since this would threaten not only {\it Iterum Shipyard}, but all Terran space exploration, {\it Kali's} trans-Martian injection is rescheduled for the next available launch window of 25 months.
\StopTimelineDate

\StartTimelineDate{35 Capricorn, 0 A.R.}
The first non-experimental production ready R1A1 battle rifles emerge from the assembly line. Although firearms are normally given a serial number incremented by one with the creation of each new weapon, at Arda's request, his rifle's receiver is stamped with {\it 109450}.
\StopTimelineDate

% Should occur during a perihelion season...
\StartTimelineDate{16 Aquarius, 0 A.R.}
In keeping with Senka's prediction, most of Mars, including {\it Arcadia Planitia}, is enveloped in a seasonal dust storm by this time. The day marks the planet's orbital perihelion where it is nearest to the Sun.
\StopTimelineDate

% Kali takes a free-return trajectory taking 180 days and drops off personnel and material. Should take place on 33 Capricorn, 0 A.R. + 25 months for the rescheduled launch from Earth + 180 days of transit...
\StartTimelineDate{11 Cancer, 2 A.R.}
{\it Kali} completes its trans-Martian journey, aerobreaking into areostationary orbit. The ship carries a payload of remotely operated equipment destined for Phobos, the sun-synchronous photographic reconnaissance satellite {\it Bhadra I},\footnote{In Hindu mythology, one of Shiva's servants and goddess of the hunt.} the areosynchronous communications satellite {\it Bhadra II}, MARSBRIG military personnel, as well as some civilian contractors and defence personnel. The combined human resources and equipment represent the {\it United Nations Mars Expeditionary Peacekeeping Force (UNMEPF)}.
\StopTimelineDate

% Two month after arrival in orbit to give Thor Battery time to finish construction...
\StartTimelineDate{9 Leo, 2 A.R.}
{\it Kali's} commanding officer issues warning orders down the ship's chain of command to prepare for insertion, areostationary orbit. All mechanized infantry company, platoon and section commanders are briefed before loading \goto{MADVs}[Mars Ascent-Descent Vehicle (MADV)]. Armour and artillery follow.
\StopTimelineDate

