% This is part of the Avaneya Project Crew Handbook.
% Copyright (C) 2010, 2011
%   Kshatra Corp.
% See the file License for copying conditions.

% Countdown section...
\StartSection{Countdown}

\StartTimelineDate{A.R.}
Footage of alleged riots erupting across Arcadia with security forces responding disproportionately are aired across Earth. Mainstream media attribute the alleged riots to grievances expressed by dissenters at variance with Free Arcadia partisans who apparently relied upon use of force to silence them, but now, like Arda, hold public office at Arcadia Hall. The dissenters are coined {\it Red Revolutionaries}. Several critics note the equivalence of Terran gravity in some of the footage allegedly obtained by citizens {\it in situ} but are rapidly dismissed by mainstream media as \quote{cranks} and \quote{conspiracy theorists}.
\StopTimelineDate

\StartTimelineDate{A.R.}
The United Nations Security Council passes resolution 12664 where it is adopted unanimously. The contents of the resolution are not as of yet made public, but kept internal. It begins with:

\startTimelineGeneralDocument
The Security Council,

Reaffirming the principles and purposes of the Charter of the United Nations,

Determined to combat by all means threats to transplanetary peace and security caused by terrorist acts,

Deploring the gross and systematic violation of human rights, including the repression of peaceful demonstrators, expressing deep concern at the deaths of civilians, and rejecting unequivocally the incitement to hostility and violence against the civilian population and infrastructure excited through {\it in situ} administrative elements of the United Nations Arcadian Settlement,

Recognizing the inherent right of individual or collective self-defence in accordance with the Charter,

\startitemize[n]
\setupwhitespace[big]
\item Unequivocally condemns in the strongest terms the horrifying terrorist attacks which took place recently at the site of the No. \type{#}2 Railgun on Mars in the United Nations Arcadian Settlement and regards such acts, like any act of international terrorism on Earth, as a threat to transplanetary peace and security;

\item Expresses its deepest sympathy and condolences to the victims, their families, and of the United Nations Arcadian Settlement;

\item Calls on all Mars Treaty signatories to work together urgently to bring to justice the perpetrators, organizers and sponsors of these terrorist attacks and stresses that those responsible for aiding, supporting or harbouring the perpetrators, organizers and sponsors of these acts will be held accountable;

\item Calls also on all Mars Treaty signatories to redouble their efforts to prevent and suppress terrorist acts including by increased cooperation and full implementation of the relevant anti-terrorist conventions and Security Council resolutions;

\item Expresses its readiness to take all necessary steps to respond to the terrorist attacks of the Arcadian Settlement railgun bombing, and to combat all forms of terrorism, in accordance with its responsibilities under the Charter of the United Nations;

\item Authorizes Mars Treaty signatories that have notified the Secretary-General, acting nationally or through regional organizations or arrangements, and acting in cooperation with the Secretary-General and the Office of the President of the Space Agency, to take all necessary measures to protect civilians, civilian populated areas, and assets under threat of attack in the United Nations Arcadian Settlement, and requests those signatories concerned to inform the Secretary-General immediately of the measures they take pursuant to the authorization conferred by this paragraph which shall be immediately reported to the Security Council;

\item Decides to remain actively seized of the matter.\footnote{Most Security Council resolutions end with this phrase. Article 12 of the United Nations charter states that \quote{While the Security Council is exercising in respect of any dispute or situation the functions assigned to it in the present Charter, the General Assembly shall not make any recommendation with regard to that dispute or situation unless the Security Council so requests.} In other words, the General Assembly of many is being instructed by the Security Council of few to not interfere in an affair the latter still wishes to handle exclusively.}
\stopitemize
\stopTimelineGeneralDocument

The resolution recommends and authorizes the {\it United Nations Department of Peacekeeping Operations} (UNDPKO) to place all four of its permanently standing {\it Rapid Reaction Force} battalions on high alert.
\StopTimelineDate

\StartTimelineDate{A.R.}
The United Nations Department of Peacekeeping Operations' {\it Joint Chiefs of Staff} assemble in Brussels, European Union. There they liaise with government officials and other high level military UN servants for the first day of a three month summit. The summit is conducted under maximum security. Though its existence is made public, the exercise of a high level of discretion in all of its activities ensures that its nature is not. This draws public criticism due to the presence of a number of high level state officials from across the planet and the lack of transparency.

Rumours spread rapidly over the solnet purporting the purpose of the summit as one of planning the first off world military occupation of another planet in history. Leaked documents suggest the summits agenda is to assess the feasibility, followed by addressing the preparation of all strategic, intelligence gathering, logistical, personnel, and financial planning necessary to conduct such an operation. Known high level UNSA liaisons, including UNSA's President Dr. Samuel Lieberman, are captured on camera being hustled in and out of meetings within the compound. Public speculation is further compounded when representatives of major international financial institutions and the NAU CIA director are also captured on camera by independent media.
\StopTimelineDate

\StartTimelineDate{A.R.}
UNDPKO's summit in Brussels concludes with the establishment of the newly formed {\it Mars Command (MARSCOM)}. It is established under the civilian administered UNDPKO and is now responsible for overseeing all of the latter's peacekeeping operations at the strategic level, in the route to, in the vicinity of, and on the Red Planet. The headquarters is located at Peterson Air Force Base, North American Union, and is jointly staffed by an international consortium of military personnel from all nations signatory to the Mars Treaty.
\StopTimelineDate

\StartTimelineDate{A.R.}
Lockheed Martin-Boeing's Skunk Works advanced research facilities begin refitting UNSA's {\it Tarikin III} personnel ferry into the {\it Kali}, a militarized, remotely piloted, spacecraft carrying a payload of personnel and materials. This is in response to a UNSA issued contract for a classified mission payload specification requested by MARSCOM.
\StopTimelineDate

\StartTimelineDate{A.R.}
Selected Rapid Reaction Force personnel are assigned to UNSA training facilities in Antarctica and Huelva, Spain, now administered by MARSCOM as two battalions MARSBAT1 and MARSBAT2.
\StopTimelineDate

\StartTimelineDate{A.R.}
Arda receives an asymmetrically encrypted message warning of an allegedly imminent plan, drafted by a consortium of Terran financial interests intending to use the United Nations as their delivery platform. The plan apparently would call for the organization and landing of two battalion of peacekeepers to Martian theatre, operating under UNSA's newly formed MARSCOM, which itself operates as an organ of the UN, which allegedly answers to the Terran central banking cartels.

The message claims the single Rapid Reaction Force battalions MARSBAT1 and MARSBAT2 provided by the UNDPKO under MARSCOM command are tasked with the following prioritized strategic objectives leaked from source material:

\startTimelineCorrespondenceDocument
TOP SECRET
DISTRIBUTION RESTRICTED / FOR NAMED RECIPIENTS ONLY
MARSCOM JOINT CHIEFS OF STAFF
{\color[colour_item] MEMORANDUM CONCERNING ARCADIAN SETTLEMENT STRATEGIC OBJECTIVES}
\startitemize[4]
%\setupwhitespace[big]
\item {\color[colour_item] ALPHA} - KILL OR CAPTURE ATATÜRK\footnote{The NAU CIA attaché under MARSCOM uses this cryptonym for Arda.}

    \startitemize[n]
    \item IF CAPTURED, HOLD IN CUSTODY FOR IMMEDIATE EXTRACTION TO KALI, AREOSTATIONARY ORBIT. WILL AWAIT EXTRAJUDICIAL RENDITION BACK TO EARTH.
    \item IF KILLED, IMPLICATE 'RED REVOLUTIONARIES'. MEDIA ALREADY BRIEFED AND WILL STANDBY. BHADRA 2 COMMSAT WILL BROADCAST NEWS BACK TO EARTH.
    \stopitemize

\item {\color[colour_item] BRAVO} - CAPTURE ARCADIA
    \startitemize[n]
    \item INSERT PLAINCLOTHES NAU CIA, SPECIAL FORCES, AND BLACKWATER ASSETS INTO CITY. WILL GATHER INTEL.
    \item SEIZE CONTROL OF MAX NO OF MATERIAL ASSETS INTACT.
    \item MAINTAIN STARPORT ARCADIA OPERATIONAL.
    \item REINSTATEMENT OF UNSA ADMIN BROKERED THROUGH MARSCOM'S INITIAL COMMAND AND CONTROL.
    \item DISARM SIERRA\footnote{MARSCOM phonetic alphabet designation for a settler.} THROUGH CONFISCATION OF ALL SMALL ARMS.
    \item NEUTRALIZE ALL LOCAL RESISTANCE, W/ NO QUARTER POSSIBLE. EXPECT MINIMAL, IF ANY, FIBA.\footnote{Fighting in built up areas.} 
    \item ENHANCED INTERROGATION OF ALL SUSPECTED FAP\footnote{Free Arcadia partisans.} CAPTURED. EXTRAJUDICIAL DETENTION, TRIAL, AND EXECUTION. ATATÜRK EXCEPTION, VIZ STRATEGIC OBJECTIVE ALPHA.
    \stopitemize

\item {\color[colour_item] CHARLIE} - ASSET EXTRACTION
    \startitemize[n]
    \item EXTRACT ALL NAU CIA ASSETS STILL OPERATING IN THEATRE.
    \stopitemize

\item {\color[colour_item] DELTA} - ECONOMIC REFORMS
    \startitemize[n]
    \item REINSTATEMENT OF SETTLEMENT LABOUR AND KEY PERSONNEL NECESSARY FOR RESUMPTION OF STABLE DEUTERIUM AND OTHER KEY EXPORTS.
    \item INSTALLATION OF 'FIRST CENTRAL BANK OF THE UNITED NATIONS ARCADIAN SETTLEMENT'.
    \item REVERSION OF ARCADIAN JENYA BACK TO TERRAN BANCOR.
    \item CAPTURE OF ALL RHODIUM BACKED JENYAS HELD IN RESERVE.
    \stopitemize

\item {\color[colour_item] ECHO} - POLITICAL REFORMS
    \startitemize[n]
    \item IMPLEMENTATION OF MARTIAL LAW W/ SUFFICIENT TIME TO ACHIEVE ALL STRATEGIC OBJECTIVES.
    \item ENGINEER SETTLEMENT ELECTION WITH SUCCESSION OF ARCADIA HALL'S INCUMBENT OOP\footnote{Office of the President.} W/ SYMPATHETIC SUCCESSOR TO WHITE LISTED TERRAN INDUSTRIAL AND FINANCIAL INTERESTS.
    \stopitemize
\stopitemize

IN UNLIKELY EVENT MARSBAT1 / MARSBAT2 UNABLE TO HOLD ARCADIA, EXECUTE EMERGENCY CONTINGENCY PLAN OF OBJECTIVE FOXTROT:
\startitemize[1]
\item {\color[colour_item] FOXTROT} - ASSET INSERTION AND OPERATION

    \startitemize[n]
    \item ALL MARSCOM ASSETS, EXCEPTION FOXTROT(TWO) FALLBACK TO RENDEZVOUS TANGO. AWAIT EXTRACTION TO KALI, AREOSTATIONARY ORBIT. FURTHER ORDERS WILL FOLLOW.

    \item IF OBJECTIVE BRAVO(ONE)'S ASSETS INSERTED SUCCESSFULLY, THESE ASSETS REMAIN BEHIND IN SIERRA SAFE HOUSES AS 'FIFTH COLUMN' CONTRAS. FIFTH COLUMN WILL PROVIDE PROTECTION FOR RETREATING MARSCOM ASSETS IF POSSIBLE.

    \item FIFTH COLUMN CONTRAS WILL CONDUCT INSURGENCY AND EXECUTE FALSE FLAGS MASQUERADING AS 'RED REVOLUTIONARIES'. CAN OPERATE INDEPENDENTLY AND CONDUCT OPERATIONS W/O NEED FOR EXTENDED COMMUNICATION CONTACT IF BROKEN OFF. PROVIDES MARSCOM W/ CLANDESTINE UNIT BEHIND ARCADIAN DEMARCATION LINE FOLLOWING STRATEGIC RETREAT.

    \item FIFTH COLUMN MUST RAISE VITAL FUNDS INDEPENDENTLY IN LOCAL JENYA CURRENCY IF OBJECTIVE DELTA(THREE) FAILED IN ORDER TO SUPPORT ACTIVITIES. SIERRA'S SEIZED LABORATORY FACILITIES POSSIBLY CONDUCIVE TO {\it IN SITU} SYNTHETIC PRODUCTION NECESSARY FOR NARCOTRAFFICKING. PROVIDES PLAUSIBLE DENIABILITY WHILE DENYING NEED FOR COSTLY TERRAN IMPORT.
    \stopitemize
\stopitemize

\stopTimelineCorrespondenceDocument

The anonymous tip-off contains detailed information on troop movement to Terran MARSCOM administered training facilities, size of units, weapons and tactics to be employed, probable locations of encampment, and route of approach. UNSA allegedly intends to transport the battalion under the handle of the {\it United Nations Emergency Assistance Peacekeeping Force (UNEAPF)}. Its numbers are drawn principally from North American Union and European Union airborne light infantry units with armoured reconnaissance and self-propelled artillery support.

The message is encrypted using a public key of Arda's not used for many years since his long departed days of a military posting with the Mountain and Commando Brigade in Hakkari, European Union. Arda and his staff deem the message credible.
\StopTimelineDate

\StartTimelineDate{A.R.}
Arda directs the systems team to take out UNSA orbital surveillance capabilities by destroying the {\it Mars Enhanced Telecommunications Orbiter} while they still have access. UNSA transmits a command sequence to the orbiter to revoke Arcadia's keypair but is eight minutes too late. The orbiter burns up its entire attitude control hydrazine fuel reserves, causing it to be incinerated during atmospheric entry after deorbiting itself.

UNSA has the NAUSA JPL lock down all four {\it Mars Positioning Satellites} in anticipation of RAP denial-of-service attacks.

All solnet communications between Earth and Mars that had been brokered by METO are now effectively disabled. The orbiter had also been used for reliable communication between Arcadians. The communications team adapt in its absence with a combination of line-of-sight UHF and shortwave communication, requiring no satellite.
\StopTimelineDate

\StartTimelineDate{A.R.}
The Security Council makes resolution 12664 and the fact of its unanimous adoption public, but stalls until today. The timing coincides with the saturation of key Terran broadcast networks airing a recent adultery scandal of a prominent Terran celebrity.
\StopTimelineDate

% Need a strategic transport...
\StartTimelineDate{A.R.}
{\it Kali's} launch window is missed due to the low-Earth orbit having become overly saturated with space debris. UNSA engineers fear they may be approaching the Kessler effect of causing a runaway chain reaction which would reduce all objects in orbit. Since this would threaten not only the {\it Iterum Shipyard}, but all Terran space exploration, its trans-Martian injection must be rescheduled for the next available launch window of 25 months.
\StopTimelineDate

\StartTimelineDate{A.R.}
Arda holds an emergency cabinet meeting at {\it Arcadia Hall} without issuing a public statement. A team of geologists and engineers are attached to a survey team and dispatched immediately on rovers at his directive.
\StopTimelineDate

% Expediency is plausible if partisan meetings had been held at such a location were surveying had already been done in the past...
\StartTimelineDate{A.R.}
Arcadia Hall issues an Arcadia-wide emergency evacuation order that is carried out over the next two weeks. All non-essential personnel are transported to a location within an unmarked subterranean lava tube installation.
\StopTimelineDate

\StartTimelineDate{A.R.}
{\it ERV Bravo} carrying Kissinger lands at Edwards Air Force Base, California, North American Union. The spacecraft is immediately seized as a UNSA asset. Kissinger leaves for Geneva the following day.
\StopTimelineDate

% Kali takes a free-return trajectory taking 180 days and drops off personnel and material...
\StartTimelineDate{A.R.}
{\it Kali} completes its trans-Martian journey, aerobreaking into Martian areostationary orbit. The ship carries a payload of remotely operated equipment destined for Phobos, the photographic reconnaissance satellite {\it Bhadra I}\footnote{In Hindu mythology, one of Shiva's servants and goddess of the hunt.} for sun-synchronous, the communications satellite {\it Bhadra II} for areosynchronous orbit, as well as two United Nations Rapid Reaction Force battalions.
\StopTimelineDate

\StartTimelineDate{A.R.}
{\it Kali's} commanding officer Lieutenant-Colonel Dragov issues warning orders down the chain of command to prepare for insertion while in Martian areostationary orbit.
\StopTimelineDate

