% This is part of the Avaneya Project Crew Handbook.
% Copyright (C) 2010, 2011, 2012
%   Kshatra Corp.
% See the file License for copying conditions.


% Rubicon Act's notable portions summarized...
\startarticle
\item % Article I

...declares the colony an independent, sovereign, self-governed, constitutional republic with the right to self determination. It is self-styled the {\it Republic of Arcadia Planitia (RAP)}, colloquially known by its capital and sole city, {\it Arcadia}. The United Nation claims this area falls under its administration and knows it only as the Arcadian Settlement.

\item % Article II

...defines the head of state, the President. The President and his secretaries are democratically elected and serve for as long as the electorate permit. Since Arcadia is only a single city at this time, there is no large legislative assembly defined akin to a congress of sorts as sometimes found in the governance of nations greater than a single city.

\item % Article III

...discharges all of Arcadia's public debt held by the {\it International Monetary Fund} and the {\it World Bank}, ceasing interest payments effective immediately. 

The Terran central banks had been using the settlers' registered biological property, that is, their UN issued extraterritorial birth certificates, and the security of the bearer's labour as monetized collateral against a debt that preceded the settlement's very existence. This had been done without the bearer's knowledge.

\item % Article IV

...defines an independent judiciary whose judges are democratically elected. They serve for as long as the electorate permit.
\stoparticle

\startarticle[start=7]
\item % Article VII

...superannuates the Terran \goto{\it bancor}[Bancor] fiat currency with the \goto{\it jenya}[Jenya], a \goto{\it rhodium}[Rhodium] standard. The rhodium currency is composed of a platinum metal considered more rare and precious than gold. It is declared the exclusive legal tender within the Republic, necessitating all Terran interests acquire Arcadian goods and services exclusively in jenyas.

\item % Article VIII

...defines the state's disposition to personal arms and a military. It states that the right of the people to bear arms must never be infringed. The underlying idea is that this is done principally for the protection of the people from their own government, or that of another's.

The creation of a permanent standing military force at the government's disposal is strictly forbidden. The raising of militias is permitted, but conditional upon executive authority vested exclusively in none other than a consensus of the people, not of government. Some of the militia's restrictions are that it is forbidden from deployment outside of Arcadia's territorial regions, that its purpose is exclusively defensive -- of people, not of government or corporations, that its commissioned officers be Arcadians, that membership is voluntary, and that it cannot be deployed domestically as an aid to civil power under any circumstance.

\item % Article IX

...denies the deployment of munitions of war, suborbital, orbital, in outer space, or anywhere outside of Arcadia's territorial regions.
\stoparticle

\startarticle[start=11]
\item % Articles XI to XXVI

...enumerate the fundamental freedoms and human rights Arcadians enjoy with respect to their new state. They are adaptations from Earth's former {\it European Convention on Human Rights} drafted in 1950 A.D.. These articles include the

    \startitemize[R,2*broad][start=11]
    \setupitemize[left=(, right=)]
    \item[item:rubicon_act_fundamental_rights_first] right of everyone to their life;
    \item right of peaceful enjoyment of one's possessions;
    \item prohibition of torture;
    \item prohibition of slavery;
    \item right of liberty and security;
    \item right to a fair trial;
    \item prohibition of retrospective criminalization of acts;
    \item right of one's privacy;
    \item right of freedom of thought, conscience, and religion;
    \item right of freedom of expression;
    \item right of freedom of assembly;
    \item right of any of marriageable age to marry and establish a family;
    \item prohibition of discrimination;
    \item right to not be denied an education;
%    \item right to regular, free, and fair elections;
    \item[item:rubicon_act_fundamental_rights_last] prohibition of the death penalty.
    \stopitemize

\stopitemize

\startarticle[start=29]
\item % Article XXIX

...prohibits the state from providing itself with a not-withstanding clause, preventing a legal override of any part of this Act.
\stoparticle

