% This is part of the Avaneya Project Crew Handbook.
% Copyright (C) 2010-2013 Cartesian Theatre <kip@thevertigo.com>.
% See the file Copying for details on copying conditions.

% Documentation section...
\StartSection{Documentation}

\startitemize[4]
\head {\em \BIBTEX}

As you have probably noticed, this book has many references. \BIBTEX\ is a well proven tool that has been used for many years in various fields ranging from the humanities to the sciences. It keeps track of references, organizes them, and presents a standardized interface for typesetting software like \CONTEXT.

\head {\em \CONTEXT}

We use \CONTEXT\ for typesetting important documents, such as this one. This document was compiled on \currentdate\ from Bzr revision \BzrRevisionClickable, typeset using \CONTEXT\ \contextversion, and rendered with \texenginename\ \texengineversion. One of the many beauties of using a programmable typesetting language like \CONTEXT\ is that documents written in them are {\tt diff} and {\tt patch} friendly. This encourages collaboration by making it easier to do.

\head {\em Doxygen}

As software increases in its complexity, so does the necessity to document its internal workings. Doxygen is a tool that can generate \index{engineering documentation}engineering documentation on the fly. The documentation is derived from metadata hints at the source code level. It can even produce graphs that show the relationship between classes. This latter feature is actually implemented internally using another tool we depend on, Graphviz.

\head {\em Graphviz}

Graphviz is a powerful set of tools for drawing graphs specified in the DOT language. Among the suite it provides, {\tt dot} was used for several of our figures, like that in \in{figure}[figure:Resources_Pipeline].

\head {\em Umbrello}

Umbrello is a powerful diagramming tool which lends itself well in the creation of \index{UML models}UML models. The AresEngine's architectural design was prepared in Umbrello, as seen in \in{section}[AresEngine's Architecture]. The two projects compliment each other well since modelling the AresEngine put very heavy demands on Umbrello, thus allowing the stress testing to identify, solve, and enhance many aspects of Umbrello. We are very grateful to the cooperative and friendly developers working on the Umbrello project. The two projects have historically had a productive relationship.
\stopitemize

