% This is part of the Avaneya Project Crew Handbook.
% Copyright (C) 2010, 2011, 2012 Cartesian Theatre <kip@thevertigo.com>.
% See the file Copying for details on copying conditions.

% Selected Technologies chapter...
\StartChapter{Selected Technologies}

It is generally a bad engineering practise to try to re--invent the wheel when there are existing tools and technologies that can be reasonably leveraged. We will list by category in this chapter all of these technologies we selected for use with this project. Sometimes we have explained each of them in the words of their maintainers, sometimes in our own to better justify the rationale behind their selection.

We always try to stick to using software {\it libre} whenever possible. The only time we might use a non--free tool is in using it as a model in designing a suitable replacement. Another time would be where writing a replacement is far too difficult and we had no other choice.

Note that not all of these technologies are components within the AresEngine, such as Audacity. It does not do anything at runtime, but we can consider it as a valuable \quote{offline} tool since it is necessary in preparing game media for the runtime environment.

\input Selected_Technologies/Audio.tex
\input Selected_Technologies/Collaboration.tex
\input Selected_Technologies/Developer_Tools.tex
\input Selected_Technologies/Documentation.tex
\input Selected_Technologies/Graphics.tex
\input Selected_Technologies/i18n_and_L10n.tex
\input Selected_Technologies/Input.tex
\input Selected_Technologies/Networking.tex
\input Selected_Technologies/Physics.tex
\input Selected_Technologies/Platform_Conveniences.tex
\input Selected_Technologies/Scripting.tex

\StopChapter

