% This is part of the Avaneya Project Crew Handbook.
% Copyright (C) 2010-2015 Cartesian Theatre <info@cartesiantheatre.com>.
% See the file Copying for details on copying conditions.

% Game Mechanics chapter...


\StartSection{Economy}
A cities economy is a diverse and complex aspect of gameplay in Avaneya yet however complex the implementation of the economy becomes the player's use of it should be fluent and intuitive, allowing for varying degrees of complexity and involvement depending on a players style of play. This can be done with a good user interface, a solid understanding of economic principles, and clever AI. 

This section deals with the variety of ways cities earn and spend money and maps out the paths that spending makes. To begin with we will discuss income streams for a players cities. These are primarily made up of taxation, exports (and imports), and city resources and infrastructure. Within each of these categories are several methods which players can use to generate income.

%Figure showing flow of city expense and income
%Overall_Economy.gv
The above figure shows the average city's economy. In it, income leaves and returns in various methods but note how income is also generated from some of the cities expenses. Expenses can be seen of as investments varying in lengths and amounts of return. Zoning for instance is a relatively cheap investment at city expense which has further investment from private capital to generate a form of growth (prosperity) which can be taxed or processed through fees to generate city income. This simple investment can be made more complex through adding infrastructure which may increase desirability and efficiency and perhaps allow for higher fees or taxes. Other forms of investment, such as funding, divert money into projects the player or city desires allowing for complex growth along specific avenues.



\StartSection{Income Streams}

City income is generated in 6 ways.

1.  Taxation
2.  Fees and ordinance
3.  Imports and Exports
4.  Commodities
5.  Land
6.  Goods and services

The forms of income used is completely up to the players. They can use as many or a little of these as they wish.



\StartSection{Expenditure}

1.  Infrastucture and City Buildings
2.  Funding
3.  Ordinance
4.  Zoning
5.  Militia
6.  Goods and Services


Just like city income, the avenues players choose to spend their city's money is entirely up to them. Different goals and situations require different strategies.


Beginning with income streams, lets start with the primary method of generating income, taxation.
%cue taxation.tex.



%DEMAND CURVES CREATE AND BREAK BUSINESSES






\StartSection{Principles}
%List and explaination of major economic priciples needed to be applied
This section notes the foundational economic princples the games requires to model economics innovatively. 
\StartSubsection{Effects of taxation on goods and businesses}
\StartSubsection{Effects of levies on imports and exports}
\StartSubsection{Elasticity}
\StartSubsection{Producer and Consumer Surplus}
\StartSubsection{Regulation and Correction

\StartSection{Behavioural Economics}%this section will most likely end up in workforce.
%Workforce interaction





