% This is part of the Avaneya Project Crew Handbook.
% Copyright (C) 2010, 2011, 2012
%   Kshatra Corp.
% See the file License for copying conditions.

% Characters chapter...
\StartChapter{Characters}

Avaneya's characters are drawn from a variety of places on Earth. This is to be expected of any space program with an international flavour. Nearly all of those deployed to Mars have strong scientific backgrounds in their respective fields. Their theoretical knowledge is further supplemented with proven practical experience.

All of the characters' idiosyncrasies can and should be revealed indirectly whenever possible. This may be through body language, object symbolism, or perhaps the circumstance of the events they participate in. Whenever more direct means are appropriate, a character's own writings or their participation in a dialogue are possibilities as well.

An introductory cinematic sequence is ideal for presenting the audience with the characters, though it may only have time to cover the most important ones. The leading figure and protagonist, Arda Baştürk, ought to be the game's most recognizable character. He is described further in \in{section}[Arda Baştürk].

% Mission Commander...
\page
\input Source/Characters/Arda_Baştürk.tex

% Chief Medical Officer, Chief Field Science... (Experimental Team Lead)
\page
\input Source/Characters/Henrik_Nørgaard.tex

% Systems Team Lead...
\page
\input Source/Characters/Khalid_Zafar.tex

% United Nations Deputy Secretary-General...
\page
\input Source/Characters/Leonard_Kissinger.tex

% Greenhouse Team Lead, Terraformation Team Lead...
\page
\input Source/Characters/Nayana_Rai.tex

% Chief Flight Engineer, Mechanics Team Lead, Survey Team Lead...
\page
\input Source/Characters/Senka_Rukavina.tex

\StopChapter

