% This is part of the Avaneya Project Crew Handbook.
% Copyright (C) 2010, 2011, 2012
%   Kshatra Corp.
% See the file License for copying conditions.

% Henrik Nørgaard section...
\StartSection{Henrik Nørgaard}
Henrik was born 49 Aries, 32 B.R., in Copenhagen, European Union. He attended Lund University\index{Lund University} where he completed his undergraduate education in cellular microbiology. Following that, he went on to complete a Masters in Molecular Biology and Microbiology. After completing his doctoral studies, he pursued research opportunities with the Max Planck Institute for Medical Research in Heidelberg\index{Max Planck Institute+for Medical Research}, European Union. 

While in Heidelberg, Henrik's principle area of investigation was in the field of orthomolecular medicine\index{Orthomolecular medicine}.\footnote{Orthomolecular medicine is a branch of medicine concerning itself with the relation of nutrition and illness, especially using the former in the treatment of the latter.} During his time with the institute, Henrik had more than 80 peer reviewed papers published on topics ranging from clinical trials involving the use of traditional herbal remedies, the biochemistry of fermented foods, and the efficient recycling of biomass.

After several years with the institute, Henrik went on to acquire his status as a naturopathic doctor at the University of Copenhagen while lecturing part time at the university. As a lecturer, he was noted for his satirical sense of humour, brilliance, and an extensive set of mushrooms and herbs from virtually every corner of the Earth in his office.

Following his accreditation as a naturopath, Henrik went on to open a medical clinic in Copenhagen. Shortly after his acceptance of an offer from UNSA to venture to Mars as part of the Avaneya Initiative, Henrik passed on care of the clinic under the supervision of a close colleague.

