% This is part of the Avaneya Project Crew Handbook.
% Copyright (C) 2010, 2011, 2012
%   Kshatra Corp.
% See the file License for copying conditions.

% Leonard Kissinger section...
\StartSection{Leonard Kissinger}
Leonard was born 8 Gemini, 37 B.R., New York, in what was then the United States prior to its unification in 30 B.R..

Leonard attended the University of Oxford\index{University of Oxford} where he majored in Politics, Philosophy, and Economics at Christchurch College. While at Oxford, Leonard was a member of the Bullingdon Club\index{Bullingdon Club}. He completed his undergraduate degree three years later, graduating with a first class Bachelor of Arts. Following that, he pursued graduate studies at Princeton University's Woodrow Wilson School\index{Woodrow Wilson School} with a focus on public policy analysis and theory of law. After completing his studies, he accepted a teaching position with Harvard.

By 19 B.R., he had been a participant in a number of policy groups. These included membership with the Trilaterial Commission\index{Trilateral Commission}, the Aspen Institute\index{Aspen Institute}, and holder of the title of President of the Council of Foreign Relations\index{Council of Foreign Relations}. His public persona increased in recognition as he appeared regularly over the years at the World Economic Forum\index{World Economic Forum} as a guest speaker. He had been involved with the RANND Corporation\index{RANND Corporation} as an external consultant throughout his time on Earth as well.

In 18 B.R., while maintaining his teaching position at Harvard, Leonard accepted a position as a member of the board of directors at Chase National Bank\index{Chase National Bank}. He maintained the position up to and following his entry into politics.

Leonard entered the North American Union's\index{North American Union} Department of State as a diplomat in 15 B.R.. He gained recognition in having been instrumental in applying diplomatic pressure on what was then Japan and South Korea to sign the Tokyo Partnership Accord\index{Tokyo Partnership Accord}. When Japan and South Korea finally became signatories, a nearly unified far eastern Asian Union\index{Asian Union} was created.

Two years later he became foreign policy adviser to the Secretary of State under the President of the North American Union\index{North American Union}, though maintaining many of the functions of a diplomat. His responsibilities included the formalization of relations between the North American Union and the newly created Asian Union. He has also been cited for his involvement in improving bilateral relations between the Asian Pacific\index{Asian Pacific Union} and Central American Unions\index{Central American Union} by promoting cooperation between their respective central banks. 

Leonard's political philosophy has been described as adhering to the {\it Realpolitik}\index{Realpolitik} school of thought. The approach is one in which politics and diplomacy do not primarily concern themselves with ideological or ethical concerns, but on use of force to achieve objectives. The term was first used in the 19\high{th} century by German writer and politician Ludwig von Rochau\index{Ludwig von Rochau}.

Leonard left his position as a foreign policy adviser to the Secretary of State in order to accept an appointment to the Office of the United Nations Deputy Secretary-General on 19 Scorpio, 11 B.R.. As part of his duties he was required to undergo a three week basic preparatory course for non-essential personnel in Antarctica to prepare him for accompanying the Avaneya crew to Mars.

Leonard has always exercised care in minimizing the publicity of his personal life. However, it is well established that he has always been a strong patron of the arts. Leonard was a Trustee of the Museum of Modern Art\index{Museum of Modern Art} in New York City since before his entry into politics. In addition, he was a noted financial contributor to a number of cancer related charity groups which included the Children's Last Wish Fund\index{Children's Last Wish Fund}. Rumours have surfaced in the past that he is said to own significant stock in North American Union-based asset management Carlyle Holdings\index{Carlyle Holdings}, along with several pharmaceutical giants.

Leonard is a descendant of the controversial Dr Henry Kissinger\index{Henry Kissinger}. The latter was a noted diplomat and businessman, highly influential in the political circles of his time.

