% This is part of the Avaneya Project Crew Handbook.
% Copyright (C) 2010-2013 Cartesian Theatre <info@cartesiantheatre.com>.
% See the file Copying for details on copying conditions.

% Licensing Rationale chapter...
\StartChapter{Licensing Rationale}

Avaneya, as a piece of software, is composed of different types of files that fall into different categories. The game is licensed under multiple licenses, with a given license for each category. These categories are for {\it artwork}, {\it code}, {\it documentation}, {\it music}, and {\it typography}. Where there is a discrepancy between the following descriptions and the contents of the {\tt Copying} file located within the project source root, the latter prevails.

\startitemize[4]
\head {\em Artwork}

This consists of all creative literary works, including this handbook, as well as any other fragments, cinematics, models, audio and voice effects, textures, materials, concept art, screenshots, and other relevant non--executable creative media.

These objects are released under the terms of the Creative Commons Attribution--NonCommercial--ShareAlike (BY--NC--SA) 3.0 Unported license, copyright \copyright\ \CopyrightDates\ \CopyrightHolder. See \in{section}[Creative Commons Attribution-NonCommercial-ShareAlike] for the full text of the license.

This license ensures that, under the default conditions and unless given permission to do otherwise, these works cannot be used for commercial purposes (NC). They must give attribution (BY), and if they alter, transform, or build upon it, they must also distribute the resulting work genetically by doing so only under the same or similar license (SA). Everyone is free to copy, distribute, and transmit this creative work, as well as to adapt it as they see fit. 

Keep in mind that these are just the {\it default} copyright conditions and can be waived at any time with the permission of the copyright holder. We felt that art does not have to be free, but should at least be shareable. This is the position Richard Stallman has taken, along with many other artists.

The Viking lander mission data archived by the Science Digital Data Preservation Task, Jet Propulsion Laboratory, was provided courtesy of the NASA Planetary Data System and was released into the public domain.

\head {\em Code}

The AresEngine, shaders, Lua scripts, build environment scripts, and any other relevant executable code, are released under the terms of the GNU General Public License 3.0 or later, copyright \copyright\ \CopyrightDates\ \CopyrightHolder. See \in{section}[GNU General Public License] for the full text of this license.

The Solnet server, build environment scripts, and any other relevant executable data are released under the terms of the GNU Affero General Public License, copyright \copyright\ \CopyrightDates\ \CopyrightHolder. See \in{section}[GNU Affero General Public License] for the full text of the license.

In a nutshell, these two licenses protect a users four fundamental freedoms. Sometimes these are colloquially referred to as the \quote{Four Pillars / Tenets of Free Software}. These are firstly the freedom to use the software for any purpose; secondly, the freedom to adapt the software to suit their needs; thirdly, the freedom to share the software with their friends and neighbours; and fourthly, the freedom to share their adaptations with others. 

The license for most software and other practical works are designed to take away your freedom to share and change the works. By contrast, these licenses are intended to guarantee your freedom to share and change all versions of a program and to make sure that it remains free for all of its users.

\head {\em Documentation}

UML and other design schematics, technical documents, doxygen output, man pages, and other relevant data are copyright \copyright\ \CopyrightDates\ \CopyrightHolder\ under the terms of the GNU Free Documentation License 1.3. See \in{section}[GNU Free Documentation License] for the full text of the license.

The Free Software Foundation explained the purpose of this license as being to make a manual, textbook, or other functional and useful document \quote{free} in the sense of freedom. This is to assure everyone the effective freedom to copy and redistribute it, with or without modifying it, either commercially or non--commercially. Secondarily, this license preserves for the author and publisher a way to get credit for their work, while not being considered responsible for modifications made by others.

\head {\em Music}

This includes all music in Ogg Vorbis, FLAC, Speex, or other formats and associated project files where provided.

These objects are released under the terms of their respective artists or publishers. We felt that music does not have to be free, but should at least be shareable. This is the position Richard Stallman has taken, along with many other musicians.

\head {\em Typography}

This includes the Avaneya Font Family typeface and all associated source files. This typeface is under development.

These objects are released under the terms of the SIL Open Font License version 1.1, copyright \copyright\ \CopyrightDates\ \CopyrightHolder. See \in{section}[SIL Open Font License] for the full text of the license.
\stopitemize

\StopChapter

