
\StartSubSection{Funding}

Funding designates city money to avenues chosen by the player or the cities population. Efficient use of this funding depends on a variety of factors through micro management. Funding can be directed to institutions such as universities or government research labs, or directed to commerical interests such as businesses or industry, or to city infrastructure or projects of similar vain. Funding allows for improvements in specific, directed areas of the city or cities development. It keeps students educated (or uneducated), transport network working (or falling apart), law inforcement honest (or corrupt), among others. It also allows for technological developments unique to a city or keeps a cities technological innovativeness in line with other cities. Funds can be spent wisely or unwisely and depending on a cities style, goal, or projected growth, certain avenues of funding become more or less important.

The management of a city requires vast amounts of funding and funding avenues. Funding options/avenues may appear or disappear as play progresses (this should be displayed or noted to a player as they happen) or as ordinance are selected. However most funding options will become permanent once implements. Not all cities will have a hospital or roads, but most will and once in place may remain for the length of the city. Therefore the funding screen is dynamic and changes with city buildings, infrastructure, and enacted ordinance.

Like most things in Avaneya, funding can be micro managed to a level of detail the player is comfortable with or prefers. Players can fund all health care, or individual hospitals / doctors, or even aspects of the health care they think is required most, such as endocrinology research or ambulance upgrades. Funding can also be limited to certain areas of the city, such as a percent of infrastructure funding being spent on the East Exit Vacuum Tunnel Transport Network.

Funding is mostly draining on city coffers however some funding can generate income for the city. An example may be funding infrastructure to be up to date with neighboring transport nextworks all for businesses to export more goods allowing the player to charge more fees and/or gain more in tax from the business. Another example is funding a thorough education, helping build a solid workforce, and reducing unemployment creating a wealth not only in taxes, but in innovation, competitiveness, and therefore a stronger economy in the long run with more wealth. Funding could also be a more brutal means of raising revenue by quelling unrest and forcing people to pay taxes at rates deemed unfair.



