\StartSubSection{Exports and Imports}
Through export and import levies, fees, charges, or taxation, income can be gained for the city from its businesses and industries exporting goods or from other cities importing goods to the players city. Additionally, exports increase the income of a city by bringing in wealth from other cities in exhange for the player's city's goods. Trade in general may increase the income of a city through established trade routes, larger markets, and peaceful diplomacy backed by financial gains for both parties. 

Charges placed on exports and imports can also be micro managed to focus on the type of goods being ported right down to individual goods themselves by brand, quantity, and quality. Cities can even ban the export or import of goods. This may become necessary in food shortages, administering certain city ordinances, or to appease interest groups (both internal and external). 

The control of exports and imports through fees is another method of controlling workforce conditions by allowing or limiting access to certain goods and setting the price at which these goods can be obtained.

Enforcing the import and export of goods may require enforcement officers, or similar and strict control against certain opinions/influences may result in varying aspects of criminal workforce. And importing and exporting requires certain building types, depending on the goods being transported and the quality measures desired. 

%From what screen do players setup exports / imports?
Imports and exports are accessed via the (TRADE?) screen. This screen shows all trade with other cities and with Earth, and can be broken down into various analyses. This screen will initially show sufficient detail but players can click on any import or export to show greater detail and edit it. 

Depending on the method of play, players may not instigate trade personally, this will be done by the cities business'. On cities ran with a more centralised approach, the player may instigate the trade of goods, however players may do this at anytime. In the trade screen, by clicking (OPEN_TRADE?) the player can seek to trade with any city (or Earth) where diplomatic conditions are appropriate to facilitate. Cities where diplomacy has been established but the underlying conditions to start trade do not exist will also be shown (at the bottom) but will be grey out.  

%compatibly transports and docking are required for trade.
%Import/Export Screen.
%Fees for imports / exports screens.
%Market backed diplomacy.
%Effects of levies on goods prices, diplomacy, and foreign policy.

