\StartSubSection{Land}
%\StartSubSection{City Land} %work in progress, may be scrapped as conflicts with zoning. may expand zoning to include land being bought by other cities.
Players are designated an area of land to be owned by them, for use as they wish, their owned land can expand or contract depending on a variety of circumstances. 

Land ownded by a players city can be leased or sold to various interests with various conditions. In this manner a player can generate income from leasing or selling its cities land. This land may be bought by a domestic business interest, terran business or government interest, or by other cities and players. The price of land may be dictated by players but its intrinsic value in dependant on its ordinance, local zoning, resource value, or growth potential. %(whatever growth potential is)
%legal action re changing or interfering with land bought by external interest. does there need to be a "trust meter" or similar? 
A potential problem for the city is that this land's use is then largely outside of city control and potential disasters or problems may not be solvable, leaving areas of disruption within the cities limits. In addition, changing city ordinance or conditions which effect the leased or bought land may bring legal action to the players city in the form of a fine or make it less trusted to do business in, hurting investment and local businesses and therefore the city's economy.

%Break up land using the (TOOL_NAME), same as that mention in th taxation screen.
This land can be selected through the (SCREEN/TOOL). From here the land can be sold, auctioned, or leased under conditions stipulated by the player. These conditions can also be micromanaged.
