
\StartSubSection{Ordinance}

City ordinance consists of projects, initiatives, or requirements a city sets in place for itself and other bodies within the city or relating to the city. Examples of ordinance may be an organic agriculture innitiative, free higher education, or polution free zones. Ordinances can be broad or specific and can be designated to certain areas of a city or apply to the city as a whole, they can also target certain groups or bodies of a city such as workforce types or income tpye/levels. An ordinance gives players more control over what is produced in the city, where it is produced, and how it is produced. It can encourage or discourage certain lifestyles, areas of development, population growth, and many other things as well as create balance or unbalace or various factors.

Therefore enacted ordinance can take and give money to specific groups for specific interests. Ordinace also allows players to implement or attempt to implement city laws. Ordinance can generate income if used correctly but this is usually through encouraging growth in certain economic sectors or encouraging movement areas the players thinks may be lucrative for the city. Costs can be initial and/or ongoing due to it being implemented by government agents. Some ordinance will have initial/outlay or specifically defined costs that cannot be modified, others can have costs varied and can be given more or less funds through the funding screen. 


Types of ordinance available depend on workforce types, funding levels, currently established interets, agents in play, and other ordinance that are active. The actual list of ordinance available is quite large and, as mentioned, is effected by a variety of factors so every ordinance is not available to all players. While there will be a large amount available to every player certain ordinance will require specific conditions to be met to become available. When enacting ordinance players can go into a degree of detail to determine who may be effected and the manner in which they will be effected. 




The primary aim for the ordance screen is to provide the player with all remaining options to modify their city that are not available to them in the other screens such as taxation, funding, building, etc. Its impact on economy is that it is enacting laws or directing public money down avenues that effect business and population directly through bans or similar or indirectly through workforce "consumption" options and limitations.

To implement ordinance players select/navigate to the ordinance screen from which there appears a list of the ordinance available that can be filtered/sorted to either zonal types or workforce types. Players then select a specific ordance they wish to utilize. Many ordinance have different processes to being implemented, some may be as simple as selecting the ordinance, others may allow for varying degrees of detail, allowing players to micro manage the effects of the ordinance or create multiple types of the ordinance. 

Micro management of ordinance involves selecting the ordinance then selecting the factors which it will influence. For starters this may be a zonal factor such as all commericial zoning. It could be limitted further to be all exporting commercially zoned businesses. Or further still to be all businesses exporting some resource, or even businesses exporting an amount of this resource to a certain city. The same could be done for workforce types, such as impacting all households with low income, or all households with low income and a single parents. Another way ordinance may be needed to be inacted is to be within certain geographical limits, this may be cities within a certain area, specific cities, or even certain areas of the players city. Geographical areas can also be limited to certain structural sites such as stadiums or train stations or starports or a variety of these. As mentioned, the micro management type and degree is unique and dependant on the ordinance itself.

The limits of ordinance are not to be confused with taxation by there effect. While taxation deals primarily with income types or wealth, ordinance deals with inacting laws or services. In the above example with the commercial business "Conglamerate Hotcakes" exporting 2 million "Martian Jaffa Waffles" to earth, this business may be required to have all shareholders attend dietary health counselling. In this example we can see that it forces the business performing an action under condition of ordinance to do something, namely dietary counselling. In the other example with the sinlge mother with two sons, the family may be given a certain amount of house cleaning service provided per month. In this way we can see that ordinance is providing what is required be a section of the city without making this happen through funds.  


%The various types of ordinance the game requires will become more evident as development progresses. This list may need to be work on by various people on the project. This list will be added to, taken from, and refined as development progresses.

%Possible Ordinance
-green import/export.
-ban import/export type/good/company.
-city volunteer projects (various).
-firearm limitations/regulations
-water regulations
-waste regulations
-engery use regulations
-farming initiatives/regulations
-food production regulations
-fire safety programs
-neighboud watch or similar
-farmers markets or similar
-community engagement programs
-public housing/free housing
-rental rates / housing rates / commision levels
-various housing affordable limits
-food stamps
-free food / other food related issues
-various education types (will effect workforce)
-clean air
-organic/biodynamic foods
-government loans or zero debt loans or similar
-anti-monopoly laws / pro-monopoly laws
-driving laws
-vehicle emmission / fuel types 
-resource use limits
-energy limits/types
-legal system admin / free legal aid
-law enforcement bodies
-martial law orders
-zoning regulations
-business trading hours
-noise polution laws
-enforcement bodies/zones
-public transport issues / types / enforcement / costs
-import / export limits, local, earth
-garbage / recycling / etc
-goods reuse / business buy back or similar
-Parking
-public areas / pets / walking / traffic
-heating
-noise levels
-health care, various.
-city amenities, various.
-import standards.
-export standards.


